### 2017-07-04 LaTeX/Paper.tex
Created first draft of paper about the program.

\documentclass[11pt,a4paper]{article}
\usepackage{termpaper}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{caption}

%opening
\title{<TITLE PENDING>}
\author{
 \authorname{Jakob Klinger} \\
 \studentnumber{1125755} \\
 \curriculum{033 534} \\
 \email{e1125755@student.tuwien.ac.at}
}

\begin{document}

\maketitle

%TODO: Abstract (\begin{abstract}...\end{abstract}

\section{Introduction}
Annotating a document is usually solved by adding footnotes or figures in an appropriate position and adding a simple reference in the text, leaving the reader to find the referenced content by themselves. Sometimes however, if a more obvious connection between the text and the referenced content is required, the reference is visibly connected to the text by drawing a line between them.

In this paper, we will look at ways to use Boundary Labeling, which means that all annotations will be placed outside of the text they are referencing. (See also \cite{Bekos2007}) While there are many papers discussing Boundary Labeling in general, only very few exist that apply this concept to written text. As a result, and since it is very easy to accidentally obscure parts of the text, this approach isn't used very often, and tends to use simplistic algorithms which produce mediocre results.

%TODO: Was genau wollen wir erreichen?
%TODO: Sektion: Terminologie

%Section Ideas: Leader Types, The Program/Framework, The Algorithm(s),

\bibliographystyle{plain}
\bibliography{references}

\end{document}

### 2017-07-05 LaTeX/Paper.tex
Extended Paper. No changes to Code.

@@
 %opening
-\title{<TITLE PENDING>}
+\title{Boundary Labeling for annotated documents}%Vielleicht genauerer Titel?
 \author{
@@
 
-%TODO: Abstract (\begin{abstract}...\end{abstract}
+%TODO: Abstract: \begin{abstract}...\end{abstract}
 
@@
 
+\section{Boundary Labeling}
+Boundary Labeling can be explained very quickly -  connect a Point of Interest with a box containing further information which is located outside of the main area.
+To easily reference important concepts, some additional Terminology will be introduced:
+
+\begin{itemize}
+\item \textbf{Label:} The additional information, usually represented as a box containing the additional information. Will also be referred to as \textbf{``Annotation''} in this paper.
+\item \textbf{Site:} The point, object or word that is labeled - without it, there wouldn't be any labeling necessary.%Is that the correct name?
+\item \textbf{Leader:} The line connecting the site to the label - depending on its shape, it can be further classified into several subgroups.
+\end{itemize}
+
+\subsection{Boundary Labeling in Documents}
+As with other labeling techniques, Boundary Labeling has guidelines on how to create an optimal labeling - the leaders should be as direct as possible, no important information should be obscured, and it should be easily discernable which Label belongs to which site.
+These three often stand in conflict with one another, especially when labeling documents, as the text usually is very dense and leaves little space for leaders in between, yet one shouldn't allow them to pass through the text, as this makes the text harder to read.
+
+%TODO: Über bisherige Versuche/Papers sprechen
 %TODO: Was genau wollen wir erreichen?

### 2017-07-06 Eclipse/Labeling/src/routingapp/GraphCreatorPanel.java
Backing up notes from last presentation

@@ -43,7 +43,7 @@ public class GraphCreatorPanel extends JComponent {
 
 	//DEBUG VALUES
 	private boolean showWordBoundaries=false;//draws rectangles around detected word boundaries in main text, if set to true
-	private boolean showGraphGrid=false;//Draws the whole routing Graph 
+	private boolean showGraphGrid=true;//Draws the whole routing Graph 
 	//DEBUG VALUES END
 
 	private int width=600;

### 2017-07-17 LaTeX/Paper.tex
Backing up notes from last presentation
Added Drawing for Terminology section of the Paper.

@@
 \usepackage{caption}
+\usepackage{wrapfig}
+%\usepackage{placeins}
 
@@
 
+%TODO: Bezug auf Programm & Ergebnisse nehmen!
+
 \section{Boundary Labeling}
 Boundary Labeling can be explained very quickly -  connect a Point of Interest with a box containing further information which is located outside of the main area.
-To easily reference important concepts, some additional Terminology will be introduced:
+To easily reference important concepts, some additional Terminology will be introduced (See Fig.~\ref{fig:term} for a visual explanation)
+
+\begin{wrapfigure}{r}{0.35\textwidth}
+ \captionsetup{justification=centering, margin=0.75cm}
+ \raggedleft
+  \includegraphics[scale=0.5]{IPE_TerminologyDrawing.pdf}
+  \caption{Test}
+ \label{fig:term}
+\end{wrapfigure}
 
@@
  \item \textbf{Label:} The additional information, usually represented as a box containing the additional information. Will also be referred to as \textbf{``Annotation''} in this paper.
-\item \textbf{Site:} The point, object or word that is labeled - without it, there wouldn't be any labeling necessary.%Is that the correct name?
+ \item \textbf{Site:} The point, object or word that is labeled - without it, there wouldn't be any labeling necessary.
  \item \textbf{Leader:} The line connecting the site to the label - depending on its shape, it can be further classified into several subgroups.
+ \item \textbf{Port:} The location where the leader connects to the label. It may be restricted pre-determined locations, like only at the corners.
 \end{itemize}
+
+%\FloatBarrier
 
@@
 These three often stand in conflict with one another, especially when labeling documents, as the text usually is very dense and leaves little space for leaders in between, yet one shouldn't allow them to pass through the text, as this makes the text harder to read.
+%Beispiele von suboptimalen Lösungen einfügen (Bilder)
 
@@
 %TODO: Was genau wollen wir erreichen?
-%TODO: Sektion: Terminologie
 
-%Section Ideas: Leader Types, The Program/Framework, The Algorithm(s),
+%Section Ideas: Leader Types, The Program/Framework (Modellerklärung), The Algorithm(s), Implementation, Evaluation, Conclusion
 

### 2017-07-21 LaTeX/Paper.tex
Paper: Expanded Terminology section and started section about previous research. Program: No changes.

@@
 \usepackage{wrapfig}
-%\usepackage{placeins}
 
@@
   \includegraphics[scale=0.5]{IPE_TerminologyDrawing.pdf}
-  \caption{Test}
+  \caption{Illustrated guide to the labeling terminology}
  \label{fig:term}
@@
  \item \textbf{Site:} The point, object or word that is labeled - without it, there wouldn't be any labeling necessary.
- \item \textbf{Leader:} The line connecting the site to the label - depending on its shape, it can be further classified into several subgroups.
+ \item \textbf{Leader:} The line connecting the site to the label - depending on its shape, it can be further classified into several subgroups which can be freely combined to form new leader types:
+  \begin{itemize}
+   \item \textit{S-Leader:} This Leader connects site and label in a straight line.
+   \item \textit{O-Leader:} This leader runs orthogonally to the border between text and label area.
+   \item \textit{P-Leader:} This Leader runs parallel to the border between text and label area. It must be combined with other leader styles, as it won't reach the label area otherwise.
+  \end{itemize}
+    For example, the leader from Fig.~\ref{fig:term} would be classified as an opo-Leader.
  \item \textbf{Port:} The location where the leader connects to the label. It may be restricted pre-determined locations, like only at the corners.
@@
 
-%\FloatBarrier
+\newpage
 
@@
 These three often stand in conflict with one another, especially when labeling documents, as the text usually is very dense and leaves little space for leaders in between, yet one shouldn't allow them to pass through the text, as this makes the text harder to read.
-%Beispiele von suboptimalen Lösungen einfügen (Bilder)
+%Beispiele von suboptimalen Lösungen einfügen? (Bilder)
+
+\section{Previous Attempts}%Sub(sub)section machen?
+
+While there haven't been many papers written about this topic, they do offer interesting contributions.
+For example, the paper\cite{Kindermann2014} about the Luatodonotes-Package illustrates some of the different leader types available, and came to the conclusion that leaders without bends are easier to follow.
+%Bild einfügen?
 

### 2017-07-24 LaTeX/Paper.tex
Paper: Added second paper to section "Previous Work". Program: No changes.

@@
 While there haven't been many papers written about this topic, they do offer interesting contributions.
-For example, the paper\cite{Kindermann2014} about the Luatodonotes-Package illustrates some of the different leader types available, and came to the conclusion that leaders without bends are easier to follow.
+For example, the paper about the Luatodonotes-Package\cite{Kindermann2014} illustrates some of the different leader types available, and came to the conclusion that leaders without bends are easier to follow.
 %Bild einfügen?
+However, most solutions proposed in that paper didn't care whether a leader overlapped with text or not, which resulted in a decrease in readability.
 
-%TODO: Über bisherige Versuche/Papers sprechen
+The paper by Loose\cite{Loose2015} on the other hand is based around only using the free space between the text, which produces slightly longer leaders, and forces curves.
+In short, the Leaders can be described as a sequence of O- and P-segments, which continue until the label's port is reached.
+%Überleitung zu eigenem Programm?
+Viele Konzepte wurden hier wiederverwendet!
+
 %TODO: Was genau wollen wir erreichen?

### 2017-08-03 LaTeX/Paper.tex
Paper: Moved Section "Previous Work" to introduction, rewrote it to replace the word "Leader" with more neutral terminology. Program: No relevant changes, only edited for demonstration.

--- 2017-07-21 LaTeX/Paper.tex
@@
 Annotating a document is usually solved by adding footnotes or figures in an appropriate position and adding a simple reference in the text, leaving the reader to find the referenced content by themselves.
-Sometimes however, if a more obvious connection between the text and the referenced content is required, the reference is visibly connected to the text by drawing a line between them.
+Sometimes however, if a more obvious connection between the text and the referenced content is required, the reference is visibly connected to the text by drawing a straight line or a more complex path between them.
 
-In this paper, we will look at ways to use Boundary Labeling, which means that all annotations will be placed outside of the text they are referencing.
+In this paper, we will look at ways to use Boundary Labeling, which means that all annotations will be placed outside of the text they are referencing and will be visually connected to the feature they are referencing.
 (See also \cite{Bekos2007}) While there are many papers discussing Boundary Labeling in general, only very few exist that apply this concept to written text.
 As a result, and since it is very easy to accidentally obscure parts of the text, this approach isn't used very often, and tends to use simplistic algorithms which produce mediocre results.
+However, the papers that do discuss boundary labeling in text offer interesting contributions.
+
+For example, the paper about the Luatodonotes-Package\cite{Kindermann2014} illustrates some of the different styles of drawing these connecting paths, and came to the conclusion that paths without bends are easier to follow.
+%Bild einfügen?
+However, most solutions proposed in that paper don't care whether a path overlapped with text or not, which results in a decrease in readability.
+
+The paper by Loose\cite{Loose2015} on the other hand is based around only using the free space between lines and words, which produces longer paths, and forces curves, but doesn't obscure any part of the text.
+%Siehe auhc andere Papers - Motivation! (Warum will ich mir dieses Thema ansehen)
 
@@
 
-\section{Boundary Labeling}
+\section{Boundary Labeling} %Evtl in Introduction eingliedern?
 Boundary Labeling can be explained very quickly -  connect a Point of Interest with a box containing further information which is located outside of the main area.
@@
  \item \textbf{Site:} The point, object or word that is labeled - without it, there wouldn't be any labeling necessary.
- \item \textbf{Leader:} The line connecting the site to the label - depending on its shape, it can be further classified into several subgroups which can be freely combined to form new leader types:
+ \item \textbf{Leader:} The path connecting the site to the label - depending on its shape, it can be further classified into several subgroups which can be freely combined to form new leader types:
   \begin{itemize}
@@
 \end{itemize}
-
-\newpage
+%Def. für Graphen hinzufügen! - Bei Koordinaten: Anmerken, dass 2d-euklidische Geometrie benutzt wird!
 
@@
 As with other labeling techniques, Boundary Labeling has guidelines on how to create an optimal labeling - the leaders should be as direct as possible, no important information should be obscured, and it should be easily discernable which Label belongs to which site.
-These three often stand in conflict with one another, especially when labeling documents, as the text usually is very dense and leaves little space for leaders in between, yet one shouldn't allow them to pass through the text, as this makes the text harder to read.
+These three easily come into conflict with one another, especially when labeling documents, as the text usually is very dense and leaves little space for leaders in between, yet one shouldn't allow them to pass through the text, as this makes the text harder to read.
+Therefore, each implementation must come to a compromise that balances these criteria.
 %Beispiele von suboptimalen Lösungen einfügen? (Bilder)
 
-\section{Previous Attempts}%Sub(sub)section machen?
-
-While there haven't been many papers written about this topic, they do offer interesting contributions.
-For example, the paper\cite{Kindermann2014} about the Luatodonotes-Package illustrates some of the different leader types available, and came to the conclusion that leaders without bends are easier to follow.
-%Bild einfügen?
-
-%TODO: Über bisherige Versuche/Papers sprechen
 %TODO: Was genau wollen wir erreichen?
 
-%Section Ideas: Leader Types, The Program/Framework (Modellerklärung), The Algorithm(s), Implementation, Evaluation, Conclusion
+%Section Ideas: The Program/Framework (Modellerklärung), The Algorithm(s), Implementation, Evaluation, Conclusion
 

### 2017-08-03 Eclipse/Labeling/src/routingapp/GraphCreatorPanel.java

@@ -55,7 +55,7 @@ public class GraphCreatorPanel extends JComponent {
 	private int spaceBetweenAnnLines=5;
 	private Font AnnotationFont;
 	
-	private int curveSize=3;
+	private int curveSize=6;
 
 	private int nextAnnotationPos=0;
 	private String routingtype;//Determines the type of routing - see also GraphCreatorPanel.getRouter() and GraphCreatorApplet.algNames[]

### 2017-08-04 LaTeX/Paper.tex
Paper: Expanded Terminology section. Program: No changes.

@@
 
-\section{Boundary Labeling} %Evtl in Introduction eingliedern?
-Boundary Labeling can be explained very quickly -  connect a Point of Interest with a box containing further information which is located outside of the main area.
-To easily reference important concepts, some additional Terminology will be introduced (See Fig.~\ref{fig:term} for a visual explanation)
+\section{Terminology and Fundamentals} %Evtl in Introduction eingliedern?
+While Boundary Labeling (or an equivalent concept) can be applied to a space with different geometry or more dimensions, this paper will only concern itself with two-dimensonal, euclidean space.
+To easily reference important concepts, some additional Terminology will be introduced as well. (See Fig.~\ref{fig:term} for a visual explanation)%Add more figures for graph-terms, etc. or change
 
@@
 \begin{itemize}
- \item \textbf{Label:} The additional information, usually represented as a box containing the additional information. Will also be referred to as \textbf{``Annotation''} in this paper.
+ \item \textbf{Graph:} A Set of Vertices and Edges (see below), often used to represent routing problems. Can be further classified depending on its properties.
+ \item \textbf{Edge:} A path connecting two vertices, usually represented as a line. Can be directional, limiting its usage, or have a cost (``Weight'') associated with its use.
+ \item \textbf{Vertex:} A featureless object that can be connected to any number other vertices via edges.
+Usually depicted as a point or small shape.
+Commonly also referred to as \textbf{``Node''}, as will be the the case in this paper.
+ \item \textbf{Label:} Additional information, usually represented as a box containing the additional information. Will also be referred to as \textbf{``Annotation''} in this paper.
  \item \textbf{Site:} The point, object or word that is labeled - without it, there wouldn't be any labeling necessary.
- \item \textbf{Leader:} The path connecting the site to the label - depending on its shape, it can be further classified into several subgroups which can be freely combined to form new leader types:
+ \item \textbf{Leader:} The path connecting the site to the label - depending on its shape, it can be described further using several subgroups which can be freely combined to form new leader types:
   \begin{itemize}
@@
 
-\subsection{Boundary Labeling in Documents}
+\section{Boundary Labeling in Documents}%Rework this
 As with other labeling techniques, Boundary Labeling has guidelines on how to create an optimal labeling - the leaders should be as direct as possible, no important information should be obscured, and it should be easily discernable which Label belongs to which site.

### 2017-08-10 LaTeX/Paper.tex
Paper: Started section about the program. Program: No changes.

@@
 While Boundary Labeling (or an equivalent concept) can be applied to a space with different geometry or more dimensions, this paper will only concern itself with two-dimensonal, euclidean space.
-To easily reference important concepts, some additional Terminology will be introduced as well. (See Fig.~\ref{fig:term} for a visual explanation)%Add more figures for graph-terms, etc. or change
+To easily reference important concepts, some additional terminology will be introduced as well. (See Fig.~\ref{fig:term} for a visual explanation)%Add more figures for graph-terms, etc. or change
 
@@
 \end{itemize}
-%Def. für Graphen hinzufügen! - Bei Koordinaten: Anmerken, dass 2d-euklidische Geometrie benutzt wird!
+%Mehr Graphenterminologie? (zb: Path)
 
@@
 Therefore, each implementation must come to a compromise that balances these criteria.
+
+%neue Section?
+In our program, %my program?
+the focus was put on keeping the text as readable as possible, which means leaders are only allowed in the space between words or lines.
+We also wanted to keep the leaders  as short as possible, given a fixed label and port position, which means that the leaders have to be monotonous.
+%Vorgehensweise erläutern
+%(Potential) Problems/Issues, solutions
 %Beispiele von suboptimalen Lösungen einfügen? (Bilder)

### 2017-08-22 LaTeX/Paper.tex
Paper: Re-worded last bit, started section about the inplementation. Program: No changes.

@@
 Therefore, each implementation must come to a compromise that balances these criteria.
+%Beispiele von suboptimalen Lösungen einfügen? (Bilder)
 
@@
 In our program, %my program?
-the focus was put on keeping the text as readable as possible, which means leaders are only allowed in the space between words or lines.
-We also wanted to keep the leaders  as short as possible, given a fixed label and port position, which means that the leaders have to be monotonous.
+the focus was put on keeping the text as readable as possible.
+This means that leaders are only allowed in the space between words or lines, which can make certain locations for sites unavailable, as there might already too many leaders routed through that area to insert yet another for the new site.
+We also wanted the leaders to be monotonous, which means that they will be one of the shortest possible connection between port and site locations.
+Our final priority was to efficiently use the space for the labels, which meant that each label should initially be placed as close to the available area's border as possible.
+While this allows for more space in which labels can be placed, it also means that the distance between Label and site often increases, which leads to longer leaders, and often more curves as well.
+
+%Subsetcion?
+\section{Implementation}
+The program was written in Java, using only JGraphT%Source & How do I link to this?
+as additional Library.
+Similar to Loose's Approach\cite{Loose2015}, we used a Graph to model the available routing space for the algorithm.
+As a result of this approach, it is important to know exactly where each word is placed.
+While Java offers a function that automatically places text in lines of given width, it does so in a way that doesn't allow the content of each line to be read as a String afterwards, which makes it unusable for our purposes.
+Therefore each word must be placed manually, and to since it isn't the main focus of the program, each series of characters between a space is considered one word that has to be fully placed within one line.
+
 %Vorgehensweise erläutern
 %(Potential) Problems/Issues, solutions
-%Beispiele von suboptimalen Lösungen einfügen? (Bilder)
 

### 2017-08-23 LaTeX/Paper.tex
Paper: Cut Section 3 and added it to the introduction, extended Section about the program. Program: Tried to smoothen curves that are located close to each other, only seemed to work for the OPO-Leaders.

@@
 \usepackage{caption}
-\usepackage{wrapfig}
+%\usepackage{wrapfig}
 
@@
 In this paper, we will look at ways to use Boundary Labeling, which means that all annotations will be placed outside of the text they are referencing and will be visually connected to the feature they are referencing.
-(See also \cite{Bekos2007}) While there are many papers discussing Boundary Labeling in general, only very few exist that apply this concept to written text.
-As a result, and since it is very easy to accidentally obscure parts of the text, this approach isn't used very often, and tends to use simplistic algorithms which produce mediocre results.
+(See also \cite{Bekos2007})
+
+The guidelines on how to create an optimal labeling are as follows: the connections should be as direct as possible, no important information should be obscured, and it should be easily discernable which Label belongs to which site.
+These three easily come into conflict with one another, especially when labeling documents, as the text usually is very dense and leaves little space for lines in between, yet one shouldn't allow them to pass through the text, as this makes the text harder to read.
+
+While there are many papers discussing Boundary Labeling in general, only very few exist that apply this concept to written text.
+Generally, this approach isn't used very often, and tends to use simplistic algorithms which produce mediocre results.
+%Beispiele von suboptimalen Lösungen einfügen?
+(Bilder)
 However, the papers that do discuss boundary labeling in text offer interesting contributions.
@@
 
-\begin{wrapfigure}{r}{0.35\textwidth}
+\begin{figure}%{r}{0.35\textwidth}
  \captionsetup{justification=centering, margin=0.75cm}
- \raggedleft
+ \centering
   \includegraphics[scale=0.5]{IPE_TerminologyDrawing.pdf}
@@
  \label{fig:term}
-\end{wrapfigure}
+\end{figure}
 
@@
 
-\section{Boundary Labeling in Documents}%Rework this
-As with other labeling techniques, Boundary Labeling has guidelines on how to create an optimal labeling - the leaders should be as direct as possible, no important information should be obscured, and it should be easily discernable which Label belongs to which site.
-These three easily come into conflict with one another, especially when labeling documents, as the text usually is very dense and leaves little space for leaders in between, yet one shouldn't allow them to pass through the text, as this makes the text harder to read.
-Therefore, each implementation must come to a compromise that balances these criteria.
-%Beispiele von suboptimalen Lösungen einfügen? (Bilder)
+\section{The program}
 
-%neue Section?
 In our program, %my program?
@@
 Our final priority was to efficiently use the space for the labels, which meant that each label should initially be placed as close to the available area's border as possible.
-While this allows for more space in which labels can be placed, it also means that the distance between Label and site often increases, which leads to longer leaders, and often more curves as well.
+While this allows for more space in which labels can be placed, it also means that the distance between Label and site often increases, which leads to longer leaders, and often more curves as well.%''Final Priority'' klingt komisch -> Ersatz finden!
 
-%Subsetcion?
-\section{Implementation}
-The program was written in Java, using only JGraphT%Source & How do I link to this?
+\subsection{Implementation}
+The program was written in Java, using only JGraphT %Source & How do I link to this?
 as additional Library.
@@
 Therefore each word must be placed manually, and to since it isn't the main focus of the program, each series of characters between a space is considered one word that has to be fully placed within one line.
+
+Using the dimensions of each word, we create the graph by creating and connecting nodes in the space between lines at the start and end of each word, and at the end of the text area.
+We also create special nodes for each annotated word, they are located above the word in a central position.
+Each node also is located at set co-ordinates, which allows the routing algorithms to easily fulfill the monotonicity constraint.
+We left some space in between the text area and the label area %Name=?
+to allow the labels to be placed more freely.
+While the routing graph only reaches up to the border of the text area, the connection between the graph and the label is done with an OPO-Leader.
+As there is the possibility of intersecting with other leaders, this means some extra adjustments have to be done.
+For demonstration purposes, we also included another algorithm that uses S-Leaders for this section.
 

### 2017-08-23 Eclipse/Labeling/src/routingapp/GraphCreatorPanel.java

@@ -489,6 +489,7 @@ public class GraphCreatorPanel extends JComponent {
 		}
 		//Draw complete or unfinished Route
 		//NOTE: Only tested with paths that go up and to the right, although it should be capable of handling any kind of path.
+		//NOTE: OPO-Leaders currently only support going up!
 		
 		Graphics2D g2d=(Graphics2D)g;
 
@@ -531,7 +532,7 @@ public class GraphCreatorPanel extends JComponent {
 							if(previewTuple.getX()==nextTuple.getX())//<=>2 Curves close to each other
 							{
 								route.lineTo(currentTuple.getX(),currentTuple.getY()+curveSize/2*Math.signum(oldTuple.getY()-currentTuple.getY()));
-								route.curveTo(currentTuple.getX(),currentTuple.getY(), nextTuple.getX(), nextTuple.getY(), nextTuple.getX(), nextTuple.getY()+curveSize/2*Math.signum(previewTuple.getY()-nextTuple.getY()));
+								route.curveTo(currentTuple.getX(),currentTuple.getY()-curveSize/2*Math.signum(oldTuple.getY()-currentTuple.getY()), nextTuple.getX(), nextTuple.getY()+curveSize/2*Math.signum(previewTuple.getY()-nextTuple.getY()), nextTuple.getX(), nextTuple.getY()+curveSize/2*Math.signum(previewTuple.getY()-nextTuple.getY()));
 							}
 							else if(Math.abs(nextTuple.getX()-currentTuple.getX())<=curveSize)//nextTuple is too close to the curve - we will just go past it
 							{
@@ -567,6 +568,7 @@ public class GraphCreatorPanel extends JComponent {
 			}
 			
 			//Connection from Graph to the Annotation
+			//NOTE: This assumes OPO-Leaders that go upwards!
 			if(route.getCurrentPoint().getX()<rightTextBorder)//Compensating for right-to-left routed paths
 			{
 					route.moveTo(nodeList.get(0).getX(), nodeList.get(0).getY());
@@ -576,12 +578,12 @@ public class GraphCreatorPanel extends JComponent {
 			{
 				route.lineTo(leftAnnotationBorder,currentAnnotationPos);
 			}
-			else if(Math.abs(info.getOpoStart()-info.getOpoEnd())<2*curveSize)
[…]

### 2017-09-06 LaTeX/Paper.tex
Paper: Split section 3 into theoretical algorithm and implementatin details, currently unwinished. Program: No changes.

@@
  \item \textbf{Graph:} A Set of Vertices and Edges (see below), often used to represent routing problems. Can be further classified depending on its properties.
- \item \textbf{Edge:} A path connecting two vertices, usually represented as a line. Can be directional, limiting its usage, or have a cost (``Weight'') associated with its use.
+ \item \textbf{Edge:} A connection between two vertices, usually represented as a line. Can be directional, limiting its usage, or have a cost (``Weight'') associated with its use.
  \item \textbf{Vertex:} A featureless object that can be connected to any number other vertices via edges.
@@
  \item \textbf{Port:} The location where the leader connects to the label. It may be restricted pre-determined locations, like only at the corners.
+  \item \textbf{Monotonous/Monotonicity:} Describes a steady progression towards a direction or goal.
+Will be used in this paper to indicate that there's no loss of progress made from the site to the label, effectively restricting the direction a leader can go at each bend.
 \end{itemize}
-%Mehr Graphenterminologie? (zb: Path)
+%Mehr Graphenterminologie? (zb: Path) +Polylinien Vertex/Node für Knoten in Polylinien bzw. Graphen -->Eindeutig!
 
@@
 
-In our program, %my program?
-the focus was put on keeping the text as readable as possible.
+In our program, the focus was put on keeping the text as readable as possible.
 This means that leaders are only allowed in the space between words or lines, which can make certain locations for sites unavailable, as there might already too many leaders routed through that area to insert yet another for the new site.
@@
 Our final priority was to efficiently use the space for the labels, which meant that each label should initially be placed as close to the available area's border as possible.
-While this allows for more space in which labels can be placed, it also means that the distance between Label and site often increases, which leads to longer leaders, and often more curves as well.%''Final Priority'' klingt komisch -> Ersatz finden!
+While this allows for more space in which labels can be placed, it also means that the distance between Label and site often increases, which leads to longer leaders.%''Final Priority'' klingt komisch -> Ersatz finden!
+
+%Algorithmus(Ansatz) und Implementierung(noch immer abstrakt) trennen! Probleme, Details, ... gegen Ende!
+\subsection{Algorithm}%''Underlying Algorithm''?
+To easily create paths that exclusively use the space between words and lines, we decided to use a graph similar to the one Loose \cite{Loose2015} used, placing nodes above and below each word's beginning and end, and connecting them with their horizontally and vertically adjacent neighbours.%Umformulieren?
+For the sites, we inserted an additional node above the center of the word, which will serve as the leader's starting point.
+We also decided to give the graph's edges a capacity of 1, which means that each leader will always use edges which haven't already be taken by other leaders.
+While this causes some sites to become un-routable due to another leader passing right through their starting node, it makes visualizing the leaders much easier.
+Our first approach was going through the labels in the order they appear in the text and placing each as far up as possible, to maximize the remaining space other labels could use.
+
+Since putting the labels right next to the text would result in the labels' positions being restricted by the text's line spacing,we decided to leave a buffer zone in between the two areas, which allows the leaders to adjust between the position forced by the line spacing and the label's final position.
+As this allowed us to place the labels more freely, we adjusted the label's positions after all labels were placed to reduce the distance that needed to be covered in the buffer area.
 
 \subsection{Implementation}
-The program was written in Java, using only JGraphT %Source & How do I link to this?
-as additional Library.
-Similar to Loose's Approach\cite{Loose2015}, we used a Graph to model the available routing space for the algorithm.
+The program was written in Java, using only JGraphT\cite{JGraphT} as additional library.
+Since we only want to create leaders that don't intersect with the text, the graph was created alongside the placement of the words on the canvas, as it was easy to extract measurements at this point.
+Due to some problems with Java's various methods of calculating linebreaks, this process was done completely by hand.
+Since the Graph's nodes represent fixed locations on the canvas, they each have coordinates associated with them, with the nodes representing the sites containing extra information, such as references to the corresponding annotation and to the leader connecting the two (if existent).
+%TODO: Code genauer durchlesen und erweitern!
+
+%Similar to Loose's Approach\cite{Loose2015}, we used a Graph to model the available routing space for the algorithm.
 As a result of this approach, it is important to know exactly where each word is placed.
@@
 
-Using the dimensions of each word, we create the graph by creating and connecting nodes in the space between lines at the start and end of each word, and at the end of the text area.
+%Using the dimensions of each word, we create the graph by creating and connecting nodes in the space between lines at the start and end of each word, and at the end of the text area.
 We also create special nodes for each annotated word, they are located above the word in a central position.
 Each node also is located at set co-ordinates, which allows the routing algorithms to easily fulfill the monotonicity constraint.
-We left some space in between the text area and the label area %Name=?
-to allow the labels to be placed more freely.
+%We left some space in between the text area and the label area %Name=?(Lass dir was einfallen)
+%to allow the labels to be placed more freely.
 While the routing graph only reaches up to the border of the text area, the connection between the graph and the label is done with an OPO-Leader.
@@
 
-%Vorgehensweise erläutern
 %(Potential) Problems/Issues, solutions
-
-%TODO: Was genau wollen wir erreichen?
 

### 2017-09-10 LaTeX/Paper.tex
Paper: Expanded terminology section and re-wrote other sections that used newly defined words in different contexts. Expanded Section 3. Program: No changes.

@@
 
-\section{Introduction}
+\section{Introduction}%Wort ``Path'' durch nicht-definierte Terminologie ersetzen?
 Annotating a document is usually solved by adding footnotes or figures in an appropriate position and adding a simple reference in the text, leaving the reader to find the referenced content by themselves.
@@
 While Boundary Labeling (or an equivalent concept) can be applied to a space with different geometry or more dimensions, this paper will only concern itself with two-dimensonal, euclidean space.
-To easily reference important concepts, some additional terminology will be introduced as well. (See Fig.~\ref{fig:term} for a visual explanation)%Add more figures for graph-terms, etc. or change
+To easily reference important concepts, some additional terminology will be introduced as well. (See Fig.~\ref{fig:term} for a visual explanation)%Add references to all terminology pictures!
+
+%Add figure for graph terminology.
 
@@
 Usually depicted as a point or small shape.
-Commonly also referred to as \textbf{``Node''}, as will be the the case in this paper.
+Also referred to as \textbf{``Node''}.%Remove second sentence and re-define node?
+ \item \textbf{Path:} An ordered series of either edges that share one vertex each with the edges listed before and after them, or an ordered series of vertices that have an edge connecting each vertex to the subsequent one.
+Also known as \textbf{``Graph Walk''}.
+
+ \item \textbf{Polyline:} A connected series of straight lines.%Ist mehr Information notwendig?
+
  \item \textbf{Label:} Additional information, usually represented as a box containing the additional information. Will also be referred to as \textbf{``Annotation''} in this paper.
  \item \textbf{Site:} The point, object or word that is labeled - without it, there wouldn't be any labeling necessary.
- \item \textbf{Leader:} The path connecting the site to the label - depending on its shape, it can be described further using several subgroups which can be freely combined to form new leader types:
+ \item \textbf{Leader:} The line or polyline connecting the site to the label - depending on its shape, it can be described further using several subgroups which can be freely combined to form new leader types:
   \begin{itemize}
@@
     For example, the leader from Fig.~\ref{fig:term} would be classified as an opo-Leader.
- \item \textbf{Port:} The location where the leader connects to the label. It may be restricted pre-determined locations, like only at the corners.
+ \item \textbf{Port:} The location where the leader connects to the label. It may be restricted to pre-determined locations, like only at the corners.
   \item \textbf{Monotonous/Monotonicity:} Describes a steady progression towards a direction or goal.
 Will be used in this paper to indicate that there's no loss of progress made from the site to the label, effectively restricting the direction a leader can go at each bend.
+  \item \textbf{Depth-first search:} A searching algorithm that continues along a set of rules as far as possible until either the goal is found, or backtracking is required.
+A common example is navigating a maze by turning the same way whenever possible.
 \end{itemize}
@@
 
-%Algorithmus(Ansatz) und Implementierung(noch immer abstrakt) trennen! Probleme, Details, ... gegen Ende!
+%Port-Location erwähnen!
 \subsection{Algorithm}%''Underlying Algorithm''?
-To easily create paths that exclusively use the space between words and lines, we decided to use a graph similar to the one Loose \cite{Loose2015} used, placing nodes above and below each word's beginning and end, and connecting them with their horizontally and vertically adjacent neighbours.%Umformulieren?
-For the sites, we inserted an additional node above the center of the word, which will serve as the leader's starting point.
+To easily create leaders that exclusively use the space between words and lines, we decided to use a graph similar to the one Loose \cite{Loose2015} used, placing vertices above and below each word's beginning and end, and connecting them with their horizontally and vertically adjacent neighbours.%Umformulieren?
+For the sites, we inserted an additional vertex above the center of the word, which will serve as the leader's starting point.
 We also decided to give the graph's edges a capacity of 1, which means that each leader will always use edges which haven't already be taken by other leaders.
-While this causes some sites to become un-routable due to another leader passing right through their starting node, it makes visualizing the leaders much easier.
+While this causes some sites to become un-routable due to another leader passing right through their starting vertex, it makes visualizing the leaders much easier.
+
 Our first approach was going through the labels in the order they appear in the text and placing each as far up as possible, to maximize the remaining space other labels could use.
-
-Since putting the labels right next to the text would result in the labels' positions being restricted by the text's line spacing,we decided to leave a buffer zone in between the two areas, which allows the leaders to adjust between the position forced by the line spacing and the label's final position.
+We also opted to use fixed ports located in the top left corner of the annotation, as this allowed us to unambiguously place any annotation by knowing either their port's location, wich is also the leader's ending point.
+However, since putting the labels right next to the text would result in the labels' positions being restricted by the text's line spacing, we then decided to leave a buffer zone in between the two areas, which allows the leaders to adjust between the position forced by the line spacing and the label's final position.
 As this allowed us to place the labels more freely, we adjusted the label's positions after all labels were placed to reduce the distance that needed to be covered in the buffer area.
@@
 Due to some problems with Java's various methods of calculating linebreaks, this process was done completely by hand.
-Since the Graph's nodes represent fixed locations on the canvas, they each have coordinates associated with them, with the nodes representing the sites containing extra information, such as references to the corresponding annotation and to the leader connecting the two (if existent).
+
+As the Graph's vertices represent fixed locations on the canvas, they each have coordinates associated with them, with the vertices representing the sites containing extra information, such as references to the corresponding annotation and to the leader connecting the two (if existent).
+This additional information is used for both routing and drawing the leaders: The routing was done via a depth-first search algorithm that prioritized vertices located further above if possible, and vertices to the right otherwise.
+After the algorithm terminates, it returns  a vertex-based graph walk describing the route from the site to the border of the text area alongside information about where to draw the OPO-segment in the buffer zone if successful, or a walk containing only the starting vertex, if no valid path was found.
+Using this information alongside the positional data stored in the vertices, the polyline representing the leader can be easily drawn.
+%TODO: Schönheitsfeatures, etc.
 %TODO: Code genauer durchlesen und erweitern!
 
-%Similar to Loose's Approach\cite{Loose2015}, we used a Graph to model the available routing space for the algorithm.
-As a result of this approach, it is important to know exactly where each word is placed.
-While Java offers a function that automatically places text in lines of given width, it does so in a way that doesn't allow the content of each line to be read as a String afterwards, which makes it unusable for our purposes.
-Therefore each word must be placed manually, and to since it isn't the main focus of the program, each series of characters between a space is considered one word that has to be fully placed within one line.
-
-%Using the dimensions of each word, we create the graph by creating and connecting nodes in the space between lines at the start and end of each word, and at the end of the text area.
-We also create special nodes for each annotated word, they are located above the word in a central position.
-Each node also is located at set co-ordinates, which allows the routing algorithms to easily fulfill the monotonicity constraint.
-%We left some space in between the text area and the label area %Name=?(Lass dir was einfallen)
-%to allow the labels to be placed more freely.
-While the routing graph only reaches up to the border of the text area, the connection between the graph and the label is done with an OPO-Leader.
-As there is the possibility of intersecting with other leaders, this means some extra adjustments have to be done.
-For demonstration purposes, we also included another algorithm that uses S-Leaders for this section.
+% For demonstration purposes, we also included another algorithm that uses S-Leaders for this section.
 

### 2017-09-27 LaTeX/Paper.tex
Paper: Changed Formatting to standard for Bachelor theses at the TU Vienna, started to re-write the paper to use more formal language. Program: No changes.

@@
-\documentclass[11pt,a4paper]{article}
-\usepackage{termpaper}
+\documentclass[11pt,a4paper]{vutinfth}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{caption}
+\usepackage{mathtools}
+\usepackage{amssymb}
 %\usepackage{wrapfig}
 
-%opening
-\title{Boundary Labeling for annotated documents}%Vielleicht genauerer Titel?
-\author{
- \authorname{Jakob Klinger} \\
- \studentnumber{1125755} \\
- \curriculum{033 534} \\
- \email{e1125755@student.tuwien.ac.at}
-}
+%TODO: CHANGE!
+\setauthor{}{Jakob Klinger}{}{male}
+\setadvisor{Pretitle}{Forename Surname}{Posttitle}{male}
+
+\setaddress{TODO: ADDRESS!}%TODO
+\setregnumber{1125755}
+\setdate{25}{09}{2017}
+
+\settitle{TestTitle}{TestTitel}
+\setsubtitle{TestSubtitle}{TestSubtitel}
+
+\setthesis{bachelor}
+\setcurriculum{EnglischerCurrName}{DeutscherCurrName} %TODO!
 
@@
 
-\maketitle
+\frontmatter
+%\addtitlepage{naustrian}
+\addtitlepage{english}
+\addstatementpage
+
+%\begin{danksagung*}
+%\todo{Ihr Text hier.}
+%\end{danksagung*}
+
+%\begin{acknowledgements*}
+%\todo{Enter your text here.}
+%\end{acknowledgements*}
+
+%\begin{kurzfassung}
+%\todo{Ihr Text hier.}
+%\end{kurzfassung}
+
+%\begin{abstract}
+%\todo{Enter your text here.}
+%\end{abstract}
+
+\selectlanguage{english}
+
+\tableofcontents
+
+\mainmatter
 
@@
 Annotating a document is usually solved by adding footnotes or figures in an appropriate position and adding a simple reference in the text, leaving the reader to find the referenced content by themselves.
-Sometimes however, if a more obvious connection between the text and the referenced content is required, the reference is visibly connected to the text by drawing a straight line or a more complex path between them.
+If a more obvious connection between the text and the referenced content is required, the reference is visibly connected to the text by drawing a straight line or a more complex path between them.
 
@@
 
+A \emph{Graph} $G=\langle V, E \rangle$ is a tuple of \emph{Vertices} (V) and \emph{Edges} (E).Vertices are... % Each edge $e_i~\in~E \, (i~\in~\mathbb{N})$ is
+
 \begin{itemize}
- \item \textbf{Graph:} A Set of Vertices and Edges (see below), often used to represent routing problems. Can be further classified depending on its properties.
+ \item \textbf{Graph:} A Tuple of Vertices and Edges (see below), often used to represent routing problems. Can be further classified depending on its properties.
  \item \textbf{Edge:} A connection between two vertices, usually represented as a line. Can be directional, limiting its usage, or have a cost (``Weight'') associated with its use.
@@
 
+\backmatter %TODO: Es gibt mehr nützliche Kommandos in VUTINFTH für zb Listen von Bildern, etc.!
+
 \bibliographystyle{plain}

### 2017-09-28 LaTeX/Paper.tex
Paper: Finished Re-work of the Graph-related terminology. Probram: No changes.

@@
 
-\settitle{TestTitle}{TestTitel}
+\settitle{Boundary Labeling for annotated documents}{Boundary Labeling in annotierten Dokumenten}
 \setsubtitle{TestSubtitle}{TestSubtitel}
@@
 %Bild einfügen?
-However, most solutions proposed in that paper don't care whether a path overlapped with text or not, which results in a decrease in readability.
+However, most solutions proposed in that paper do not consider whether a path overlapped with text or not, which results in a decrease in readability.
 
@@
 
-A \emph{Graph} $G=\langle V, E \rangle$ is a tuple of \emph{Vertices} (V) and \emph{Edges} (E).Vertices are... % Each edge $e_i~\in~E \, (i~\in~\mathbb{N})$ is
+A \emph{Graph} $G=\langle V, E \rangle$ is a tuple of \emph{Vertices} $(V=\{v_1, v_2, ..., v_n\})$ and \emph{Edges} $(E=\{e_1, e_2, ..., e_m\})$.
+A Vertex $v_i (i~\in~\mathbb{N})$ is a featureless object.
+%Wie  sage ich, dass die Umsetzung der Vertices sehr flexibel ist?
+ Each edge $e_i \, (i~\in~\mathbb{N})$ is a relation between two vertices (Therefore: $E:V\rightarrow  V\times V$).
+The vertices forming the edge are considered \em{adjacent} to each other.
+Edges can also be directional, or have a weight, which affects how they are treated by algorithms.
+ A \em{Path} $(P=v_1, ..., v_h)$ is an ordered series of vertices, where each vertex must have an edge connecting it to the subesquent one.
 
 \begin{itemize}
- \item \textbf{Graph:} A Tuple of Vertices and Edges (see below), often used to represent routing problems. Can be further classified depending on its properties.
- \item \textbf{Edge:} A connection between two vertices, usually represented as a line. Can be directional, limiting its usage, or have a cost (``Weight'') associated with its use.
- \item \textbf{Vertex:} A featureless object that can be connected to any number other vertices via edges.
-Usually depicted as a point or small shape.
-Also referred to as \textbf{``Node''}.%Remove second sentence and re-define node?
- \item \textbf{Path:} An ordered series of either edges that share one vertex each with the edges listed before and after them, or an ordered series of vertices that have an edge connecting each vertex to the subsequent one.
+% \item \textbf{Graph:} A Tuple of Vertices and Edges (see below), often used to represent routing problems. Can be further classified depending on its properties.
+% \item \textbf{Edge:} A connection between two vertices, usually represented as a line. Can be directional, limiting its usage, or have a cost (``Weight'') associated with its use.
+% \item \textbf{Vertex:} A featureless object that can be connected to any number other vertices via edges. Usually depicted as a point or small shape.
+% \item \textbf{Path:} An ordered series of either edges that share one vertex each with the edges listed before and after them, or an ordered series of vertices that have an edge connecting each vertex to the subsequent one.
 Also known as \textbf{``Graph Walk''}.

### 2017-09-29 LaTeX/Paper.tex
Paper: Continued re-write of terminology section. Program: No changes.

@@
 The guidelines on how to create an optimal labeling are as follows: the connections should be as direct as possible, no important information should be obscured, and it should be easily discernable which Label belongs to which site.
-These three easily come into conflict with one another, especially when labeling documents, as the text usually is very dense and leaves little space for lines in between, yet one shouldn't allow them to pass through the text, as this makes the text harder to read.
+These three easily come into conflict with one another,  as the text usually is very dense and leaves little space for lines in between, yet one shouldn't allow them to pass through the text, as this makes the text harder to read.
 
@@
 
-For example, the paper about the Luatodonotes-Package\cite{Kindermann2014} illustrates some of the different styles of drawing these connecting paths, and came to the conclusion that paths without bends are easier to follow.
+In the paper about the Luatodonotes-Package\cite{Kindermann2014} illustrates some of the different styles of drawing these connecting paths, and came to the conclusion that paths without bends are easier to follow.
 %Bild einfügen?
-However, most solutions proposed in that paper do not consider whether a path overlapped with text or not, which results in a decrease in readability.
+However, most solutions proposed in that paper do not consider whether a path overlaps with text or not, which results in a decrease in readability.
 
@@
 A \emph{Graph} $G=\langle V, E \rangle$ is a tuple of \emph{Vertices} $(V=\{v_1, v_2, ..., v_n\})$ and \emph{Edges} $(E=\{e_1, e_2, ..., e_m\})$.
-A Vertex $v_i (i~\in~\mathbb{N})$ is a featureless object.
+A vertex $v_i (i~\in~\mathbb{N})$ is a featureless object.
 %Wie  sage ich, dass die Umsetzung der Vertices sehr flexibel ist?
- Each edge $e_i \, (i~\in~\mathbb{N})$ is a relation between two vertices (Therefore: $E:V\rightarrow  V\times V$).
-The vertices forming the edge are considered \em{adjacent} to each other.
+ Each edge $e_i \, (i~\in~\mathbb{N})$ is a relation between two vertices $(E:V\rightarrow  V\times V)$.
+The vertices forming the edge are considered \emph{adjacent} to each other.
 Edges can also be directional, or have a weight, which affects how they are treated by algorithms.
- A \em{Path} $(P=v_1, ..., v_h)$ is an ordered series of vertices, where each vertex must have an edge connecting it to the subesquent one.
+ A \emph{Path} $(P=v_1, ..., v_h)$ is an ordered series of vertices, where each vertex must have an edge connecting it to the subesquent one.
+
+\emph{Polylines} are a connected series of \emph{Line segments}. Line segments are straight lines that contain each point between their starting and end point. %Open vs. closed line segments?
+\emph{Labels} hold additional information and are represented as boxes containing this information.
+They are usually placed in the \emph{Label area} which is an area designated to hold labels and is located off to a side so the labels don't obscure anything.
+The point or object that this information refers to is called a \emph{Site}.
+The site and the label are connected via a \emph{Leader}, a polyline that can be further classified by looking at the orientation of its segments: \emph{O-Segments} run orthogonally to the border of the label area.
+\emph{P-Segments} run parallel to the border of the label area, and as such must be combined with other segments for the leader to reach its destination.
+\emph{S-Segments} are not required to have any particular orientation, and simply connect their start and einding points in a straight line.
+The leader's name is created by combining the name of the segments - for example, the leader from Fig~\ref{fig:term} would be classified as an OPO-Leader.
+The location where a leader connects to the label is called the \emph{Port}. It can either be located freely along the label's edge, or restricted to pre-defined positions.
 
@@
 
- \item \textbf{Polyline:} A connected series of straight lines.%Ist mehr Information notwendig?
+ %\item \textbf{Polyline:} A connected series of straight lines.%Ist mehr Information notwendig?
 
- \item \textbf{Label:} Additional information, usually represented as a box containing the additional information. Will also be referred to as \textbf{``Annotation''} in this paper.
- \item \textbf{Site:} The point, object or word that is labeled - without it, there wouldn't be any labeling necessary.
- \item \textbf{Leader:} The line or polyline connecting the site to the label - depending on its shape, it can be described further using several subgroups which can be freely combined to form new leader types:
-  \begin{itemize}
-   \item \textit{S-Leader:} This Leader connects site and label in a straight line.
-   \item \textit{O-Leader:} This leader runs orthogonally to the border between text and label area.
-   \item \textit{P-Leader:} This Leader runs parallel to the border between text and label area. It must be combined with other leader styles, as it won't reach the label area otherwise.
-  \end{itemize}
-    For example, the leader from Fig.~\ref{fig:term} would be classified as an opo-Leader.
- \item \textbf{Port:} The location where the leader connects to the label. It may be restricted to pre-determined locations, like only at the corners.
+% \item \textbf{Label:} Additional information, usually represented as a box containing the additional information. Will also be referred to as \textbf{``Annotation''} in this paper.
+% \item \textbf{Site:} The point, object or word that is labeled - without it, there wouldn't be any labeling necessary.
+% \item \textbf{Leader:} The line or polyline connecting the site to the label - depending on its shape, it can be described further using several subgroups which can be freely combined to form new leader types:
+%  \begin{itemize}
+ %  \item \textit{S-Leader:} This Leader connects site and label in a straight line.
+  % \item \textit{O-Leader:} This leader runs orthogonally to the border between text and label area.
+  % \item \textit{P-Leader:} This Leader runs parallel to the border between text and label area. It must be combined with other leader styles, as it won't reach the label area otherwise.
+  %\end{itemize}
+   % For example, the leader from Fig.~\ref{fig:term} would be classified as an opo-Leader.
+% \item \textbf{Port:} The location where the leader connects to the label. It may be restricted to pre-determined locations, like only at the corners.
   \item \textbf{Monotonous/Monotonicity:} Describes a steady progression towards a direction or goal.

### 2017-10-01 LaTeX/Paper.tex
Paper: Finished first re-write. Program: No changes.

@@
 \setauthor{}{Jakob Klinger}{}{male}
-\setadvisor{Pretitle}{Forename Surname}{Posttitle}{male}
+\setadvisor{}{}{}{}%{Pretitle}{Forename Surname}{Posttitle}{male}
+%Ass.Prof. Dipl.-Inform. Dr.rer.nat. Martin Nöllenburg
+% University Assistant, M.Sc., B.Sc. Fabian Klute
 
@@
 \settitle{Boundary Labeling for annotated documents}{Boundary Labeling in annotierten Dokumenten}
-\setsubtitle{TestSubtitle}{TestSubtitel}
+\setsubtitle{}{}%{TestSubtitleENG}{TestSubtitelGER}
 
 \setthesis{bachelor}
-\setcurriculum{EnglischerCurrName}{DeutscherCurrName} %TODO!
+\setcurriculum{Bachelor's programme Software \& Information Engineering}{Bachelorstudium Software \& Information Engineering} %TODO!
 
@@
 
-\section{Introduction}%Wort ``Path'' durch nicht-definierte Terminologie ersetzen?
-Annotating a document is usually solved by adding footnotes or figures in an appropriate position and adding a simple reference in the text, leaving the reader to find the referenced content by themselves.
-If a more obvious connection between the text and the referenced content is required, the reference is visibly connected to the text by drawing a straight line or a more complex path between them.
+\chapter{Introduction}%Wort ``Path'' durch nicht-definierte Terminologie ersetzen?
+Annotating a document is usually solved by adding footnotes in an appropriate position and adding a simple reference in the text, leaving the reader to find the referenced content by themselves.
+If a more obvious connection between the text and the referenced content is required, the reference placed to the side of the document and visibly connected to the text by drawing a straight line or a more complex path between them.
 
-In this paper, we will look at ways to use Boundary Labeling, which means that all annotations will be placed outside of the text they are referencing and will be visually connected to the feature they are referencing.
+In this paper, we will look at ways to use Boundary Labeling, which means that all annotations will be placed somewhere outside of the text they are referencing and will be visually connected to the feature they are referencing.
 (See also \cite{Bekos2007})
 
-The guidelines on how to create an optimal labeling are as follows: the connections should be as direct as possible, no important information should be obscured, and it should be easily discernable which Label belongs to which site.
-These three easily come into conflict with one another,  as the text usually is very dense and leaves little space for lines in between, yet one shouldn't allow them to pass through the text, as this makes the text harder to read.
+The guidelines on how to create suitable labelings are as follows: the connections should be as direct as possible, no important information should be obscured, and it should be easily discernable which Label belongs where.
+These three criteria easily come into conflict with one another, as the text usually is very dense and leaves little space for lines in between, yet one shouldn't allow them to pass through the text, as this makes the text harder to read.
 
@@
  A \emph{Path} $(P=v_1, ..., v_h)$ is an ordered series of vertices, where each vertex must have an edge connecting it to the subesquent one.
+ \emph{Depth-first search} is a searching algorithm on a graph G that starts at a given vertex $v_s \in V$ and explores the graph by traversing its edges as far as possible before backtracking.
+In our algorithm, the goal is to reach a target vertex $v_t \in V$, at which point the algorithm terminates and returns the route taken from $v_s$ to $v_t$.
+% A search algorithm is \emph{monotonous} if, for each vertex $v$ and each successor $v'$ generated by the algorithm, the estimated cost of reaching the goal from $v$ is not greater than the cost of reaching $v'$ plus the estimated cost of reaching the goal from $v'$.
 
@@
 The leader's name is created by combining the name of the segments - for example, the leader from Fig~\ref{fig:term} would be classified as an OPO-Leader.
-The location where a leader connects to the label is called the \emph{Port}. It can either be located freely along the label's edge, or restricted to pre-defined positions.
+The location where a leader connects to the label is called the \emph{Port}. It can be restricted to pre-defined positions. %TODO: Re-insert monotonicity
 
-\begin{itemize}
-% \item \textbf{Graph:} A Tuple of Vertices and Edges (see below), often used to represent routing problems. Can be further classified depending on its properties.
-% \item \textbf{Edge:} A connection between two vertices, usually represented as a line. Can be directional, limiting its usage, or have a cost (``Weight'') associated with its use.
-% \item \textbf{Vertex:} A featureless object that can be connected to any number other vertices via edges. Usually depicted as a point or small shape.
-% \item \textbf{Path:} An ordered series of either edges that share one vertex each with the edges listed before and after them, or an ordered series of vertices that have an edge connecting each vertex to the subsequent one.
-Also known as \textbf{``Graph Walk''}.
+% \textbf{Monotonous/Monotonicity:} Describes a steady progression towards a direction or goal.
+Will be used in this paper to indicate that there's no loss of progress made from the site to the label, effectively restricting the direction a leader can go at each bend.
+%+Node für Knoten in Polylinien?
 
- %\item \textbf{Polyline:} A connected series of straight lines.%Ist mehr Information notwendig?
+\chapter{The Algorithm}
 
-% \item \textbf{Label:} Additional information, usually represented as a box containing the additional information. Will also be referred to as \textbf{``Annotation''} in this paper.
-% \item \textbf{Site:} The point, object or word that is labeled - without it, there wouldn't be any labeling necessary.
-% \item \textbf{Leader:} The line or polyline connecting the site to the label - depending on its shape, it can be described further using several subgroups which can be freely combined to form new leader types:
-%  \begin{itemize}
- %  \item \textit{S-Leader:} This Leader connects site and label in a straight line.
-  % \item \textit{O-Leader:} This leader runs orthogonally to the border between text and label area.
-  % \item \textit{P-Leader:} This Leader runs parallel to the border between text and label area. It must be combined with other leader styles, as it won't reach the label area otherwise.
-  %\end{itemize}
-   % For example, the leader from Fig.~\ref{fig:term} would be classified as an opo-Leader.
-% \item \textbf{Port:} The location where the leader connects to the label. It may be restricted to pre-determined locations, like only at the corners.
-  \item \textbf{Monotonous/Monotonicity:} Describes a steady progression towards a direction or goal.
-Will be used in this paper to indicate that there's no loss of progress made from the site to the label, effectively restricting the direction a leader can go at each bend.
-  \item \textbf{Depth-first search:} A searching algorithm that continues along a set of rules as far as possible until either the goal is found, or backtracking is required.
-A common example is navigating a maze by turning the same way whenever possible.
-\end{itemize}
-%Mehr Graphenterminologie? (zb: Path) +Polylinien Vertex/Node für Knoten in Polylinien bzw. Graphen -->Eindeutig!
+In our algorithm, the focus was put on keeping the text as readable as possible.
+This means that leaders aren't allowed to pass through words, which was simplified by enclosing each word in a bounding rectangle that leaders aren't allowed to pass through.
+The leaders should also be kept as short as possible, so we restricted them to always be moving toward the label in at least one dimension.%Wie drücke ich das als Formel aus?
+We also wanted to use the space available in the labeling area as efficiently as possible, so labels are placed as far up as possible in order to maximize the space available to future labels at the cost of possibly increased leader length.
 
-\section{The program}
+%\subsection{Algorithm}%''Underlying Algorithm''?
+To easily create leaders that exclusively use the space not taken up by a word's bounding rectangle, we decided to use a graph similar to the one Loose \cite{Loose2015} used.
+As each vertex represents a physical location, they will have co-ordinates associated with them, and the vertices representing the sites will hold additional information regarding its leader and label.
+The graph is constructed by placing vertices in the space between lines next to each corner of a word's bounding rectangle, with two consecutive words in a line sharing the two vertices associated with their adjacent corners.
+%Insert picture
+For the sites, we inserted an additional vertex above the center of the word, which will serve as the leader's starting point.
+After placing all vertices, we created edges between each vertex and the closest horizontal neighbour to both sides, and between nodes that are located exactly above or below each other, and exactly one line apart.
 
-In our program, the focus was put on keeping the text as readable as possible.
-This means that leaders are only allowed in the space between words or lines, which can make certain locations for sites unavailable, as there might already too many leaders routed through that area to insert yet another for the new site.
-We also wanted the leaders to be monotonous, which means that they will be one of the shortest possible connection between port and site locations.
-Our final priority was to efficiently use the space for the labels, which meant that each label should initially be placed as close to the available area's border as possible.
-While this allows for more space in which labels can be placed, it also means that the distance between Label and site often increases, which leads to longer leaders.%''Final Priority'' klingt komisch -> Ersatz finden!
+In our labeling algorithm, we decided to work through the labels in the order they appear in the text, and placing each as far up as possible, skipping any label that would've required us to move a lower line.
+We also opted to use fixed ports located in the top left corner of each label, as this allowed us to unabiguously place annotations by only knowing their port location, which is equal to their leader's ending point.
+To not restrict the label's placement by the line spacing, we left a buffer zone between the text area and the label area, which allows to place labels even further above to optimally use the available space.
+The routing was implemented using a depth-first search algorithm, restricted to only selecting vertices located above or to the right of the current vertex, prioritizimg locations above whenever possible.
+After the algorithm terminates, it returns a path from the site to the right text border and information on where to draw the OPO-segment that connects the text area with the label's port, if successful.
 
-%Port-Location erwähnen!
-\subsection{Algorithm}%''Underlying Algorithm''?
-To easily create leaders that exclusively use the space between words and lines, we decided to use a graph similar to the one Loose \cite{Loose2015} used, placing vertices above and below each word's beginning and end, and connecting them with their horizontally and vertically adjacent neighbours.%Umformulieren?
-For the sites, we inserted an additional vertex above the center of the word, which will serve as the leader's starting point.
-We also decided to give the graph's edges a capacity of 1, which means that each leader will always use edges which haven't already be taken by other leaders.
-While this causes some sites to become un-routable due to another leader passing right through their starting vertex, it makes visualizing the leaders much easier.
-
-Our first approach was going through the labels in the order they appear in the text and placing each as far up as possible, to maximize the remaining space other labels could use.
-We also opted to use fixed ports located in the top left corner of the annotation, as this allowed us to unambiguously place any annotation by knowing either their port's location, wich is also the leader's ending point.
-However, since putting the labels right next to the text would result in the labels' positions being restricted by the text's line spacing, we then decided to leave a buffer zone in between the two areas, which allows the leaders to adjust between the position forced by the line spacing and the label's final position.
-As this allowed us to place the labels more freely, we adjusted the label's positions after all labels were placed to reduce the distance that needed to be covered in the buffer area.
-
-\subsection{Implementation}
-The program was written in Java, using only JGraphT\cite{JGraphT} as additional library.
-Since we only want to create leaders that don't intersect with the text, the graph was created alongside the placement of the words on the canvas, as it was easy to extract measurements at this point.
+\section{Implementation}
+The program was written in Java 1.8.0u40, using JGraphT1.0.1\cite{JGraphT} as graph library.
+Since we only want to create leaders that don't intersect with the text, the graph was created alongside the placement of the words on the canvas.%, as it was easy to extract measurements at this point.
 Due to some problems with Java's various methods of calculating linebreaks, this process was done completely by hand.
 
-As the Graph's vertices represent fixed locations on the canvas, they each have coordinates associated with them, with the vertices representing the sites containing extra information, such as references to the corresponding annotation and to the leader connecting the two (if existent).
-This additional information is used for both routing and drawing the leaders: The routing was done via a depth-first search algorithm that prioritized vertices located further above if possible, and vertices to the right otherwise.
+%As the Graph's vertices represent fixed locations on the canvas, they each have coordinates associated with them, with the vertices representing the sites containing extra information, such as references to the corresponding annotation and to the leader connecting the two (if existent).
+%The additional information stored in the vertices is used for both routing and drawing the leaders: The routing was done via a depth-first search algorithm that prioritized vertices located further up if possible, and vertices to the right otherwise.
 After the algorithm terminates, it returns  a vertex-based graph walk describing the route from the site to the border of the text area alongside information about where to draw the OPO-segment in the buffer zone if successful, or a walk containing only the starting vertex, if no valid path was found.
-Using this information alongside the positional data stored in the vertices, the polyline representing the leader can be easily drawn.
+Using this information alongside the positional data stored in the vertices, the polyline representing the leader can be drawn.
 %TODO: Schönheitsfeatures, etc.

### 2017-10-28 LaTeX/Paper.tex
Paper: Minor corrections. Program: No changes.

@@
 
-A \emph{Graph} $G=\langle V, E \rangle$ is a tuple of \emph{Vertices} $(V=\{v_1, v_2, ..., v_n\})$ and \emph{Edges} $(E=\{e_1, e_2, ..., e_m\})$.
-A vertex $v_i (i~\in~\mathbb{N})$ is a featureless object.
+A \emph{graph} $G=\langle V, E \rangle$ is a tuple of \emph{vertices} $V=\{v_1, v_2, ..., v_n\}$ and \emph{edges} $E=\{e_1, e_2, ..., e_m\}$.
+A vertex $v$ is a featureless object.
 %Wie  sage ich, dass die Umsetzung der Vertices sehr flexibel ist?
- Each edge $e_i \, (i~\in~\mathbb{N})$ is a relation between two vertices $(E:V\rightarrow  V\times V)$.
-The vertices forming the edge are considered \emph{adjacent} to each other.
-Edges can also be directional, or have a weight, which affects how they are treated by algorithms.
- A \emph{Path} $(P=v_1, ..., v_h)$ is an ordered series of vertices, where each vertex must have an edge connecting it to the subesquent one.
- \emph{Depth-first search} is a searching algorithm on a graph G that starts at a given vertex $v_s \in V$ and explores the graph by traversing its edges as far as possible before backtracking.
+ Each edge $e$ is a relation between two vertices $E:V\rightarrow  V\times V$. %Edges can also be directional, or have a weight, which affects how they are treated by algorithms.
+We call two vertices $u,v \in V$ \emph{adjacent}, if the edge $e=(u,v) \in E$.
+ A \emph{path} $P=v_1, ..., v_h$ is an ordered series of vertices, where each vertex must have an edge connecting it to the subesquent one.
+ \emph{Depth-first search} is a searching algorithm on a graph G that starts at a given vertex $v \in V$ and explores the graph by traversing its edges as far as possible before backtracking.
 In our algorithm, the goal is to reach a target vertex $v_t \in V$, at which point the algorithm terminates and returns the route taken from $v_s$ to $v_t$.
@@
 
-\emph{Polylines} are a connected series of \emph{Line segments}. Line segments are straight lines that contain each point between their starting and end point. %Open vs. closed line segments?
-\emph{Labels} hold additional information and are represented as boxes containing this information.
+\emph{polylines} are a connected series of \emph{Line segments}. Line segments are straight lines that contain each point between their starting and end point. %Open vs. closed line segments?
+\emph{labels} hold additional information and are represented as boxes containing this information.
 They are usually placed in the \emph{Label area} which is an area designated to hold labels and is located off to a side so the labels don't obscure anything.
-The point or object that this information refers to is called a \emph{Site}.
-The site and the label are connected via a \emph{Leader}, a polyline that can be further classified by looking at the orientation of its segments: \emph{O-Segments} run orthogonally to the border of the label area.
+The point or object that this information refers to is called a \emph{site}.
+The site and the label are connected via a \emph{leader}, a polyline that can be further classified by looking at the orientation of its segments: \emph{O-Segments} run orthogonally to the border of the label area.
 \emph{P-Segments} run parallel to the border of the label area, and as such must be combined with other segments for the leader to reach its destination.
@@
 The leader's name is created by combining the name of the segments - for example, the leader from Fig~\ref{fig:term} would be classified as an OPO-Leader.
-The location where a leader connects to the label is called the \emph{Port}. It can be restricted to pre-defined positions. %TODO: Re-insert monotonicity
+The location where a leader connects to the label is called the \emph{port}. It can be restricted to pre-defined positions. %TODO: Re-insert monotonicity
 

### 2017-11-01 LaTeX/Paper.tex
Paper: Polished Section 2, added image showing the routing graph. Program: Added Debug variable 'hideLeaders'.

@@
 In our algorithm, the focus was put on keeping the text as readable as possible.
-This means that leaders aren't allowed to pass through words, which was simplified by enclosing each word in a bounding rectangle that leaders aren't allowed to pass through.
-The leaders should also be kept as short as possible, so we restricted them to always be moving toward the label in at least one dimension.%Wie drücke ich das als Formel aus?
-We also wanted to use the space available in the labeling area as efficiently as possible, so labels are placed as far up as possible in order to maximize the space available to future labels at the cost of possibly increased leader length.
+This means that leaders aren't allowed to pass through words, which was implemented by enclosing each word in a bounding rectangle that leaders aren't allowed to pass through.
+The leaders should also be kept as short as possible, so we restricted them to always be moving toward the label, even if the direct route is unavailable.%Wie drücke ich das als Formel aus?
+We also wanted to use the space available in the labeling area as efficiently as possible, so labels are placed as far up as possible in order to maximize the space available to future labels at the cost of increased leader length.
 
-%\subsection{Algorithm}%''Underlying Algorithm''?
-To easily create leaders that exclusively use the space not taken up by a word's bounding rectangle, we decided to use a graph similar to the one Loose \cite{Loose2015} used.
+\begin{figure}%{r}{0.35\textwidth}
+ \captionsetup{justification=centering, margin=0.75cm}
+ \centering
+  \includegraphics[scale=1.0]{RoutingGraphPartial_NoBorders.png}
+  \caption{Visual representation of the routing graph. The highlighted nodes represent leader sources.}
+ \label{fig:RGraph}
+\end{figure}
+To create leaders that exclusively use the space not taken up by a word's bounding rectangle, we decided to use a graph similar to the one Loose \cite{Loose2015} used.
 As each vertex represents a physical location, they will have co-ordinates associated with them, and the vertices representing the sites will hold additional information regarding its leader and label.
-The graph is constructed by placing vertices in the space between lines next to each corner of a word's bounding rectangle, with two consecutive words in a line sharing the two vertices associated with their adjacent corners.
+The graph is constructed by placing vertices between the lines, located next to each corner of a word's bounding rectangle, with two consecutive words in a line sharing the two vertices associated with their adjacent corners.
 %Insert picture
@@
 After placing all vertices, we created edges between each vertex and the closest horizontal neighbour to both sides, and between nodes that are located exactly above or below each other, and exactly one line apart.
+(For a visual representation, see Fig.~\ref{fig:RGraph})
 
-In our labeling algorithm, we decided to work through the labels in the order they appear in the text, and placing each as far up as possible, skipping any label that would've required us to move a lower line.
+In our labeling algorithm, we decided to work through the labels in the order they appear in the text, placing each as far up as possible, and skipping any label that would've required us place the label below its leader's source node.
 We also opted to use fixed ports located in the top left corner of each label, as this allowed us to unabiguously place annotations by only knowing their port location, which is equal to their leader's ending point.
-To not restrict the label's placement by the line spacing, we left a buffer zone between the text area and the label area, which allows to place labels even further above to optimally use the available space.
+To not restrict the label's placement by the line spacing, we left a buffer zone between the text area and the label area, which allows us to place labels even further above to optimally use the available space.
 The routing was implemented using a depth-first search algorithm, restricted to only selecting vertices located above or to the right of the current vertex, prioritizimg locations above whenever possible.

### 2017-11-05 LaTeX/Paper.tex
Paper: Modified Section 2. Program: Changes from previous commit (new Debug variable) failed to commit. Adding them now.

@@
 To not restrict the label's placement by the line spacing, we left a buffer zone between the text area and the label area, which allows us to place labels even further above to optimally use the available space.
-The routing was implemented using a depth-first search algorithm, restricted to only selecting vertices located above or to the right of the current vertex, prioritizimg locations above whenever possible.
-After the algorithm terminates, it returns a path from the site to the right text border and information on where to draw the OPO-segment that connects the text area with the label's port, if successful.
+The algorithm's input consists of the Graph $G$ and a set of sites $V_{ann} \subset V$ which will be routed using a depth-first search algorithm, restricted to only selecting vertices located above or to the right of the current vertex, prioritizing moving up whenever possible.
+After the algorithm terminates, it returns a set with routing information for each vertex, which contains a Path $P=\{v_1, \cdots, v_n\}$, leading from the site to the text area's border, along with some additional information on how to draw the OPO-Segment that connects to the Label's port.
+If the routing for any given site failed, the path consists of a single vertex - the site.
 

### 2017-11-05 Eclipse/Labeling/src/routingapp/GraphCreatorPanel.java

@@ -44,6 +44,7 @@ public class GraphCreatorPanel extends JComponent {
 	//DEBUG VALUES
 	private boolean showWordBoundaries=false;//draws rectangles around detected word boundaries in main text, if set to true
 	private boolean showGraphGrid=false;//Draws the whole routing Graph 
+	private boolean hideLeaders=false; //Hides the leaders and unsuccessfully routed nodes, if some other feature needs to be inspected visually 
 	//DEBUG VALUES END
 
 	private int width=600;
@@ -361,17 +362,20 @@ public class GraphCreatorPanel extends JComponent {
 		//Calculate all Routes via Router.findRoutes();
 
 		List<RouteInfo> results=router.findRoutes(graph, annotatedTuples);
-		Iterator<RouteInfo> it=results.iterator();
-		while(it.hasNext())
+		if(!hideLeaders)
 		{
-			drawRouteInfo(g,graph,it.next());
+			Iterator<RouteInfo> it=results.iterator();
+			while(it.hasNext())
+			{
+				drawRouteInfo(g,graph,it.next());
+			}
 		}
 
 	}
 
 	/**
-	 * Helper/Debug method - draws a simple visualization of the supplied graph
-	 * NOTE: The current implementation will draw all nodes multiple times, once for each of their edges.
+	 * Helper/Debug method - draws a simple visualization of the generated graph
+	 * NOTE: This implementation will draw all nodes multiple times, once for each of their edges.
 	 * @param graph the graph
 	 * @param g the Graphics object of the visualization's basis
 	 * @param col the color that will be used to draw the graph - the previously used color will be restored afterwards. 

### 2017-11-15 LaTeX/Paper.tex
Added Pseudocode.tex, esperimented a bit with it.
Paper: Added Supervisor names, re-wrote imprecise parts of terminology section, started work on picture to aid terminology explanation. Program: No changes.

@@
 \usepackage{amssymb}
+\usepackage{algorithm}
+\usepackage{algpseudocode}
 %\usepackage{wrapfig}
@@
 \setauthor{}{Jakob Klinger}{}{male}
-\setadvisor{}{}{}{}%{Pretitle}{Forename Surname}{Posttitle}{male}
+\setadvisor{Ass.Prof. Dipl.-Inform. Dr.rer.nat}{Martin Nöllenburg}{}{male}%{Pretitle}{Forename Surname}{Posttitle}{male}
+\setfirstassistant{Univ.Ass.}{Fabian Klute}{ M.Sc., B.Sc.}{male}
 %Ass.Prof. Dipl.-Inform. Dr.rer.nat. Martin Nöllenburg
@@
 \section{Terminology and Fundamentals} %Evtl in Introduction eingliedern?
-While Boundary Labeling (or an equivalent concept) can be applied to a space with different geometry or more dimensions, this paper will only concern itself with two-dimensonal, euclidean space.
+While Boundary Labeling (or an equivalent concept) can be applied to a space with different geometry or more dimensions, this paper will only concern itself with two-dimensonal, Euclidean space.
 To easily reference important concepts, some additional terminology will be introduced as well. (See Fig.~\ref{fig:term} for a visual explanation)%Add references to all terminology pictures!
@@
 %Wie  sage ich, dass die Umsetzung der Vertices sehr flexibel ist?
- Each edge $e$ is a relation between two vertices $E:V\rightarrow  V\times V$. %Edges can also be directional, or have a weight, which affects how they are treated by algorithms.
+ Each edge $e$ is a relation between two vertices $E \subseteq V\times V$. %Edges can also be directional, or have a weight, which affects how they are treated by algorithms.
 We call two vertices $u,v \in V$ \emph{adjacent}, if the edge $e=(u,v) \in E$.
- A \emph{path} $P=v_1, ..., v_h$ is an ordered series of vertices, where each vertex must have an edge connecting it to the subesquent one.
+ A \emph{path} $P=v_1, ..., v_h$ is an ordered sequence of vertices, where each vertex must have an edge connecting it to the subesquent one.
  \emph{Depth-first search} is a searching algorithm on a graph G that starts at a given vertex $v \in V$ and explores the graph by traversing its edges as far as possible before backtracking.
@@
 
-\emph{polylines} are a connected series of \emph{Line segments}. Line segments are straight lines that contain each point between their starting and end point. %Open vs. closed line segments?
-\emph{labels} hold additional information and are represented as boxes containing this information.
-They are usually placed in the \emph{Label area} which is an area designated to hold labels and is located off to a side so the labels don't obscure anything.
-The point or object that this information refers to is called a \emph{site}.
+\emph{Polylines} are a connected series of \emph{line segments}. Line segments are straight lines that contain each point between their starting and end point. 
+\emph{Labels} hold additional information and are represented as boxes containing this information.
+They are usually placed in the \emph{Label area} which is a rectangular area designated to hold labels.
+It is located next to of the bigger \emph{text area}, which contains the document's text and all \emph{sites}, the points or objects that a label's information refers to.
 The site and the label are connected via a \emph{leader}, a polyline that can be further classified by looking at the orientation of its segments: \emph{O-Segments} run orthogonally to the border of the label area.
@@
 \emph{S-Segments} are not required to have any particular orientation, and simply connect their start and einding points in a straight line.
-The leader's name is created by combining the name of the segments - for example, the leader from Fig~\ref{fig:term} would be classified as an OPO-Leader.
+The leader's name is created by combining the name of the segments - for example, the leader from Fig~\ref{fig:term} would be classified as an OPO-Leader.%Re-write once picture is updated
 The location where a leader connects to the label is called the \emph{port}. It can be restricted to pre-defined positions. %TODO: Re-insert monotonicity
@@
 
+%\begin{algorithm}
+%\caption{My algorithm}\label{euclid}
+%\begin{algorithmic}[1]
+%\Procedure{MyProcedure}{}
+%\State $\textit{stringlen} \gets \text{length of }\textit{string}$
+%\State $i \gets \textit{patlen}$
+%\BState \emph{top}:
+%\If {$i > \textit{stringlen}$} \Return false
+%\EndIf
+%\State $j \gets \textit{patlen}$
+%\BState \emph{loop}:
+%\If {$\textit{string}(i) = \textit{path}(j)$}
+%\State $j \gets j-1$.
+%\State $i \gets i-1$.
+%\State \textbf{goto} \emph{loop}.
+%\State \textbf{close};
+%\EndIf
+%\State $i \gets i+\max(\textit{delta}_1(\textit{string}(i)),\textit{delta}_2(j))$.
+%\State \textbf{goto} \emph{top}.
+%\EndProcedure
+%\end{algorithmic}
+%\end{algorithm}
+
 % For demonstration purposes, we also included another algorithm that uses S-Leaders for this section.

### 2017-11-16 LaTeX/Paper.tex
Paper: Completed new Graph Terminology drawing, added to paper. Program: No changes.

@@
  \centering
-  \includegraphics[scale=0.5]{IPE_TerminologyDrawing.pdf}
+  \includegraphics[scale=0.90]{GraphTerminologyExtended.png}
   \caption{Illustrated guide to the labeling terminology}
@@
  A \emph{path} $P=v_1, ..., v_h$ is an ordered sequence of vertices, where each vertex must have an edge connecting it to the subesquent one.
- \emph{Depth-first search} is a searching algorithm on a graph G that starts at a given vertex $v \in V$ and explores the graph by traversing its edges as far as possible before backtracking.
-In our algorithm, the goal is to reach a target vertex $v_t \in V$, at which point the algorithm terminates and returns the route taken from $v_s$ to $v_t$.
+ \emph{Depth-first search} is a searching algorithm on a graph G that starts at a given vertex $v \in V$ and explores the graph by traversing its edges as far as possible before backtracking, and continues to do so until a pre-defined goal is met.
+%In our algorithm, the goal is to reach a target vertex $v_t \in V$, at which point the algorithm terminates and returns the route taken from $v_s$ to $v_t$.
 % A search algorithm is \emph{monotonous} if, for each vertex $v$ and each successor $v'$ generated by the algorithm, the estimated cost of reaching the goal from $v$ is not greater than the cost of reaching $v'$ plus the estimated cost of reaching the goal from $v'$.
@@
 \emph{S-Segments} are not required to have any particular orientation, and simply connect their start and einding points in a straight line.
-The leader's name is created by combining the name of the segments - for example, the leader from Fig~\ref{fig:term} would be classified as an OPO-Leader.%Re-write once picture is updated
+The leader's name is created by combining the name of the segments - for example, the blue leader from Fig~\ref{fig:term} would be classified as an OPOPO-Leader.%Re-write once picture is updated
 The location where a leader connects to the label is called the \emph{port}. It can be restricted to pre-defined positions. %TODO: Re-insert monotonicity

### 2017-11-18 LaTeX/Paper.tex
Paper: Rewrote Pseudocode using the 'algorithm2e'-Package and inserted it into paper. Program: No changes.

@@
 \usepackage{amssymb}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
+\usepackage[boxed,linesnumbered]{algorithm2e}
 %\usepackage{wrapfig}
@@
 
-\begin{figure}%{r}{0.35\textwidth}
- \captionsetup{justification=centering, margin=0.75cm}
- \centering
-  \includegraphics[scale=1.0]{RoutingGraphPartial_NoBorders.png}
-  \caption{Visual representation of the routing graph. The highlighted nodes represent leader sources.}
- \label{fig:RGraph}
-\end{figure}
+\begin{algorithm}%{r}{0.35\textwidth}
+\DontPrintSemicolon
+\SetKwData{Words}{words}
+\SetKwData{PrevW}{previousWord}
+\SetKwData{G}{Graph}
+\SetKwData{W}{w} \SetKwData{V}{v}\SetKwData{Va}{v1} \SetKwData{Vb}{v2} \SetKwData{Vc}{v3} \SetKwData{Vd}{v4}
+\SetKwData{Up}{UpperVerticesList}
+\SetKwData{Low}{LowerVerticesList}
+
+\SetKwFunction{setAnn}{setAnnotation}
+\SetKwFunction{NewVert}{new Vertex}
+\SetKwFunction{AddVert}{addVertex}
+\SetKwFunction{AddAll}{addAll}
+\SetKwFunction{CEdge}{createEdgeBetween}
+\SetKwFunction{PosConn}{connectBasedOnPosition}
+\SetKwFunction{ClrList}{emptyList}
+\SetKwFunction{getC}{getCenter}
+\SetKwFunction{tLeft}{getTopLeft}
+\SetKwFunction{tRight}{getTopRight}
+\SetKwFunction{bLeft}{getBottomLeft}
+\SetKwFunction{bRight}{getBottomRight}
+\SetKwFunction{Newline}{startNewLine}
+
+\KwData{A text with annotations,stored as a String-Array}
+\KwResult{A Graph (as described above)}
+
+ initialization\;
+
+ \ForEach{\W in \Words}{
+  \eIf{\W is annotation}{
+   \V$\gets$\NewVert{\PrevW.\getC{}}\;
+   \V.\setAnn{w}\;
+   \G.\AddVert{\V}\;
+   \Up.\AddVert{\V}\;
+  }{
+   \If{\W is too big for the line}{
+    \Newline{}\;
+    \PosConn{\Up}\;
+    \Up$\gets$\Low\;
+    \ClrList{\Low}\;
+   }
+   \Va $\gets$\NewVert{\W.\tLeft{}}\;
+   \Vb $\gets$\NewVert{\W.\tRight{}}\;
+   \Vc $\gets$\NewVert{\W.\bLeft{}}\;
+   \Vd $\gets$\NewVert{\W.\bRight{}}\;
+   \;
+   \G.\AddAll{\Va,\Vb,\Vc,\Vd}\;
+   \Up.\AddAll{\Va,\Vb}\;
+   \Low.\AddAll{\Vc,\Vd}\;
+   \G.\CEdge{\Va,\Vc}\;
+   \G.\CEdge{\Vb,\Vd}\;
+  }
+ } 
+\caption{Representation of the Graph-creation algorithm in pseudocode}
+\label{alg:GraphCreation}
+\end{algorithm}
 To create leaders that exclusively use the space not taken up by a word's bounding rectangle, we decided to use a graph similar to the one Loose \cite{Loose2015} used.
@@
 After placing all vertices, we created edges between each vertex and the closest horizontal neighbour to both sides, and between nodes that are located exactly above or below each other, and exactly one line apart.
-(For a visual representation, see Fig.~\ref{fig:RGraph})
+(For a representation in Pseudocode, see Alg.~\ref{alg:GraphCreation})
 
@@
 If the routing for any given site failed, the path consists of a single vertex - the site.
+(For an illustration in Pseudocode see Alg.~\ref{alg:DFS}.)
+
+\begin{algorithm}
+
+\KwData{A single annotation's source and its Graph}
+\KwResult{A List of vertices describing the Leader's Path}
+
+\SetKwData{G}{Graph}
+\SetKwData{Path}{Path}
+\SetKwData{Src}{Source}
+\SetKwData{Curr}{currentVertex}
+\SetKwData{Old}{oldVertex}
+\SetKwData{Back}{backtracking}
+
+\SetKwFunction{Top}{getTopNeighbourOf}
+\SetKwFunction{Right}{getRightNeighbourOf}
+\SetKwFunction{AddVert}{addVertex}
+\SetKwFunction{RemVert}{RemoveVertex}
+\SetKwFunction{Last}{getLastEntry}
+\SetKwFunction{Below}{isBelow}
+\SetKwFunction{Break}{break}
+
+initialization\;
+\While{\Curr not at right text border}{
+
+ \uIf{$($\G.\Top{\Curr}$\neq null)\land \neg$\Back}{
+  \Path.\AddVert{\Curr}\;
+  \Curr $\gets$ \G.\Top{\Curr}\;\;
+ }
+ \uElseIf{\G.\Right{\Curr}$\neq null$}{
+ \Path.\AddVert{\Curr}\;
+  \Curr $\gets$ \G.\Top{\Curr}\;
+  \Back $\gets$ False\;
+ }
+ \Else{
+  \Back $\gets$ True\;
+  \Repeat{\Curr's Position is below \Old or\; ~~~\Path is Empty}{
+   \Old $\gets$ \Curr\;
+   \Curr $\gets$ \Path.\Last{}\;
+   \Path.\RemVert{\Curr}\;
+  }
+  \If(//No path found){\Curr not below \Old}{\Break}
+ }
+}
+\caption{The Depth-First-Search algorithm used in the program.}
+\label{alg:DFS}
+\end{algorithm}
 
@@
 
-%\begin{algorithm}
-%\caption{My algorithm}\label{euclid}
-%\begin{algorithmic}[1]
-%\Procedure{MyProcedure}{}
-%\State $\textit{stringlen} \gets \text{length of }\textit{string}$
-%\State $i \gets \textit{patlen}$
-%\BState \emph{top}:
-%\If {$i > \textit{stringlen}$} \Return false
-%\EndIf
-%\State $j \gets \textit{patlen}$
-%\BState \emph{loop}:
-%\If {$\textit{string}(i) = \textit{path}(j)$}
-%\State $j \gets j-1$.
-%\State $i \gets i-1$.
-%\State \textbf{goto} \emph{loop}.
-%\State \textbf{close};
-%\EndIf
-%\State $i \gets i+\max(\textit{delta}_1(\textit{string}(i)),\textit{delta}_2(j))$.
-%\State \textbf{goto} \emph{top}.
-%\EndProcedure
-%\end{algorithmic}
-%\end{algorithm}
-
 % For demonstration purposes, we also included another algorithm that uses S-Leaders for this section.

### 2017-12-03 LaTeX/Paper.tex
Paper: Re-wrote start of introduction to emphasize why annotations are important. Program: No changes.

@@
 
-\chapter{Introduction}%Wort ``Path'' durch nicht-definierte Terminologie ersetzen?
-Annotating a document is usually solved by adding footnotes in an appropriate position and adding a simple reference in the text, leaving the reader to find the referenced content by themselves.
-If a more obvious connection between the text and the referenced content is required, the reference placed to the side of the document and visibly connected to the text by drawing a straight line or a more complex path between them.
+\chapter{Introduction}
+Whenever additional information neets to be inserted into an existing document without altering the original text, we can make use of annotations.
+They usually take the form of footnotes, which require only a minimal reference in the main text, and are used for a variety of reasons - for example, to provide additional information that would hinder the text's flow if inserted directly, or as a result of a commenting tool that is used for communication between an author and their editor.
+If a more obvious connection between the text and the referenced content is required, for example when lengthy comments are added, or if a change-tracking tool is used, the reference is  placed to the side of the document and visibly connected to the part of the text it is referring to.
+This style of attotation is easily implemented on virtual documents, however if the annotations need to be included in a printed version, there are several issues that arise regarding readability of the final product and ambiguity of text-annotation assignments.
 
-In this paper, we will look at ways to use Boundary Labeling, which means that all annotations will be placed somewhere outside of the text they are referencing and will be visually connected to the feature they are referencing.
+In this paper, we will look at ways to use Boundary Labeling for this problem, which means that all annotations will be placed somewhere outside of the text they are referencing and will be visually connected to the feature they are referencing.
 (See also \cite{Bekos2007})
@@
 
-The paper by Loose\cite{Loose2015} on the other hand is based around only using the free space between lines and words, which produces longer paths, and forces curves, but doesn't obscure any part of the text.
+The thesis by Loose\cite{Loose2015} on the other hand is based around only using the free space between lines and words, which produces longer paths, and forces curves, but doesn't obscure any part of the text.
 %Siehe auhc andere Papers - Motivation! (Warum will ich mir dieses Thema ansehen)
@@
 \emph{P-Segments} run parallel to the border of the label area, and as such must be combined with other segments for the leader to reach its destination.
-\emph{S-Segments} are not required to have any particular orientation, and simply connect their start and einding points in a straight line.
+\emph{S-Segments} are not required to have any particular orientation, and simply connect their start and ending points in a straight line.
 The leader's name is created by combining the name of the segments - for example, the blue leader from Fig~\ref{fig:term} would be classified as an OPOPO-Leader.%Re-write once picture is updated

### 2017-12-13 LaTeX/Paper.tex
Paper: Added future chapter titles, small fixes. Program: No changes.

@@
 \usepackage[boxed,linesnumbered]{algorithm2e}
+\usepackage{color}
+%\usepackage[]{luatodonotes}
 %\usepackage{wrapfig}
@@
 \setcurriculum{Bachelor's programme Software \& Information Engineering}{Bachelorstudium Software \& Information Engineering} %TODO!
+
+\newcommand{\change}[1]{\textcolor{red}{#1}}
 
@@
 
-%TODO: Bezug auf Programm & Ergebnisse nehmen!
-
 \section{Terminology and Fundamentals} %Evtl in Introduction eingliedern?
 While Boundary Labeling (or an equivalent concept) can be applied to a space with different geometry or more dimensions, this paper will only concern itself with two-dimensonal, Euclidean space.
-To easily reference important concepts, some additional terminology will be introduced as well. (See Fig.~\ref{fig:term} for a visual explanation)%Add references to all terminology pictures!
-
-%Add figure for graph terminology.
+To easily reference important concepts, some additional terminology will be introduced as well. (See Fig.~\ref{fig:term} for a visual explanation)
 
@@
 \emph{S-Segments} are not required to have any particular orientation, and simply connect their start and ending points in a straight line.
-The leader's name is created by combining the name of the segments - for example, the blue leader from Fig~\ref{fig:term} would be classified as an OPOPO-Leader.%Re-write once picture is updated
+The leader's name is created by combining the name of the segments - for example, the blue leader from Fig~\ref{fig:term} would be classified as an OPOPO-Leader.
+\change{(Farben angepasst, hoffentlich ist es jetzt in Schwarz-Weiß besser zu erkennen)}
 The location where a leader connects to the label is called the \emph{port}. It can be restricted to pre-defined positions. %TODO: Re-insert monotonicity
@@
 The graph is constructed by placing vertices between the lines, located next to each corner of a word's bounding rectangle, with two consecutive words in a line sharing the two vertices associated with their adjacent corners.
-%Insert picture
 For the sites, we inserted an additional vertex above the center of the word, which will serve as the leader's starting point.
@@
 
+\DontPrintSemicolon
 \KwData{A single annotation's source and its Graph}
@@
   \Back $\gets$ True\;
-  \Repeat{\Curr's Position is below \Old or\; ~~~\Path is Empty}{
+  \Repeat{\Curr's Position is below \Old or \Path is Empty}{
    \Old $\gets$ \Curr\;
@@
   }
+  \;
   \If(//No path found){\Curr not below \Old}{\Break}
@@
 
-\section{Implementation}
+\chapter{Implementation}
 The program was written in Java 1.8.0u40, using JGraphT1.0.1\cite{JGraphT} as graph library.
@@
 Due to some problems with Java's various methods of calculating linebreaks, this process was done completely by hand.
+\section{Challenges}
 
@@
 Using this information alongside the positional data stored in the vertices, the polyline representing the leader can be drawn.
+
 %TODO: Schönheitsfeatures, etc.
@@
 
-%(Potential) Problems/Issues, solutions
+\chapter{Evaluation and Testing}
+
+\section{Data generation} \change{(Bin mit dem Titel unzufrieden, aber hier soll beschrieben werden, wie ich die einzelnen Testfälle erstellt habe.)}
+
+\section{Testing methods}
+
+\section{Results}
+
+\chapter{Conclusion}
+
+\section{Further notes}
 

### 2017-12-27 LaTeX/Paper.tex
Paper: Added Section 1.2 - Related Work, and changed parts of Chapter 1 to fit the new section. Program: No changes.

@@
 
-\setaddress{TODO: ADDRESS!}%TODO
+\setaddress{Scherzergasse 10/8, 1020 Wien}
 \setregnumber{1125755}
-\setdate{25}{09}{2017}
+\setdate{25}{09}{2017}%TODO: Update
 
@@
 They usually take the form of footnotes, which require only a minimal reference in the main text, and are used for a variety of reasons - for example, to provide additional information that would hinder the text's flow if inserted directly, or as a result of a commenting tool that is used for communication between an author and their editor.
-If a more obvious connection between the text and the referenced content is required, for example when lengthy comments are added, or if a change-tracking tool is used, the reference is  placed to the side of the document and visibly connected to the part of the text it is referring to.
-This style of attotation is easily implemented on virtual documents, however if the annotations need to be included in a printed version, there are several issues that arise regarding readability of the final product and ambiguity of text-annotation assignments.
+If a more obvious connection between the text and the referenced content is required, for example when lengthy comments are added, or if a change-tracking tool is used, the reference is often placed to the side of the document and visibly connected to the part of the text it is referring to.
+This style of annotation is easily implemented on virtual documents, since they can be hidden on demand, however if the annotations need to be included in a printed version, there are several issues that arise regarding readability of the final product and ambiguity of text-annotation assignments.
 
-In this paper, we will look at ways to use Boundary Labeling for this problem, which means that all annotations will be placed somewhere outside of the text they are referencing and will be visually connected to the feature they are referencing.
+In this thesis, we will look at ways to use Boundary Labeling for this problem, which means that all annotations will be placed somewhere outside of the text they are referencing and will be visually connected to the feature they are referencing.
 (See also \cite{Bekos2007})
@@
 
-While there are many papers discussing Boundary Labeling in general, only very few exist that apply this concept to written text.
-Generally, this approach isn't used very often, and tends to use simplistic algorithms which produce mediocre results.
-%Beispiele von suboptimalen Lösungen einfügen?
-(Bilder)
-However, the papers that do discuss boundary labeling in text offer interesting contributions.
-
-In the paper about the Luatodonotes-Package\cite{Kindermann2014} illustrates some of the different styles of drawing these connecting paths, and came to the conclusion that paths without bends are easier to follow.
-%Bild einfügen?
-However, most solutions proposed in that paper do not consider whether a path overlaps with text or not, which results in a decrease in readability.
-
-The thesis by Loose\cite{Loose2015} on the other hand is based around only using the free space between lines and words, which produces longer paths, and forces curves, but doesn't obscure any part of the text.
-%Siehe auhc andere Papers - Motivation! (Warum will ich mir dieses Thema ansehen)
-
 \section{Terminology and Fundamentals} %Evtl in Introduction eingliedern?
-While Boundary Labeling (or an equivalent concept) can be applied to a space with different geometry or more dimensions, this paper will only concern itself with two-dimensonal, Euclidean space.
+Boundary Labeling (or equivalent concepts) can be applied to a space with different geometry or more dimensions, but this thesis will only concern itself with two-dimensonal, Euclidean space.
 To easily reference important concepts, some additional terminology will be introduced as well. (See Fig.~\ref{fig:term} for a visual explanation)
@@
 The leader's name is created by combining the name of the segments - for example, the blue leader from Fig~\ref{fig:term} would be classified as an OPOPO-Leader.
-\change{(Farben angepasst, hoffentlich ist es jetzt in Schwarz-Weiß besser zu erkennen)}
 The location where a leader connects to the label is called the \emph{port}. It can be restricted to pre-defined positions. %TODO: Re-insert monotonicity
@@
 %+Node für Knoten in Polylinien?
+
+\section{Related Work}
+
+While there are many papers discussing Boundary Labeling in general, only very few exist that apply this concept to written text.
+Generally, this approach isn't used very often, and tends to use simplistic algorithms which produce mediocre results.
+%Beispiele von suboptimalen Lösungen einfügen?
+(Bilder)
+
+The concept of boundary labeling is introduced in \cite{Bekos2007}, where several approaches with varying leader types are discussed, along with any problems that arise. 
+However, the papers that do discuss boundary labeling in text offer interesting contributions.
+
+The paper about the Luatodonotes-Package\cite{Kindermann2014} illustrates some of the different styles of drawing these connecting paths, and came to the conclusion that paths without bends are easier to follow.
+%Bsp-Bild einfügen!
+However, most solutions proposed in that paper do not consider whether a path overlaps with text or not, which results in a decrease in readability.
+
+The thesis by Loose\cite{Loose2015} on the other hand is based around only using the free space between lines and words, which produces longer paths, and forces curves, but doesn't obscure any part of the text.
 

### 2017-12-29 LaTeX/Paper.tex
Paper: Added additional references, extended section 'Related Work'. Program: No changes.

@@
 \emph{Labels} hold additional information and are represented as boxes containing this information.
-They are usually placed in the \emph{Label area} which is a rectangular area designated to hold labels.
+They are usually placed in the \emph{label area} which is a rectangular area designated to hold labels.
 It is located next to of the bigger \emph{text area}, which contains the document's text and all \emph{sites}, the points or objects that a label's information refers to.
@@
 
-While there are many papers discussing Boundary Labeling in general, only very few exist that apply this concept to written text.
+Boundary Labeling was first introduced by Bekos et al.
+in 2006 (see \cite{Bekos2007}), where several approaches with different leader types are looked into.
+Since then, several papers have been written about this topic, such as \cite{Barth2015}, which looks into the readability of different leader styles.
+Interestingly, some leader styles perform quite well, despite the study's participants preferring others over them, with OPO-Leaders being both least preferred and the hardest to follow.
+%Add sth about NP-Hardness of 2-side label placement!
+%While there are many papers discussing Boundary Labeling in general, only very few exist that apply this concept to written text.
 Generally, this approach isn't used very often, and tends to use simplistic algorithms which produce mediocre results.
-%Beispiele von suboptimalen Lösungen einfügen?
-(Bilder)
 
-The concept of boundary labeling is introduced in \cite{Bekos2007}, where several approaches with varying leader types are discussed, along with any problems that arise. 
-However, the papers that do discuss boundary labeling in text offer interesting contributions.
+%Beispiele von suboptimalen Lösungen einfügen? (Bilder)
+Boundary Labeling in text documents however, is rarely discussed, and only few papers exist about this topic.
+The programs that employ this style of annotation often also use rather simple algorithms, to mediocre results.
+However, the few papers that discuss this topic add interesting information to the discussion.
+%Rewrite Sentence, I don't like it!
 
-The paper about the Luatodonotes-Package\cite{Kindermann2014} illustrates some of the different styles of drawing these connecting paths, and came to the conclusion that paths without bends are easier to follow.
-%Bsp-Bild einfügen!
+The paper about the Luatodonotes-Package\cite{Kindermann2014} uses several styles of drawing leaders, and came to the conclusion that leaders without bends are easier to follow. %Bsp-Bild einfügen!
 However, most solutions proposed in that paper do not consider whether a path overlaps with text or not, which results in a decrease in readability.
 
-The thesis by Loose\cite{Loose2015} on the other hand is based around only using the free space between lines and words, which produces longer paths, and forces curves, but doesn't obscure any part of the text.
+The thesis by Loose\cite{Loose2015} on the other hand is based around only using the free space between lines and words, which produces longer leaders, and forces curves, but doesn't obscure any part of the text.
+It uses a graph-based algorithm to model the possible leader positions.
+
+The paper by Lin et al.\cite{Lin2009} uses only OPO-Leaders that have their P-Segment located outside of the text area, but allows the leaders to use the text area's border that's opposite of the label area to route upwards or down to place more labels as close as possible to their leader's source, at the cost of increasing select leaders' length.
 

### 2018-01-02 LaTeX/Paper.tex
Paper: Found earlier version of Bekos et al.'s article. Extended 'Related Work'-Section. Added new definition for one- and multi-sided labeling. Program: No changes.

@@
 In this thesis, we will look at ways to use Boundary Labeling for this problem, which means that all annotations will be placed somewhere outside of the text they are referencing and will be visually connected to the feature they are referencing.
-(See also \cite{Bekos2007})
+(See also \cite{Bekos2005})
 
@@
 It is located next to of the bigger \emph{text area}, which contains the document's text and all \emph{sites}, the points or objects that a label's information refers to.
-The site and the label are connected via a \emph{leader}, a polyline that can be further classified by looking at the orientation of its segments: \emph{O-Segments} run orthogonally to the border of the label area.
+\change{If multiple label areas exist on different sides of the text area, we speak of \emph{multi-sided labeling}, otherwise we speak of \emph{one-sided labeling}.} The site and the label are connected via a \emph{leader}, a polyline that can be further classified by looking at the orientation of its segments: \emph{O-Segments} run orthogonally to the border of the label area.
 \emph{P-Segments} run parallel to the border of the label area, and as such must be combined with other segments for the leader to reach its destination.
@@
 
-Boundary Labeling was first introduced by Bekos et al.
-in 2006 (see \cite{Bekos2007}), where several approaches with different leader types are looked into.
-Since then, several papers have been written about this topic, such as \cite{Barth2015}, which looks into the readability of different leader styles.
+Boundary labeling was first introduced by Bekos et al.
+in 2004 (see \cite{Bekos2005}), where both one-sided and multi-sided labelings with different leader types are looked into.
+They also showed that the optimal placement of arbitrarily-sized labels on two sides of the text area can be NP-hard by drawing comparisons to the Partition-Problem.
+However, a pseudo-polynomial solution exists for this problem, which was adapted to this variation of the problem.
+
+Since then, several papers have been written about boundary labeling.
+One of these is \cite{Barth2015}, which looks into the readability of different leader styles.
 Interestingly, some leader styles perform quite well, despite the study's participants preferring others over them, with OPO-Leaders being both least preferred and the hardest to follow.
-%Add sth about NP-Hardness of 2-side label placement!
-%While there are many papers discussing Boundary Labeling in general, only very few exist that apply this concept to written text.
+
+Another article using boundary labeling is \cite{Goetzelmann2006} by G{\"o}etzelmann et al., which creates boundary labeling-style annotations along other methods to label different parts of three-dimensional figures, resulting in pictures similar to what could be found in a textbook.
+As this algorithm works in real-time, it is suitable for labeling interactive models and allows for user interaction.
+
+%While there are many papers discussing boundary labeling in general, only very few exist that apply this concept to written text.
 Generally, this approach isn't used very often, and tends to use simplistic algorithms which produce mediocre results.
@@
 %Beispiele von suboptimalen Lösungen einfügen? (Bilder)
-Boundary Labeling in text documents however, is rarely discussed, and only few papers exist about this topic.
-The programs that employ this style of annotation often also use rather simple algorithms, to mediocre results.
-However, the few papers that discuss this topic add interesting information to the discussion.
+Boundary labeling in text documents however, is rarely discussed, and only few papers exist about this topic.
+The programs that employ this style of annotation often also use rather simple algorithms, to mediocre results or make extensive use of the interactivity of a digital medium, showing annotations only on demand.
+However, the few papers that approach this topic add interesting information to the discussion.
 %Rewrite Sentence, I don't like it!
 
-The paper about the Luatodonotes-Package\cite{Kindermann2014} uses several styles of drawing leaders, and came to the conclusion that leaders without bends are easier to follow. %Bsp-Bild einfügen!
-However, most solutions proposed in that paper do not consider whether a path overlaps with text or not, which results in a decrease in readability.
+The paper about the Luatodonotes-Package\cite{Kindermann2014} uses several styles of drawing leaders, and came to the conclusion that leaders without bends are easier to follow, which fits with \cite{Barth2015}'s observations, which ranks OPO- and PO-Leaders lower than other variants.
+However, most solutions proposed in \cite{Kindermann2014} do not consider whether a path overlaps with text or not, which results in a decrease in readability.%Bsp-Bild einfügen!
 
@@
 
-The paper by Lin et al.\cite{Lin2009} uses only OPO-Leaders that have their P-Segment located outside of the text area, but allows the leaders to use the text area's border that's opposite of the label area to route upwards or down to place more labels as close as possible to their leader's source, at the cost of increasing select leaders' length.
+Lin et al.\cite{Lin2009} use only OPO-Leaders that have their P-Segment located outside of the text area in their paper, but allow the leaders to use the text area's border on the opposite side of the label area to route upwards or down.
+This allows for more labels to be placed as close as possible to their leader's source, at the cost of increasing select leaders' length and placing some labels out of order.
 
@@
 
-\section{Data generation} \change{(Bin mit dem Titel unzufrieden, aber hier soll beschrieben werden, wie ich die einzelnen Testfälle erstellt habe.)}
+\section{Data generation} %\change{(Bin mit dem Titel unzufrieden, aber hier soll beschrieben werden, wie ich die einzelnen Testfälle erstellt habe.)}
 

### 2018-01-04 LaTeX/Paper.tex
Paper: Finished Section 1.2; added definition of 'Routing Area'. Program: No changes.

@@
 They are usually placed in the \emph{label area} which is a rectangular area designated to hold labels.
-It is located next to of the bigger \emph{text area}, which contains the document's text and all \emph{sites}, the points or objects that a label's information refers to.
-\change{If multiple label areas exist on different sides of the text area, we speak of \emph{multi-sided labeling}, otherwise we speak of \emph{one-sided labeling}.} The site and the label are connected via a \emph{leader}, a polyline that can be further classified by looking at the orientation of its segments: \emph{O-Segments} run orthogonally to the border of the label area.
+It is located next to of the bigger, rectangular \emph{text area}, which contains the document's text and all \emph{sites}, the points or objects that a label's information refers to.
+\change{If multiple label areas exist on different sides of the text area, we speak of \emph{multi-sided labeling}, otherwise we speak of \emph{one-sided labeling}.
+We will be using one-sided labeling in our implementation.
+Some space was left in between the text and label area, to make connecting sites to their labels easier, which we will call the \emph{routing area}.} The site and the label are connected via a \emph{leader}, a polyline that can be further classified by looking at the orientation of its segments: \emph{O-Segments} run orthogonally to the border of the label area.
 \emph{P-Segments} run parallel to the border of the label area, and as such must be combined with other segments for the leader to reach its destination.
@@
 
-%While there are many papers discussing boundary labeling in general, only very few exist that apply this concept to written text.
-Generally, this approach isn't used very often, and tends to use simplistic algorithms which produce mediocre results.
-
 %Beispiele von suboptimalen Lösungen einfügen? (Bilder)
@@
 The paper about the Luatodonotes-Package\cite{Kindermann2014} uses several styles of drawing leaders, and came to the conclusion that leaders without bends are easier to follow, which fits with \cite{Barth2015}'s observations, which ranks OPO- and PO-Leaders lower than other variants.
-However, most solutions proposed in \cite{Kindermann2014} do not consider whether a path overlaps with text or not, which results in a decrease in readability.%Bsp-Bild einfügen!
+However, most solutions proposed in \cite{Kindermann2014} do not consider whether a path overlaps with text or not, which results in a decrease in readability.
+While we do not use the routing and leader styles introduced in this paper, the results can be used in comparisons regarding readability of the main text and ease of use of the different leader styles.%Bsp-Bild einfügen!
 
 The thesis by Loose\cite{Loose2015} on the other hand is based around only using the free space between lines and words, which produces longer leaders, and forces curves, but doesn't obscure any part of the text.
-It uses a graph-based algorithm to model the possible leader positions.
+The two different approaches in this paper were a clustering-based algorithm, which was previously described in \cite{Nollenburg2010}, and a flow network-based approach.
+Several concepts of this paper, such as the graph-based strategy and the usage of a routing area will be adopted in our thesis and it is by far the biggest influence on our approach to the problem.
+%Anmerken, dass viele Konzepte hier aufgegriffen werden!
 
@@
 This allows for more labels to be placed as close as possible to their leader's source, at the cost of increasing select leaders' length and placing some labels out of order.
+While this is an interesting way to avoid longer leaders in general, it is hard to combine with the graph-based routing that happens inside the text area, so we won't make use of it.
 

### 2018-01-11 LaTeX/Paper.tex
Paper: Expanded 2.1 - Problem specification. Program: No changes.

@@
 \chapter{The Algorithm}
+%Evtl 1 Zeile Einleitung hier einfügen
+\section{Problem Specification}
+In our algorithm, the focus was put on keeping the text as readable as possible.
+This means that leaders aren't allowed to pass through words - they can only use the space between words and in between lines.
+Additionally, we wanted to maximize the usage of the label area's space, even if we have to move labels further from their sites than needed. 
+Finally, we wanted to keep the leader's length low for any given combination of labels and site placements, so every segment of a leader has to decrease the remaining distance between the site and its label's port.
 
-In our algorithm, the focus was put on keeping the text as readable as possible.
-This means that leaders aren't allowed to pass through words, which was implemented by enclosing each word in a bounding rectangle that leaders aren't allowed to pass through.
-The leaders should also be kept as short as possible, so we restricted them to always be moving toward the label, even if the direct route is unavailable.%Wie drücke ich das als Formel aus?
-We also wanted to use the space available in the labeling area as efficiently as possible, so labels are placed as far up as possible in order to maximize the space available to future labels at the cost of increased leader length.
+These properties were modelled by creating a rectangle around each word that no leaders may pass through.
+If this is done for all words in a paragraph, the remaining available space for routing resembles a grid of perfectly parallel streets.
+\change{NOTE: Insert example image for this and graph.} Since the layout encourages the usage of P- and O-segments, we decided to represent the possible routes by creating a graph which has its vertices located whenever multiple directions are available - in between lines, above or below whitespace characters.
+The graph's edges represent the different possible segments of a leader and connect each vertex with the closest vertices directly above, below or to the sides, given that no text is blocking the connection.
+\change{The resulting graph looks similar to the one in Fig.~<TODO>}
+
+For ease of visualisation, we limited the number of leaders per edge to one, and forbid the leaders to cross one another, as this could be mistaken for two curves touching each other at the corners.
+We also decided to place the annotations in the order in which they appear in the text, and always place their port above the leader's source, which means that crossings are inevitable if a leader is located above and to the left of an earlier leader.
+To ensure this never happens, leaders aren't allowed to enter the top left quadrant of a source vertex, unless no routing could be found for that vertex, which means there won't be any leaders to collide with.
+Additionally, to keep the leader's length minimal, each leader is only allowed to move towards the source, which places a hard limit on how far up a leader is allowed to go, and restricts the possible choices at each node to edges that allow the leader to move in that direction.
+To maximize the space available to subsequent annotations, each label is placed as far up as possible.
+This also means that if there's no placement available that is above its leader's source, the annotation cannot be placed, as we previously restricted the leader's movement.
+
+%This was implemented by enclosing each word in a bounding rectangle that leaders aren't allowed to pass through.
+%The leaders should also be kept as short as possible, so we restricted them to always be moving toward the label, even if the direct route is unavailable.%Wie drücke ich das als Formel aus?
+%We also wanted to use the space available in the labeling area as efficiently as possible, so labels are placed as far up as possible in order to maximize the space available to future labels at the cost of increased leader length.
+
+\section{Description}
 
@@
    \V$\gets$\NewVert{\PrevW.\getC{}}\;
-   \V.\setAnn{w}\;
+   \V.\setAnn{\W}\;
    \G.\AddVert{\V}\;

### 2018-01-12 LaTeX/Paper.tex
Paper: Added FromTextToGraph.png (needs further work) and included references to it. Program: No changes.

@@
 
+\begin{figure}
+ \captionsetup{justification=centering, margin=0.75cm}
+ \centering
+  \includegraphics[scale=0.85]{FromTextToGraph.png}
+  \caption{Visualization of the space reserved for the text, and the resulting graph.}
+ \label{fig:ttg}
+\end{figure}
+
 These properties were modelled by creating a rectangle around each word that no leaders may pass through.
-If this is done for all words in a paragraph, the remaining available space for routing resembles a grid of perfectly parallel streets.
-\change{NOTE: Insert example image for this and graph.} Since the layout encourages the usage of P- and O-segments, we decided to represent the possible routes by creating a graph which has its vertices located whenever multiple directions are available - in between lines, above or below whitespace characters.
+If this is done for all words in a paragraph, the remaining available space for routing resembles a grid, like in Fig.\ref{fig:ttg}a.
+
+Since the layout encourages the usage of P- and O-segments, we decided to represent the possible routes by creating a graph which has its vertices located whenever multiple directions are available - in between lines, above or below whitespace characters.
 The graph's edges represent the different possible segments of a leader and connect each vertex with the closest vertices directly above, below or to the sides, given that no text is blocking the connection.
-\change{The resulting graph looks similar to the one in Fig.~<TODO>}
+The resulting graph looks similar to the one in Fig.~\ref{fig:ttg}b.
 

### 2018-01-20 LaTeX/Paper.tex
Paper: Re-wrote Section 2.1, changed Fig. 2.1. Program: No changes.

--- 2018-01-04 LaTeX/Paper.tex
@@
 \usepackage{color}
+\usepackage{subcaption}
 %\usepackage[]{luatodonotes}
@@
 It is located next to of the bigger, rectangular \emph{text area}, which contains the document's text and all \emph{sites}, the points or objects that a label's information refers to.
-\change{If multiple label areas exist on different sides of the text area, we speak of \emph{multi-sided labeling}, otherwise we speak of \emph{one-sided labeling}.
+If multiple label areas exist on different sides of the text area, we speak of \emph{multi-sided labeling}, otherwise we speak of \emph{one-sided labeling}.
 We will be using one-sided labeling in our implementation.
-Some space was left in between the text and label area, to make connecting sites to their labels easier, which we will call the \emph{routing area}.} The site and the label are connected via a \emph{leader}, a polyline that can be further classified by looking at the orientation of its segments: \emph{O-Segments} run orthogonally to the border of the label area.
+Some space was left in between the text and label area, to make connecting sites to their labels easier, which we will call the \emph{routing area}.
+The site and the label are connected via a \emph{leader}, a polyline that can be further classified by looking at the orientation of its segments: \emph{O-Segments} run orthogonally to the border of the label area.
 \emph{P-Segments} run parallel to the border of the label area, and as such must be combined with other segments for the leader to reach its destination.
@@
 However, the few papers that approach this topic add interesting information to the discussion.
-%Rewrite Sentence, I don't like it!
 
@@
 \chapter{The Algorithm}
+%1 Zeile Einleitung hier einfügen - zb: Kapitelübersicht
+\section{Problem Specification}
+We limited the leaders to use exclusively O- and P-Segments, and banned them from passing through words.
+We also place labels as far up as possible to maximize the space remaining for  remaining placements.
+Additionally, a leader isn't allowed to be any longer as is necessary to connect a given Site with its label's port.
 
-In our algorithm, the focus was put on keeping the text as readable as possible.
-This means that leaders aren't allowed to pass through words, which was implemented by enclosing each word in a bounding rectangle that leaders aren't allowed to pass through.
-The leaders should also be kept as short as possible, so we restricted them to always be moving toward the label, even if the direct route is unavailable.%Wie drücke ich das als Formel aus?
-We also wanted to use the space available in the labeling area as efficiently as possible, so labels are placed as far up as possible in order to maximize the space available to future labels at the cost of increased leader length.
+These restrictions are implemented as follows: We divided the text $T$ into separate lines $L=\{W,H\}$ which are in turn made up of words $W$ and whitespace $H$.
+Each word $w \in W$ is surrounded by a bounding rectangle $r$, %Definition schreiben?
+which marks the space leaders aren't allowed to cross. For a graphical representation see Fig.~\ref{fig:wbound}.
+
+\begin{figure}
+ \centering
+ \begin{subfigure}[b]{\textwidth}
+ \centering
+  \includegraphics[]{WordBoundaries.png}
+  \caption{\label{fig:wbound}}
+ \end{subfigure}
+ \\
+ ~\\%There's probably a better way for this...
+ \begin{subfigure}[b]{\textwidth}
+ \centering
+  \includegraphics[]{RoutingGraph_edited.png}
+  \caption{\label{fig:rgraph}}
+ \end{subfigure}
+ \caption{Visualization of the space reserved for the text, and the resulting graph.}
+\end{figure}
+
+Since we only allowed the Usage of O- and P-Segments in leaders, the only way for a leader to cross through a line of text is with a P-Segment through whitespace, whereas O-Segments are only usable between lines.
+Therefore, we can create a routing graph $G=\langle V,E\rangle$ whose Edges $E$ reflect the legal paths a leader can take within $T$.
+For each whitespace character $w \in W$, one vertex is placed above and below, so that it has the maximum distance from both of its adjacent lines.
+The start and end of each line also recieve a pair of vertices each.
+Sites will be represented by the insertion of additional vertices at appropriate places.
+The edges between the vertices are all perfectly horizontal or vertical, and do neither intersect with any bounding rectangle, nor any vertices other than their starting and ending vertex.The resulting graph looks similar to Fig.~\ref{fig:rgraph}.
+
+Each edge has a capacity of 1, meaning that no more than one leader is allowed to pass through it.
+Leaders also aren't allowed to cross one another, as it is hard to discern between intersections and two leaders curving away from each other.
+This is realized by forcing leaders to incorporate as many P-Segments as early as possible while still following all previous restrictions, and forbidding leaders from entering the space to the top left of already successfully routed leader's sites.
+%In our algorithm, the focus was put on keeping the text as readable as possible.
+This means that leaders aren't allowed to pass through words - they can only use the space between words and in between lines.
+%Additionally, we wanted to maximize the usage of the label area's space, even if we have to move labels further from their sites than needed. 
+%Finally, we wanted to keep the leader's length low for any given combination of labels and site placements, so every segment of a leader has to decrease the remaining distance between the site and its label's port.
+
+%\begin{figure}
+% \captionsetup{justification=centering, margin=0.75cm}
+% \centering
+%  \makebox[\textwidth][c]{\includegraphics[scale=0.9]{FromTextToGraph.png}}
+%  \caption{Visualization of the space reserved for the text, and the resulting graph.}
+% \label{fig:ttg}
+%\end{figure}
+
+%These properties were modelled by creating a rectangle around each word that no leaders may pass through.
+If this is done for all words in a paragraph, the remaining available space for routing resembles a grid, like in Fig.\ref{fig:ttg}a.
+
+%Since the layout encourages the usage of P- and O-segments, we decided to represent the possible routes by creating a graph which has its vertices located whenever multiple directions are available - in between lines, above or below whitespace characters.
+The graph's edges represent the different possible segments of a leader and connect each vertex with the closest vertices directly above, below or to the sides, given that no text is blocking the connection.
+The resulting graph looks similar to the one in Fig.~\ref{fig:ttg}b.
+
+%For ease of visualisation, we limited the number of leaders per edge to one, and forbid the leaders to cross one another, as this could be mistaken for two leaders touching each other at the corners.
+We also decided to place the annotations in the order in which they appear in the text, and always place their port above the leader's source, which means that crossings are inevitable if a leader is located above and to the left of an earlier leader.
+To ensure this never happens, leaders aren't allowed to enter the top left quadrant of a source vertex, unless no routing could be found for that vertex, which means there won't be any leaders to collide with.
+Additionally, to keep the leader's length minimal, each leader is only allowed to move towards the source, which places a hard limit on how far up a leader is allowed to go, and restricts the possible choices at each node to edges that allow the leader to move in that direction.
+%To maximize the space available to subsequent annotations, each label is placed as far up as possible.
+This also means that if there's no placement available that is above its leader's source, the annotation cannot be placed, as we previously restricted the leader's movement.
+
+%This was implemented by enclosing each word in a bounding rectangle that leaders aren't allowed to pass through.
+%The leaders should also be kept as short as possible, so we restricted them to always be moving toward the label, even if the direct route is unavailable.%Wie drücke ich das als Formel aus?
+%We also wanted to use the space available in the labeling area as efficiently as possible, so labels are placed as far up as possible in order to maximize the space available to future labels at the cost of increased leader length.
+
+\section{Description}
 

### 2018-01-23 LaTeX/Paper.tex
Paper:Added suggested changes to section 2.1. Program: No changes.

@@
 These restrictions are implemented as follows: We divided the text $T$ into separate lines $L=\{W,H\}$ which are in turn made up of words $W$ and whitespace $H$.
-Each word $w \in W$ is surrounded by a bounding rectangle $r$, %Definition schreiben?
-which marks the space leaders aren't allowed to cross. For a graphical representation see Fig.~\ref{fig:wbound}.
+The remaining space between lines shall be defined as $S(l_1, l_2)$%improve on this!
+with $l_1,l_2 \in L$ and will have no restrictions regarding leader placement.
+For each word $w \in W$, we define $R(w)$  as its bounding rectangle, which marks the space leaders aren't allowed to cross.
+For a graphical representation see Fig.~\ref{fig:wbound}.
 
@@
 Therefore, we can create a routing graph $G=\langle V,E\rangle$ whose Edges $E$ reflect the legal paths a leader can take within $T$.
-For each whitespace character $w \in W$, one vertex is placed above and below, so that it has the maximum distance from both of its adjacent lines.
+For each whitespace character $w_x \in W$ in a given line $l_i \in L$, one vertex is placed in $S(l_{i-1},l_i)$ and $S(l_i,l_{i+1})$ each, located above or below the whitespace character, and with a maximum distance from each of its neighbouring lines.
 The start and end of each line also recieve a pair of vertices each.
-Sites will be represented by the insertion of additional vertices at appropriate places.
-The edges between the vertices are all perfectly horizontal or vertical, and do neither intersect with any bounding rectangle, nor any vertices other than their starting and ending vertex.The resulting graph looks similar to Fig.~\ref{fig:rgraph}.
+Sites will be represented by the insertion of additional vertices in $S(l_i, l_{i-1})$ on a similar height as those next to whitespace of the same line and located directly above the center point of its words $w_j$ bounding rectangle $R(w_j)$.
+Should either $l_{i-1}$ or $l_{i+1}$ not exist, the vertices' vertical placement will be similar to that of the previous or following line.
+The edges between the vertices are all perfectly horizontal or vertical, and do neither intersect with any bounding rectangle, nor any vertices other than their starting and ending vertex.
+The resulting graph looks similar to Fig.~\ref{fig:rgraph}.
 

### 2018-01-24 LaTeX/Paper.tex
Paper: Started Re-work of Section 2.2. Program: No changes.

@@
 Therefore, we can create a routing graph $G=\langle V,E\rangle$ whose Edges $E$ reflect the legal paths a leader can take within $T$.
-For each whitespace character $w_x \in W$ in a given line $l_i \in L$, one vertex is placed in $S(l_{i-1},l_i)$ and $S(l_i,l_{i+1})$ each, located above or below the whitespace character, and with a maximum distance from each of its neighbouring lines.
+For each whitespace character $h_x \in W$ in a given line $l_i \in L$, one vertex is placed in $S(l_{i-1},l_i)$ and $S(l_i,l_{i+1})$ each, located above or below the whitespace character, and with a maximum distance from each of its neighbouring lines.
 The start and end of each line also recieve a pair of vertices each.
@@
 \end{algorithm}
-To create leaders that exclusively use the space not taken up by a word's bounding rectangle, we decided to use a graph similar to the one Loose \cite{Loose2015} used.
-As each vertex represents a physical location, they will have co-ordinates associated with them, and the vertices representing the sites will hold additional information regarding its leader and label.
-The graph is constructed by placing vertices between the lines, located next to each corner of a word's bounding rectangle, with two consecutive words in a line sharing the two vertices associated with their adjacent corners.
-For the sites, we inserted an additional vertex above the center of the word, which will serve as the leader's starting point.
+
+In our Algorithm we first create the routing graph that will be used from here onward.
+To achieve this, we ge through the annotated text word by word, and measure the length $le(w_i)$ of each word $w_i \in l_j$ that is not part of an annotation to determine its placement in the text.
+Afterwards, we create new vertices above and below the whitespace characters $h_k, h_{k+1}$ preceding and succeeding $w_i$.
+These vertices are located in the space $S(l_{j-1},l_j), S(l_j,l_{j+1})$ between the word's line and its surrounding lines $l_{j-1}$ and $l_{j+1}$, as described in the previous section.
+Whenever an annotation is encountered, a single vertex is added above the last non-annotation word in $S(l_{j-1},l_j)$ at $pos(h_k)+le(w_i)/2$.
+For a representation of this process in pseudocode, see Alg.~\ref{alg:GraphCreation}.
+%To create leaders that exclusively use the space not taken up by a word's bounding rectangle, we decided to use a graph similar to the one Loose \cite{Loose2015} used.
+%As each vertex represents a physical location, they will have co-ordinates associated with them, and the vertices representing the sites will hold additional information regarding its leader and label.
+%The graph is constructed by placing vertices between the lines, located next to each corner of a word's bounding rectangle, with two consecutive words in a line sharing the two vertices associated with their adjacent corners.
+%For the sites, we inserted an additional vertex above the center of the word, which will serve as the leader's starting point.
 After placing all vertices, we created edges between each vertex and the closest horizontal neighbour to both sides, and between nodes that are located exactly above or below each other, and exactly one line apart.
@@
 
-In our labeling algorithm, we decided to work through the labels in the order they appear in the text, placing each as far up as possible, and skipping any label that would've required us place the label below its leader's source node.
-We also opted to use fixed ports located in the top left corner of each label, as this allowed us to unabiguously place annotations by only knowing their port location, which is equal to their leader's ending point.
-To not restrict the label's placement by the line spacing, we left a buffer zone between the text area and the label area, which allows us to place labels even further above to optimally use the available space.
-The algorithm's input consists of the Graph $G$ and a set of sites $V_{ann} \subset V$ which will be routed using a depth-first search algorithm, restricted to only selecting vertices located above or to the right of the current vertex, prioritizing moving up whenever possible.
-After the algorithm terminates, it returns a set with routing information for each vertex, which contains a Path $P=\{v_1, \cdots, v_n\}$, leading from the site to the text area's border, along with some additional information on how to draw the OPO-Segment that connects to the Label's port.
+For the routing part of the algorithm, we work through the annotations $a \in A$ in the order they appear in the text, placing each as far up as possible, skipping any annotation that cannot be placed above its leader's source node.
+We will also use fixed ports located in the top left corner of each label.
+This allows us to place annotations by only knowing their port location, which can be inferred from their leader's ending point.
+The routing area between the text and label areas will be used to bridge the gap between the highest reachable point within the routing graph and the highest possible label placement by connecting the two via an OPO-leader.%Kann man das so sagen?
+
+%TODO: Expand/rework this!
+The algorithm's input consists of the Graph $G$ and a set of sites $V_{ann} \subset V$ which will be routed using a depth-first search algorithm, restricted to only selecting vertices located above or to the right of the current vertex $v_{curr}$, prioritizing moving up whenever possible.
+After the algorithm terminates, it returns a set with routing information for each vertex, which contains a Path $P=\{v_1, \cdots, v_n\}$, leading from the site to the text area's border, along with information on how to draw the OPO-Segment that connects to the Label's port.
 If the routing for any given site failed, the path consists of a single vertex - the site.

### 2018-01-25 LaTeX/Paper.tex
Paper: Finished initial re-write of Section 2.2. Program: No changes.

@@
 
-In our Algorithm we first create the routing graph that will be used from here onward.
-To achieve this, we ge through the annotated text word by word, and measure the length $le(w_i)$ of each word $w_i \in l_j$ that is not part of an annotation to determine its placement in the text.
+First, we create the graph that will be used for the routing part of the algorithm.
+To achieve this, we go through the annotated text word by word in reading order, and measure the length $le(w_i)$ of each word $w_i$ that is not part of an annotation to determine its placement in the text.
+At this point, we can also assign a line $l_j$ to it.
 Afterwards, we create new vertices above and below the whitespace characters $h_k, h_{k+1}$ preceding and succeeding $w_i$.
@@
 We will also use fixed ports located in the top left corner of each label.
-This allows us to place annotations by only knowing their port location, which can be inferred from their leader's ending point.
-The routing area between the text and label areas will be used to bridge the gap between the highest reachable point within the routing graph and the highest possible label placement by connecting the two via an OPO-leader.%Kann man das so sagen?
+%This allows us to place annotations by only knowing their port location, which can be inferred from their leader's ending point.
+The routing area, which is located between the text and label areas, will be used to bridge the gap between the highest reachable point within the routing graph and the highest possible label placement by connecting the two via an OPO-leader.%Kann man das so sagen?
 
-%TODO: Expand/rework this!
-The algorithm's input consists of the Graph $G$ and a set of sites $V_{ann} \subset V$ which will be routed using a depth-first search algorithm, restricted to only selecting vertices located above or to the right of the current vertex $v_{curr}$, prioritizing moving up whenever possible.
-After the algorithm terminates, it returns a set with routing information for each vertex, which contains a Path $P=\{v_1, \cdots, v_n\}$, leading from the site to the text area's border, along with information on how to draw the OPO-Segment that connects to the Label's port.
+The routing algorithm itself uses only the Graph $G$ and the set of sites $V_{ann} \subset V$ which will be routed using a depth-first search algorithm, restricted to only selecting vertices located above or to the right of the current vertex $v_{curr}$, prioritizing moving up whenever possible.
+It will terminate upon reaching the right text border or if all possible paths failed to reach the text border.
+At this point, it will return a set with routing information for each vertex, which contains a Path $P=\{v_1, \cdots, v_n\}$, leading from the site to the text area's border, along with information on how to draw the OPO-Segment that connects to the Label's port.
 If the routing for any given site failed, the path consists of a single vertex - the site.

### 2018-01-27 LaTeX/Paper.tex
Paper: Edited Fig. 1.1 to include the Routing Area; minor corrections. Program: No changes.

@@
  \centering
-  \includegraphics[scale=0.90]{GraphTerminologyExtended.png}
+  \includegraphics[scale=0.95]{GraphTerminologyExtended.png}
   \caption{Illustrated guide to the labeling terminology}
@@
 
-For the routing part of the algorithm, we work through the annotations $a \in A$ in the order they appear in the text, placing each as far up as possible, skipping any annotation that cannot be placed above its leader's source node.
+For the routing part of the algorithm, we work through the annotations $a \in A$ in the order they appear in the text, placing each as far up as possible, skipping any annotation that cannot be placed in a reachable position above its leader's source.
 We will also use fixed ports located in the top left corner of each label.
 %This allows us to place annotations by only knowing their port location, which can be inferred from their leader's ending point.
-The routing area, which is located between the text and label areas, will be used to bridge the gap between the highest reachable point within the routing graph and the highest possible label placement by connecting the two via an OPO-leader.%Kann man das so sagen?
+The routing area, which is located between the text and label areas, will be used to bridge the gap between the highest currently reachable point within the routing graph and the highest possible label placement by connecting the two via an OPO-leader.%Kann man das so sagen?
 
-The routing algorithm itself uses only the Graph $G$ and the set of sites $V_{ann} \subset V$ which will be routed using a depth-first search algorithm, restricted to only selecting vertices located above or to the right of the current vertex $v_{curr}$, prioritizing moving up whenever possible.
+The routing algorithm itself uses only the Graph $G$ and the set of sites $V_{ann} \subset V$ which will be routed using a depth-first search algorithm, using only vertices located above or to the right of the current vertex $v_{curr}$, prioritizing moving up whenever possible.
 It will terminate upon reaching the right text border or if all possible paths failed to reach the text border.
@@
 \KwData{A single annotation's source and its Graph}
-\KwResult{A List of vertices describing the Leader's Path}
+\KwResult{A List of vertices describing the leader's path}
 

### 2018-02-05 LaTeX/Paper.tex
Paper: Added Definition of monotonicity to 1.2, fixed typo in 2.1. ProNo changes.

@@
  \emph{Depth-first search} is a searching algorithm on a graph G that starts at a given vertex $v \in V$ and explores the graph by traversing its edges as far as possible before backtracking, and continues to do so until a pre-defined goal is met.
-%In our algorithm, the goal is to reach a target vertex $v_t \in V$, at which point the algorithm terminates and returns the route taken from $v_s$ to $v_t$.
+\change{A function is called \emph{monotone} if it is either non-decreasing or non-increasing, meaning that for each $x<y$, $f(x) \leq f(y)$ for \emph{monotonically increasing functions}, or $f(x) \geq f(y)$ for \emph{monotonically decreasing} functions.}
 % A search algorithm is \emph{monotonous} if, for each vertex $v$ and each successor $v'$ generated by the algorithm, the estimated cost of reaching the goal from $v$ is not greater than the cost of reaching $v'$ plus the estimated cost of reaching the goal from $v'$.
@@
 \emph{Labels} hold additional information and are represented as boxes containing this information.
-They are usually placed in the \emph{label area} which is a rectangular area designated to hold labels.
+\change{They always have a \emph{port}, a special point on the label's border which will be defined later.} Labels are usually placed in the \emph{label area} which is a rectangular area designated to hold labels.
 It is located next to of the bigger, rectangular \emph{text area}, which contains the document's text and all \emph{sites}, the points or objects that a label's information refers to.
@@
 Will be used in this paper to indicate that there's no loss of progress made from the site to the label, effectively restricting the direction a leader can go at each bend.
-%+Node für Knoten in Polylinien?
 
@@
  \\
- ~\\%There's probably a better way for this...
+ ~\\%There's probably a better way to force multiple newlines...
  \begin{subfigure}[b]{\textwidth}
@@
 Therefore, we can create a routing graph $G=\langle V,E\rangle$ whose Edges $E$ reflect the legal paths a leader can take within $T$.
-For each whitespace character $h_x \in W$ in a given line $l_i \in L$, one vertex is placed in $S(l_{i-1},l_i)$ and $S(l_i,l_{i+1})$ each, located above or below the whitespace character, and with a maximum distance from each of its neighbouring lines.
+For each whitespace character $h \in H$ in a given line $l_i \in L$, one vertex is placed in $S(l_{i-1},l_i)$ and $S(l_i,l_{i+1})$ each, located above or below the whitespace character, and with a maximum distance from each of its neighbouring lines.
 The start and end of each line also recieve a pair of vertices each.

### 2018-02-07 LaTeX/Paper.tex
Paper: Changed parts of section 2.1 (highlighted) and re-wrote section 2.2. Program: No changes.

@@
 These restrictions are implemented as follows: We divided the text $T$ into separate lines $L=\{W,H\}$ which are in turn made up of words $W$ and whitespace $H$.
-The remaining space between lines shall be defined as $S(l_1, l_2)$%improve on this!
-with $l_1,l_2 \in L$ and will have no restrictions regarding leader placement.
+The remaining space between lines is called $S=\{s_1, ...
+s_{n+1}\}$, with $n$ being the number of lines in $L$.
+For each $l_i \in L$,  $s_i$ is the space directly above that line.
+All $s \in S$ are the same height and width, even $s_1$ and $s_{n+1}$, which only have lines on one of their sides.
 For each word $w \in W$, we define $R(w)$  as its bounding rectangle, which marks the space leaders aren't allowed to cross.
@@
 Therefore, we can create a routing graph $G=\langle V,E\rangle$ whose Edges $E$ reflect the legal paths a leader can take within $T$.
-For each whitespace character $h \in H$ in a given line $l_i \in L$, one vertex is placed in $S(l_{i-1},l_i)$ and $S(l_i,l_{i+1})$ each, located above or below the whitespace character, and with a maximum distance from each of its neighbouring lines.
-The start and end of each line also recieve a pair of vertices each.
-Sites will be represented by the insertion of additional vertices in $S(l_i, l_{i-1})$ on a similar height as those next to whitespace of the same line and located directly above the center point of its words $w_j$ bounding rectangle $R(w_j)$.
-Should either $l_{i-1}$ or $l_{i+1}$ not exist, the vertices' vertical placement will be similar to that of the previous or following line.
+\change{For each line $l_i \in L$ and each whitespace character $h \in l_i$, one vertex is placed in $s_i$ and $s_{i+1}$ each, located above or below the whitespace character, and with a maximum distance from each of its neighbouring lines.
+The start and end of each line are also assigned a pair of vertices each.} Sites will be represented by the insertion of additional vertices in $S_i$ on a similar height as those next to whitespace of the same line and located directly above the center point of its words $w_j$ bounding rectangle $R(w_j)$.
 The edges between the vertices are all perfectly horizontal or vertical, and do neither intersect with any bounding rectangle, nor any vertices other than their starting and ending vertex.
@@
 Leaders also aren't allowed to cross one another, as it is hard to discern between intersections and two leaders curving away from each other.
-This is realized by forcing leaders to incorporate as many P-Segments as early as possible while still following all previous restrictions, and forbidding leaders from entering the space to the top left of already successfully routed leader's sites.
+\change{They are also monotonically increasing, and go as far up as possible, limited only by the previous label's placement.} %This is realized by forcing leaders to incorporate as many P-Segments as early as possible while still following all previous restrictions, and forbidding leaders from entering the space to the top left of already successfully routed leader's sites.
 %In our algorithm, the focus was put on keeping the text as readable as possible.
@@
 \section{Description}
+
+The routing part of the algorithm, works through the annotations $a \in A$ in the order they appear in the text, placing each as far up as possible, skipping any annotation that cannot be placed in a reachable position that is also above its leader's source.
+We will also use fixed ports located in the top left corner of each label.
+The routing area will be used to bridge the gap between the highest currently reachable point within the routing graph and the highest possible label placement by connecting the two via an OPO-leader.
+\change{We will call the combination of this connecting leader and the leader generated within the routing graph a \emph{Source-Port-Path} or \emph{SP-Path}.}%Kann man das so sagen?
+
+The routing algorithm itself uses a depth-first-search algorithm, with the Graph $G$ and the set of sites $V_{ann} \subset V$ as input.
+The path within $G$ will be monotonically increasing.
+The Algorithm terminates either when reaching the right text border or if all possible paths failed to reach the text border.
+The SP-Path generated by the algorithm is split into two parts: The path within the graph $(P \subset V)$, which contains all vertices travelled through on the way to the text area's border, and the OPO-Leader that connects to the Label's port.
+If the routing for any given site failed, the SP-path only consists of a single vertex - the site.
+(For an illustration in Pseudocode see Alg.~\ref{alg:DFS}.)
+
+\begin{algorithm}
+
+\DontPrintSemicolon
+\KwData{A single annotation's source and its Graph}
+\KwResult{A List of vertices describing the leader's path}
+
+\SetKwData{G}{Graph}
+\SetKwData{Path}{Path}
+\SetKwData{Src}{Source}
+\SetKwData{Curr}{currentVertex}
+\SetKwData{Old}{oldVertex}
+\SetKwData{Back}{backtracking}
+
+\SetKwFunction{Top}{getTopNeighbourOf}
+\SetKwFunction{Right}{getRightNeighbourOf}
+\SetKwFunction{AddVert}{addVertex}
+\SetKwFunction{RemVert}{RemoveVertex}
+\SetKwFunction{Last}{getLastEntry}
+\SetKwFunction{Below}{isBelow}
+\SetKwFunction{Break}{break}
+
+initialization\;
+\While{\Curr not at right text border}{
+
+ \uIf{$($\G.\Top{\Curr}$\neq null)\land \neg$\Back}{
+  \Path.\AddVert{\Curr}\;
+  \Curr $\gets$ \G.\Top{\Curr}\;\;
+ }
+ \uElseIf{\G.\Right{\Curr}$\neq null$}{
+ \Path.\AddVert{\Curr}\;
+  \Curr $\gets$ \G.\Top{\Curr}\;
+  \Back $\gets$ False\;
+ }
+ \Else{
+  \Back $\gets$ True\;
+  \Repeat{\Curr's Position is below \Old or \Path is Empty}{
+   \Old $\gets$ \Curr\;
+   \Curr $\gets$ \Path.\Last{}\;
+   \Path.\RemVert{\Curr}\;
+  }
+  \;
+  \If(//No path found){\Curr not below \Old}{\Break}
+ }
+}
+\caption{The Depth-First-Search algorithm used in the program.}
+\label{alg:DFS}
+\end{algorithm}
+
+\section{Analysis and Proof}
 
@@
 
+%TODO: Re-Write!
+
 First, we create the graph that will be used for the routing part of the algorithm.
@@
 Afterwards, we create new vertices above and below the whitespace characters $h_k, h_{k+1}$ preceding and succeeding $w_i$.
-These vertices are located in the space $S(l_{j-1},l_j), S(l_j,l_{j+1})$ between the word's line and its surrounding lines $l_{j-1}$ and $l_{j+1}$, as described in the previous section.
-Whenever an annotation is encountered, a single vertex is added above the last non-annotation word in $S(l_{j-1},l_j)$ at $pos(h_k)+le(w_i)/2$.
+These vertices are located in the space $s_j, s_{j+1}$ between the word's line and its surrounding lines $l_{j-1}$ and $l_{j+1}$, as described in the previous section.
+Whenever an annotation is encountered, a single vertex is added above the last non-annotation word in $s_j$ at $pos(h_k)+le(w_i)/2$.
 For a representation of this process in pseudocode, see Alg.~\ref{alg:GraphCreation}.
@@
 (For a representation in Pseudocode, see Alg.~\ref{alg:GraphCreation})
-
-For the routing part of the algorithm, we work through the annotations $a \in A$ in the order they appear in the text, placing each as far up as possible, skipping any annotation that cannot be placed in a reachable position above its leader's source.
-We will also use fixed ports located in the top left corner of each label.
-%This allows us to place annotations by only knowing their port location, which can be inferred from their leader's ending point.
-The routing area, which is located between the text and label areas, will be used to bridge the gap between the highest currently reachable point within the routing graph and the highest possible label placement by connecting the two via an OPO-leader.%Kann man das so sagen?
-
-The routing algorithm itself uses only the Graph $G$ and the set of sites $V_{ann} \subset V$ which will be routed using a depth-first search algorithm, using only vertices located above or to the right of the current vertex $v_{curr}$, prioritizing moving up whenever possible.
-It will terminate upon reaching the right text border or if all possible paths failed to reach the text border.
-At this point, it will return a set with routing information for each vertex, which contains a Path $P=\{v_1, \cdots, v_n\}$, leading from the site to the text area's border, along with information on how to draw the OPO-Segment that connects to the Label's port.
-If the routing for any given site failed, the path consists of a single vertex - the site.
-(For an illustration in Pseudocode see Alg.~\ref{alg:DFS}.)
-
-\begin{algorithm}
-
-\DontPrintSemicolon
-\KwData{A single annotation's source and its Graph}
-\KwResult{A List of vertices describing the leader's path}
-
-\SetKwData{G}{Graph}
-\SetKwData{Path}{Path}
-\SetKwData{Src}{Source}
-\SetKwData{Curr}{currentVertex}
-\SetKwData{Old}{oldVertex}
-\SetKwData{Back}{backtracking}
-
-\SetKwFunction{Top}{getTopNeighbourOf}
-\SetKwFunction{Right}{getRightNeighbourOf}
-\SetKwFunction{AddVert}{addVertex}
-\SetKwFunction{RemVert}{RemoveVertex}
-\SetKwFunction{Last}{getLastEntry}
-\SetKwFunction{Below}{isBelow}
-\SetKwFunction{Break}{break}
-
-initialization\;
-\While{\Curr not at right text border}{
-
- \uIf{$($\G.\Top{\Curr}$\neq null)\land \neg$\Back}{
-  \Path.\AddVert{\Curr}\;
-  \Curr $\gets$ \G.\Top{\Curr}\;\;
- }
- \uElseIf{\G.\Right{\Curr}$\neq null$}{
- \Path.\AddVert{\Curr}\;
-  \Curr $\gets$ \G.\Top{\Curr}\;
-  \Back $\gets$ False\;
- }
- \Else{
-  \Back $\gets$ True\;
-  \Repeat{\Curr's Position is below \Old or \Path is Empty}{
-   \Old $\gets$ \Curr\;
-   \Curr $\gets$ \Path.\Last{}\;
-   \Path.\RemVert{\Curr}\;
-  }
-  \;
-  \If(//No path found){\Curr not below \Old}{\Break}
- }
-}
-\caption{The Depth-First-Search algorithm used in the program.}
-\label{alg:DFS}
-\end{algorithm}
 

### 2018-02-09 LaTeX/Paper.tex
Paper: Finished Re-Write of 2.2. Program: No changes.

@@
 
-The routing part of the algorithm, works through the annotations $a \in A$ in the order they appear in the text, placing each as far up as possible, skipping any annotation that cannot be placed in a reachable position that is also above its leader's source.
-We will also use fixed ports located in the top left corner of each label.
+The routing part of the algorithm works through the annotations $a \in A$ in the order they appear in the text, placing each as far up as possible, skipping any annotation that cannot be placed in a reachable position that is also above its leader's source.
+It will use fixed ports located in the top left corner of each label.
 The routing area will be used to bridge the gap between the highest currently reachable point within the routing graph and the highest possible label placement by connecting the two via an OPO-leader.
@@
 
-The routing algorithm itself uses a depth-first-search algorithm, with the Graph $G$ and the set of sites $V_{ann} \subset V$ as input.
-The path within $G$ will be monotonically increasing.
-The Algorithm terminates either when reaching the right text border or if all possible paths failed to reach the text border.
-The SP-Path generated by the algorithm is split into two parts: The path within the graph $(P \subset V)$, which contains all vertices travelled through on the way to the text area's border, and the OPO-Leader that connects to the Label's port.
-If the routing for any given site failed, the SP-path only consists of a single vertex - the site.
-(For an illustration in Pseudocode see Alg.~\ref{alg:DFS}.)
+\change{The routing algorithm itself uses a depth-first-search algorithm, with the Graph $G$ and the set of sites $V_{ann} \subset V$ as input.
+The Algorithm terminates either when reaching the right text border or if all possible paths failed to reach the text border, and the resulting SP-Path will be monotonically increasing.
+This path is split into two parts: The path within the graph $(P \subset V)$, leading from the source to the text area's border, and the OPO-Leader that connects the text area's border with the Label's port.
+The former is realized as a list of vertices $V_{Path} \subset V$ the leader travels through, whereas the latter is determined by the rightmost vertex of $V_{Path}$, the Port, and the location of the P-Segment.
+If the routing for any given site failed, the SP-path only consists of a single vertex - the site.} (For an illustration in Pseudocode see Alg.~\ref{alg:DFS}.)
 

### 2018-02-11 LaTeX/Paper.tex
Paper: Changed 2.3 to a short list of planned topics, previous content is temporarily hidden. Program: No changes.

@@
 
-The routing part of the algorithm works through the annotations $a \in A$ in the order they appear in the text, placing each as far up as possible, skipping any annotation that cannot be placed in a reachable position that is also above its leader's source.
+The routing algorithm works through the annotations $a \in A$ in the order they appear in the text, placing each as far up as possible, skipping any annotation that cannot be placed above its leader's source or is impossible to route.
 It will use fixed ports located in the top left corner of each label.
 The routing area will be used to bridge the gap between the highest currently reachable point within the routing graph and the highest possible label placement by connecting the two via an OPO-leader.
-\change{We will call the combination of this connecting leader and the leader generated within the routing graph a \emph{Source-Port-Path} or \emph{SP-Path}.}%Kann man das so sagen?
+\change{We will call the combination of this connecting leader and the path generated within the routing graph a \emph{Source-Port-Path} or \emph{SP-path}.}%Kann man das so sagen?
 
 \change{The routing algorithm itself uses a depth-first-search algorithm, with the Graph $G$ and the set of sites $V_{ann} \subset V$ as input.
-The Algorithm terminates either when reaching the right text border or if all possible paths failed to reach the text border, and the resulting SP-Path will be monotonically increasing.
+The Algorithm terminates either when reaching the right text border or if all possible paths failed to reach the text border, and the resulting SP-path will be monotonically increasing.
 This path is split into two parts: The path within the graph $(P \subset V)$, leading from the source to the text area's border, and the OPO-Leader that connects the text area's border with the Label's port.
-The former is realized as a list of vertices $V_{Path} \subset V$ the leader travels through, whereas the latter is determined by the rightmost vertex of $V_{Path}$, the Port, and the location of the P-Segment.
+The former is implemented as a list of vertices $V_{Path} \subset V$ the leader travels through, whereas the latter is determined by the rightmost vertex of $V_{Path}$, the Port, and the location of the P-Segment.
 If the routing for any given site failed, the SP-path only consists of a single vertex - the site.} (For an illustration in Pseudocode see Alg.~\ref{alg:DFS}.)
+
+\change{Each SP-path's P-Segment in the Routing Area is only placed after all SP-Paths have been generated, and will be placed with equal spacing between the borders of the Routing Area and each P-segment, to ensure readability of the resulting leaders.}
 
@@
 
-\begin{algorithm}%{r}{0.35\textwidth}
-\DontPrintSemicolon
-\SetKwData{Words}{words}
-\SetKwData{PrevW}{previousWord}
-\SetKwData{G}{Graph}
-\SetKwData{W}{w} \SetKwData{V}{v}\SetKwData{Va}{v1} \SetKwData{Vb}{v2} \SetKwData{Vc}{v3} \SetKwData{Vd}{v4}
-\SetKwData{Up}{UpperVerticesList}
-\SetKwData{Low}{LowerVerticesList}
+In this section we will analyze the following:
 
-\SetKwFunction{setAnn}{setAnnotation}
-\SetKwFunction{NewVert}{new Vertex}
-\SetKwFunction{AddVert}{addVertex}
-\SetKwFunction{AddAll}{addAll}
-\SetKwFunction{CEdge}{createEdgeBetween}
-\SetKwFunction{PosConn}{connectBasedOnPosition}
-\SetKwFunction{ClrList}{emptyList}
-\SetKwFunction{getC}{getCenter}
-\SetKwFunction{tLeft}{getTopLeft}
-\SetKwFunction{tRight}{getTopRight}
-\SetKwFunction{bLeft}{getBottomLeft}
-\SetKwFunction{bRight}{getBottomRight}
-\SetKwFunction{Newline}{startNewLine}
+\begin{itemize}
+ \item Computation time of graph-creation
+ \item Computation time of the routing algorithm for single leaders
+ \item Overall computation time of the routing algorithm
+\end{itemize}
+~\\
+Furthermore, we will prove that:
+\begin{itemize}
+\item The algorithm always returns a shortest path
+\item The algorithm generates only crossing-free routings
+\item The algorithm places as many annotations as possible
+\item The algorithm keeps the annnotations in reading order
+\end{itemize}
 
-\KwData{A text with annotations,stored as a String-Array}
-\KwResult{A Graph (as described above)}
-
- initialization\;
-
- \ForEach{\W in \Words}{
-  \eIf{\W is annotation}{
-   \V$\gets$\NewVert{\PrevW.\getC{}}\;
-   \V.\setAnn{\W}\;
-   \G.\AddVert{\V}\;
-   \Up.\AddVert{\V}\;
-  }{
-   \If{\W is too big for the line}{
-    \Newline{}\;
-    \PosConn{\Up}\;
-    \Up$\gets$\Low\;
-    \ClrList{\Low}\;
-   }
-   \Va $\gets$\NewVert{\W.\tLeft{}}\;
-   \Vb $\gets$\NewVert{\W.\tRight{}}\;
-   \Vc $\gets$\NewVert{\W.\bLeft{}}\;
-   \Vd $\gets$\NewVert{\W.\bRight{}}\;
-   \;
-   \G.\AddAll{\Va,\Vb,\Vc,\Vd}\;
-   \Up.\AddAll{\Va,\Vb}\;
-   \Low.\AddAll{\Vc,\Vd}\;
-   \G.\CEdge{\Va,\Vc}\;
-   \G.\CEdge{\Vb,\Vd}\;
-  }
- } 
-\caption{Representation of the Graph-creation algorithm in pseudocode}
-\label{alg:GraphCreation}
-\end{algorithm}
+%\begin{algorithm}%{r}{0.35\textwidth}
+%\DontPrintSemicolon
+%\SetKwData{Words}{words}
+%\SetKwData{PrevW}{previousWord}
+%\SetKwData{G}{Graph}
+%\SetKwData{W}{w} \SetKwData{V}{v}\SetKwData{Va}{v1} \SetKwData{Vb}{v2} \SetKwData{Vc}{v3} \SetKwData{Vd}{v4}
+%\SetKwData{Up}{UpperVerticesList}
+%\SetKwData{Low}{LowerVerticesList}
+%
+%\SetKwFunction{setAnn}{setAnnotation}
+%\SetKwFunction{NewVert}{new Vertex}
+%\SetKwFunction{AddVert}{addVertex}
+%\SetKwFunction{AddAll}{addAll}
+%\SetKwFunction{CEdge}{createEdgeBetween}
+%\SetKwFunction{PosConn}{connectBasedOnPosition}
+%\SetKwFunction{ClrList}{emptyList}
+%\SetKwFunction{getC}{getCenter}
+%\SetKwFunction{tLeft}{getTopLeft}
+%\SetKwFunction{tRight}{getTopRight}
+%\SetKwFunction{bLeft}{getBottomLeft}
+%\SetKwFunction{bRight}{getBottomRight}
+%\SetKwFunction{Newline}{startNewLine}
+%
+%\KwData{A text with annotations,stored as a String-Array}
+%\KwResult{A Graph (as described above)}
+%
+% initialization\;
+% 
+% \ForEach{\W in \Words}{
+%  \eIf{\W is annotation}{
+%   \V$\gets$\NewVert{\PrevW.\getC{}}\;
+%   \V.\setAnn{\W}\;
+%   \G.\AddVert{\V}\;
+%   \Up.\AddVert{\V}\;
+%  }{
+%   \If{\W is too big for the line}{
+%    \Newline{}\;
+%    \PosConn{\Up}\;
+%    \Up$\gets$\Low\;
+%    \ClrList{\Low}\;
+%   }
+%   \Va $\gets$\NewVert{\W.\tLeft{}}\;
+%   \Vb $\gets$\NewVert{\W.\tRight{}}\;
+%   \Vc $\gets$\NewVert{\W.\bLeft{}}\;
+%   \Vd $\gets$\NewVert{\W.\bRight{}}\;
+%   \;
+%   \G.\AddAll{\Va,\Vb,\Vc,\Vd}\;
+%   \Up.\AddAll{\Va,\Vb}\;
+%   \Low.\AddAll{\Vc,\Vd}\;
+%   \G.\CEdge{\Va,\Vc}\;
+%   \G.\CEdge{\Vb,\Vd}\;
+%  }
+% } 
+%\caption{Representation of the Graph-creation algorithm in pseudocode}
+%\label{alg:GraphCreation}
+%\end{algorithm}
 
@@
 
-First, we create the graph that will be used for the routing part of the algorithm.
+%First, we create the graph that will be used for the routing part of the algorithm.
 To achieve this, we go through the annotated text word by word in reading order, and measure the length $le(w_i)$ of each word $w_i$ that is not part of an annotation to determine its placement in the text.
@@
 Whenever an annotation is encountered, a single vertex is added above the last non-annotation word in $s_j$ at $pos(h_k)+le(w_i)/2$.
-For a representation of this process in pseudocode, see Alg.~\ref{alg:GraphCreation}.
+%For a representation of this process in pseudocode, see Alg.~\ref{alg:GraphCreation}.
 %To create leaders that exclusively use the space not taken up by a word's bounding rectangle, we decided to use a graph similar to the one Loose \cite{Loose2015} used.

### 2018-02-20 LaTeX/Paper.tex
Paper: Re-wrote definitions for Polylines and Monotonicity. Program: No changes.

@@
 \usepackage{subcaption}
-%\usepackage[]{luatodonotes}
-%\usepackage{wrapfig}
 
@@
 
-\section{Terminology and Fundamentals} %Evtl in Introduction eingliedern?
+\section{Terminology and Fundamentals} 
 Boundary Labeling (or equivalent concepts) can be applied to a space with different geometry or more dimensions, but this thesis will only concern itself with two-dimensonal, Euclidean space.
@@
 
-\begin{figure}%{r}{0.35\textwidth}
+\begin{figure}
  \captionsetup{justification=centering, margin=0.75cm}
@@
  \emph{Depth-first search} is a searching algorithm on a graph G that starts at a given vertex $v \in V$ and explores the graph by traversing its edges as far as possible before backtracking, and continues to do so until a pre-defined goal is met.
-\change{A function is called \emph{monotone} if it is either non-decreasing or non-increasing, meaning that for each $x<y$, $f(x) \leq f(y)$ for \emph{monotonically increasing functions}, or $f(x) \geq f(y)$ for \emph{monotonically decreasing} functions.}
-% A search algorithm is \emph{monotonous} if, for each vertex $v$ and each successor $v'$ generated by the algorithm, the estimated cost of reaching the goal from $v$ is not greater than the cost of reaching $v'$ plus the estimated cost of reaching the goal from $v'$.
+%A function is called \emph{monotone} if it is either non-decreasing or non-increasing, meaning that for each $x<y$, $f(x) \leq f(y)$ for \emph{monotonically increasing functions}, or $f(x) \geq f(y)$ for \emph{monotonically decreasing} functions.
 
-\emph{Polylines} are a connected series of \emph{line segments}. Line segments are straight lines that contain each point between their starting and end point. 
+\change{A \emph{Polyline} is a curve defined by a sequence of points $O=p_1, p_2, ...,p_n$, connecting each point to its successor with straight lines.
+Each point $p \in O$ cannot be equal to any other point in $O$.
+We will call a polyline \emph{monotone}, if it satisfies certain conditions. For \emph{monotonically increasing} polylines, each two vertices must satisfy the following: \\
+$\forall p_i, p_j \in O: i<j \iff (X(p_i)\leq X(p_j) \land Y(p_i)\leq Y(p_j)\land p_i \neq p_j)$, where $X(.)$ returns a point's X-coordinate, and $Y(.)$ returns its Y-coordinate.
+For \emph{monotonically decreasing} polylines, the inverse must be true - each point must have at least one coordinate with a value smaller than its predecessor.}
 \emph{Labels} hold additional information and are represented as boxes containing this information.
-\change{They always have a \emph{port}, a special point on the label's border which will be defined later.} Labels are usually placed in the \emph{label area} which is a rectangular area designated to hold labels.
+They always have a \emph{port}, a special point on the label's border which will be defined later.
+Labels are usually placed in the \emph{label area} which is a rectangular area designated to hold labels.
 It is located next to of the bigger, rectangular \emph{text area}, which contains the document's text and all \emph{sites}, the points or objects that a label's information refers to.
@@
 The location where a leader connects to the label is called the \emph{port}. It can be restricted to pre-defined positions. %TODO: Re-insert monotonicity
-
-% \textbf{Monotonous/Monotonicity:} Describes a steady progression towards a direction or goal.
-Will be used in this paper to indicate that there's no loss of progress made from the site to the label, effectively restricting the direction a leader can go at each bend.
 
@@
 Therefore, we can create a routing graph $G=\langle V,E\rangle$ whose Edges $E$ reflect the legal paths a leader can take within $T$.
-\change{For each line $l_i \in L$ and each whitespace character $h \in l_i$, one vertex is placed in $s_i$ and $s_{i+1}$ each, located above or below the whitespace character, and with a maximum distance from each of its neighbouring lines.
-The start and end of each line are also assigned a pair of vertices each.} Sites will be represented by the insertion of additional vertices in $S_i$ on a similar height as those next to whitespace of the same line and located directly above the center point of its words $w_j$ bounding rectangle $R(w_j)$.
+For each line $l_i \in L$ and each whitespace character $h \in l_i$, one vertex is placed in $s_i$ and $s_{i+1}$ each, located above or below the whitespace character, and with a maximum distance from each of its neighbouring lines.
+The start and end of each line are also assigned a pair of vertices each.
+Sites will be represented by the insertion of additional vertices in $S_i$ on a similar height as those next to whitespace of the same line and located directly above the center point of its words $w_j$ bounding rectangle $R(w_j)$.
 The edges between the vertices are all perfectly horizontal or vertical, and do neither intersect with any bounding rectangle, nor any vertices other than their starting and ending vertex.
@@
 Leaders also aren't allowed to cross one another, as it is hard to discern between intersections and two leaders curving away from each other.
-\change{They are also monotonically increasing, and go as far up as possible, limited only by the previous label's placement.} %This is realized by forcing leaders to incorporate as many P-Segments as early as possible while still following all previous restrictions, and forbidding leaders from entering the space to the top left of already successfully routed leader's sites.
+They are also monotonically increasing, and go as far up as possible, limited only by the previous label's placement.
+%This is realized by forcing leaders to incorporate as many P-Segments as early as possible while still following all previous restrictions, and forbidding leaders from entering the space to the top left of already successfully routed leader's sites.
 %In our algorithm, the focus was put on keeping the text as readable as possible.
@@
 The routing area will be used to bridge the gap between the highest currently reachable point within the routing graph and the highest possible label placement by connecting the two via an OPO-leader.
-\change{We will call the combination of this connecting leader and the path generated within the routing graph a \emph{Source-Port-Path} or \emph{SP-path}.}%Kann man das so sagen?
+We will call the combination of this connecting leader and the path generated within the routing graph a \emph{Source-Port-Path} or \emph{SP-path}.%Kann man das so sagen?
 
-\change{The routing algorithm itself uses a depth-first-search algorithm, with the Graph $G$ and the set of sites $V_{ann} \subset V$ as input.
+The routing algorithm itself uses a depth-first-search algorithm, with the Graph $G$ and the set of sites $V_{ann} \subset V$ as input.
 The Algorithm terminates either when reaching the right text border or if all possible paths failed to reach the text border, and the resulting SP-path will be monotonically increasing.
@@
 The former is implemented as a list of vertices $V_{Path} \subset V$ the leader travels through, whereas the latter is determined by the rightmost vertex of $V_{Path}$, the Port, and the location of the P-Segment.
-If the routing for any given site failed, the SP-path only consists of a single vertex - the site.} (For an illustration in Pseudocode see Alg.~\ref{alg:DFS}.)
+\change{If the routing for any given site failed, the algorithm returns $\bot$.} (For an illustration in Pseudocode see Alg.~\ref{alg:DFS}.)
 
-\change{Each SP-path's P-Segment in the Routing Area is only placed after all SP-Paths have been generated, and will be placed with equal spacing between the borders of the Routing Area and each P-segment, to ensure readability of the resulting leaders.}
+Each SP-path's P-Segment in the Routing Area is only placed after all SP-Paths have been generated, and will be placed with equal spacing between the borders of the Routing Area and each P-segment, to ensure readability of the resulting leaders.
 
@@
 
-\section{Data generation} %\change{(Bin mit dem Titel unzufrieden, aber hier soll beschrieben werden, wie ich die einzelnen Testfälle erstellt habe.)}
+\section{Data generation} %(Bin mit dem Titel unzufrieden, aber hier soll beschrieben werden, wie ich die einzelnen Testfälle erstellt habe.)
 

### 2018-02-21 LaTeX/Paper.tex
Paper: Moved and re-wrote Definition of SP-Paths, cleared up unused comments and files. Program: no changes.

@@
 The resulting graph looks similar to Fig.~\ref{fig:rgraph}.
+\change{We wil call the vertices closest to the text area's border to the routing area \emph{Border Vertices} $B \subset V$.
+They serve as a goal for the depth-first search, and are located at the end of each line in our implementation.}
 
@@
 Leaders also aren't allowed to cross one another, as it is hard to discern between intersections and two leaders curving away from each other.
-They are also monotonically increasing, and go as far up as possible, limited only by the previous label's placement.
-%This is realized by forcing leaders to incorporate as many P-Segments as early as possible while still following all previous restrictions, and forbidding leaders from entering the space to the top left of already successfully routed leader's sites.
-%In our algorithm, the focus was put on keeping the text as readable as possible.
-This means that leaders aren't allowed to pass through words - they can only use the space between words and in between lines.
-%Additionally, we wanted to maximize the usage of the label area's space, even if we have to move labels further from their sites than needed. 
-%Finally, we wanted to keep the leader's length low for any given combination of labels and site placements, so every segment of a leader has to decrease the remaining distance between the site and its label's port.
+Furthermore, they are monotonically increasing, and go as far up as possible, limited only by the previous label's placement.
 
-%\begin{figure}
-% \captionsetup{justification=centering, margin=0.75cm}
-% \centering
-%  \makebox[\textwidth][c]{\includegraphics[scale=0.9]{FromTextToGraph.png}}
-%  \caption{Visualization of the space reserved for the text, and the resulting graph.}
-% \label{fig:ttg}
-%\end{figure}
-
-%These properties were modelled by creating a rectangle around each word that no leaders may pass through.
-If this is done for all words in a paragraph, the remaining available space for routing resembles a grid, like in Fig.\ref{fig:ttg}a.
-
-%Since the layout encourages the usage of P- and O-segments, we decided to represent the possible routes by creating a graph which has its vertices located whenever multiple directions are available - in between lines, above or below whitespace characters.
-The graph's edges represent the different possible segments of a leader and connect each vertex with the closest vertices directly above, below or to the sides, given that no text is blocking the connection.
-The resulting graph looks similar to the one in Fig.~\ref{fig:ttg}b.
-
+\change{The Routing Area will be used to connect each site's path with its source, using OPO-Leaders.
+Combined with the path $P=\{s, v_1, v_2, ..., v_b\}; P \subset V$ leading from the source $s \in S$ to the graph's border $(v_b \in B)$ it creates an unbroken connection between a source and its port, which shall be called a \emph{Source-Port-Path}, or \emph{SP-Path}.}
 %For ease of visualisation, we limited the number of leaders per edge to one, and forbid the leaders to cross one another, as this could be mistaken for two leaders touching each other at the corners.
@@
 
-%This was implemented by enclosing each word in a bounding rectangle that leaders aren't allowed to pass through.
-%The leaders should also be kept as short as possible, so we restricted them to always be moving toward the label, even if the direct route is unavailable.%Wie drücke ich das als Formel aus?
-%We also wanted to use the space available in the labeling area as efficiently as possible, so labels are placed as far up as possible in order to maximize the space available to future labels at the cost of increased leader length.
-
 \section{Description}
@@
 It will use fixed ports located in the top left corner of each label.
-The routing area will be used to bridge the gap between the highest currently reachable point within the routing graph and the highest possible label placement by connecting the two via an OPO-leader.
-We will call the combination of this connecting leader and the path generated within the routing graph a \emph{Source-Port-Path} or \emph{SP-path}.%Kann man das so sagen?
 

### 2018-02-23 LaTeX/Paper.tex
Paper: Wrote computation time analysis for graph creation algorithm. Re-introduced and edited graph-creation pseudocode. Program: No changes.

@@
 \section{Problem Specification}
+\label{sec:ProbSpec}
 We limited the leaders to use exclusively O- and P-Segments, and banned them from passing through words.
@@
 \begin{itemize}
-\item The algorithm always returns a shortest path
 \item The algorithm generates only crossing-free routings
-\item The algorithm places as many annotations as possible
-\item The algorithm keeps the annnotations in reading order
+\item The algorithm always returns a highest (define!) legal path
+
+NOTE: The algorithm keeps the annnotations in reading order
 \end{itemize}
 
-%\begin{algorithm}%{r}{0.35\textwidth}
-%\DontPrintSemicolon
-%\SetKwData{Words}{words}
-%\SetKwData{PrevW}{previousWord}
-%\SetKwData{G}{Graph}
-%\SetKwData{W}{w} \SetKwData{V}{v}\SetKwData{Va}{v1} \SetKwData{Vb}{v2} \SetKwData{Vc}{v3} \SetKwData{Vd}{v4}
-%\SetKwData{Up}{UpperVerticesList}
-%\SetKwData{Low}{LowerVerticesList}
-%
-%\SetKwFunction{setAnn}{setAnnotation}
-%\SetKwFunction{NewVert}{new Vertex}
-%\SetKwFunction{AddVert}{addVertex}
-%\SetKwFunction{AddAll}{addAll}
-%\SetKwFunction{CEdge}{createEdgeBetween}
-%\SetKwFunction{PosConn}{connectBasedOnPosition}
-%\SetKwFunction{ClrList}{emptyList}
-%\SetKwFunction{getC}{getCenter}
-%\SetKwFunction{tLeft}{getTopLeft}
-%\SetKwFunction{tRight}{getTopRight}
-%\SetKwFunction{bLeft}{getBottomLeft}
-%\SetKwFunction{bRight}{getBottomRight}
-%\SetKwFunction{Newline}{startNewLine}
-%
-%\KwData{A text with annotations,stored as a String-Array}
-%\KwResult{A Graph (as described above)}
-%
-% initialization\;
-% 
-% \ForEach{\W in \Words}{
-%  \eIf{\W is annotation}{
-%   \V$\gets$\NewVert{\PrevW.\getC{}}\;
-%   \V.\setAnn{\W}\;
-%   \G.\AddVert{\V}\;
-%   \Up.\AddVert{\V}\;
-%  }{
-%   \If{\W is too big for the line}{
-%    \Newline{}\;
-%    \PosConn{\Up}\;
-%    \Up$\gets$\Low\;
-%    \ClrList{\Low}\;
-%   }
-%   \Va $\gets$\NewVert{\W.\tLeft{}}\;
-%   \Vb $\gets$\NewVert{\W.\tRight{}}\;
-%   \Vc $\gets$\NewVert{\W.\bLeft{}}\;
-%   \Vd $\gets$\NewVert{\W.\bRight{}}\;
-%   \;
-%   \G.\AddAll{\Va,\Vb,\Vc,\Vd}\;
-%   \Up.\AddAll{\Va,\Vb}\;
-%   \Low.\AddAll{\Vc,\Vd}\;
-%   \G.\CEdge{\Va,\Vc}\;
-%   \G.\CEdge{\Vb,\Vd}\;
-%  }
-% } 
-%\caption{Representation of the Graph-creation algorithm in pseudocode}
-%\label{alg:GraphCreation}
-%\end{algorithm}
+To determine the program's computation time, we will first take a look at the creation of the graph.
+As discussed in section~\ref{sec:ProbSpec}, we will be placing a pair of vertices per whitespace character, as well as one additional pair at both the start and end of each line, which amounts to $2\times|H|+4\times|L|$ vertices total.
+Together with the sites $I$, which are also represented as vertices, this results in the total number of vertices $|V|=|I|+2\times|H|+4\times|L|$.
+Additionally, each vertex in a pair is connected to its counterpart, which creates $(|V|-|I|)/2$ edges, representing all possible P-Segments.
+Finally, each edge placed in a space $s \in S$ above or below a line will be connected to each other, which is solved by connecting each vertex to the next unconnected vertex remaining in $s$.
+This leads to the creation of $|s|-1$ edges representing possible P-segments, which leads to a total of $|V|-|S|$ additional edges, making the total number of edges $|E|=(3\times|V|-2\times|S|-|I|)/2$.%Todo: look up fractions.
+If we assume that the creation of edges and vertices both take a similar amount of time, we are looking at roughly $5/2 \times |V|$ operations, which means that the algorithm scales linearly with the amount of vertices placed, which in turn is directly proportional to the number of words $|W|$ in the text.
+For a representation of the graph-generation algorithm in pseudocode, see Alg.~\ref{alg:GraphCreation}.
+
+\begin{algorithm}
+\DontPrintSemicolon
+\SetKwData{Words}{words}
+\SetKwData{PrevW}{previousWord}
+\SetKwData{G}{Graph}
+\SetKwData{W}{w} \SetKwData{V}{v}\SetKwData{Va}{v1} \SetKwData{Vb}{v2}
+\SetKwData{Up}{UpperVerticesList}
+\SetKwData{Low}{LowerVerticesList}
+
+\SetKwFunction{setAnn}{setAnnotation}
+\SetKwFunction{NewVert}{new Vertex}
+\SetKwFunction{AddVert}{addVertex}
+\SetKwFunction{AddAll}{addAll}
+\SetKwFunction{CEdge}{createEdgeBetween}
+\SetKwFunction{PosConn}{connectBasedOnPosition}
+\SetKwFunction{ClrList}{emptyList}
+\SetKwFunction{getC}{getCenter}
+\SetKwFunction{tLeft}{getTopLeft}
+\SetKwFunction{tRight}{getTopRight}
+\SetKwFunction{bLeft}{getBottomLeft}
+\SetKwFunction{bRight}{getBottomRight}
+\SetKwFunction{Newline}{startNewLine}
+
+\KwData{A text with annotations,stored as a String-Array}
+\KwResult{A Graph (as described above)}
+
+ initialization\;
+
+ \ForEach{\W in \Words}{
+  \eIf{\W is annotation}{
+   \V$\gets$\NewVert{\PrevW.\getC{}}\;
+   \V.\setAnn{\W}\;
+   \G.\AddVert{\V}\;
+   \Up.\AddVert{\V}\;
+  }{
+   \If{\W is too big for the line}{
+    //Create last pair of vertices in current line\;
+    \Va $\gets$\NewVert{\PrevW.\tRight{}}\;
+    \Vb $\gets$\NewVert{\PrevW.\bRight{}}\;
+    \G.\AddAll{\Va,\Vb}\;
+    \Up.\AddVert{\Va}\;
+    \Low.\AddVert{\Vb}\;
+    \G.\CEdge{\Va,\Vb}\;
+    \;
+    //Start new line\;
+    \Newline{}\;
+    \PosConn{\Up}\;
+    \Up$\gets$\Low\;
+    \ClrList{\Low}\;
+   }
+   \Va $\gets$\NewVert{\W.\tLeft{}}\;
+   \Vb $\gets$\NewVert{\W.\bLeft{}}\;
+   \;
+   \G.\AddAll{\Va,\Vb}\;
+   \Up.\AddVert{\Va}\;
+   \Low.\AddVert{\Vb}\;
+   \G.\CEdge{\Va,\Vb}\;
+  }
+ } 
+\caption{Representation of the Graph-creation algorithm in pseudocode}
+\label{alg:GraphCreation}
+\end{algorithm}
 

### 2018-02-24 LaTeX/Paper.tex
Paper: Added section about routing algorithm's computation time. (TODO: Inspect how leaders influence each other's search space) Program: Changed routing algorithm to properly use DFS - it now remembers nodes that were backtracked through.

@@
 
+Next, we shall look at the routing algorithm - since we use depth-first search, which is a well-known algorithm, the worst-case computation time is known to be $|V|+|E|$.
+To reach this time, we'd have to try and use every single edge in the graph, reaching each of the graph's vertices in the process.
+However, as our algorithm can't use any edges leading away from its target, our search space is reduced to only the vertices either in lower-numbered $s \in S$ (``above'') or those in the same $s$, but located between the source and the border vertices.
+If we also take into account that leaders aren't allowed to cross one another, this further limits our search space, removing all vertices above the previous leader's vertices.
+
 \begin{algorithm}

### 2018-02-24 Eclipse/Labeling/src/routingapp/GreedyTopRouting.java

@@ -35,6 +35,7 @@ public class GreedyTopRouting implements Routing{
 		GraphTuple currentNode=source;
 
 		ArrayList<GraphTuple> pathNodes=new ArrayList<GraphTuple>();
+		ArrayList<GraphTuple> visitedNodes=new ArrayList<GraphTuple>(); //Stores already visited nodes after backtracking, so we don't run into the same dead end twice.
 		pathNodes.add(index, currentNode);
 		index++;
 
@@ -61,11 +62,11 @@ public class GreedyTopRouting implements Routing{
 
 						if(temp.equals(currentNode)) temp=graph.getEdgeTarget(edge);
 
-						if((temp.getX()>currentNode.getX()))
+						if((!visitedNodes.contains(temp))&&(temp.getX()>currentNode.getX()))
 						{
 							horizontalCandidate=temp;
 						}
-						else if((!backtrack)&&(temp.getY()<currentNode.getY())&&(temp.getY()>nextAnnotationPos))
+						else if((!backtrack)&&(!visitedNodes.contains(temp))&&(temp.getY()<currentNode.getY())&&(temp.getY()>nextAnnotationPos))
 						{
 							if((currentNode.getX()>lastAnnotatedWord.getX())||(currentNode.getY()>lastAnnotatedWord.getY()))
 							{
@@ -96,6 +97,7 @@ public class GreedyTopRouting implements Routing{
 						while((tempNode.getY()==currentNode.getY())&&(currentNode!=pathNodes.get(0)))	
 						{
 							index--;
+							visitedNodes.add(pathNodes.get(index));
 							pathNodes.remove(index);
 							currentNode=pathNodes.get(index-1);
 						}
@@ -118,6 +120,7 @@ public class GreedyTopRouting implements Routing{
 							GraphTuple newTuple=iter.next();
 							DefaultWeightedEdge temp=graph.getEdge(oldTuple, newTuple);
 							if(temp!=null)graph.setEdgeWeight(temp, graph.getEdgeWeight(temp)-1);
+							
 							oldTuple=newTuple;
 						}
 					}
@@ -125,6 +128,7 @@ public class GreedyTopRouting implements Routing{
 			}//end while
 		}
 
+		visitedNodes.clear();
 		return new GraphWalk<GraphTuple,DefaultWeightedEdge>(graph, pathNodes,pathNodes.size());
 	}
 

### 2018-02-25 LaTeX/Paper.tex
Paper: Completed Proof section of the paper (first version). Re-wrote parts of previous sections to include definitions and concepts neccesessary for this. Program: No changes.

@@
 The resulting graph looks similar to Fig.~\ref{fig:rgraph}.
-\change{We wil call the vertices closest to the text area's border to the routing area \emph{Border Vertices} $B \subset V$.
+\change{In the following, we define vertex $v_a \in s_i$ to be \emph{above} another vertex $v_b \in s_j$, if $i<j$.
+If $v_a$ and $v_b$ are adjacent to each other, $v_a$ is \emph{directly above} $v_b$.
+We will also define a vertex $v_b$ to be \emph{between} two vertices $v_a$ and $v_c$, if $v_a, v_b, v_c \in s$ $(s \in S)$ and it is impossible to create a path $P \subseteq s$ using only vertices from $s$ that leads from $v_a$ to $v_c$ without including $v_b$.
+Furthermore,we will call the vertices closest to the text area's border to the routing area \emph{Border Vertices} $B \subset V$.
 They serve as a goal for the depth-first search, and are located at the end of each line in our implementation.}
@@
 \section{Description}
+\label{sec:AlgDesc}
 
@@
 
-The routing algorithm itself uses a depth-first-search algorithm, with the Graph $G$ and the set of sites $V_{ann} \subset V$ as input.
-The Algorithm terminates either when reaching the right text border or if all possible paths failed to reach the text border, and the resulting SP-path will be monotonically increasing.
-This path is split into two parts: The path within the graph $(P \subset V)$, leading from the source to the text area's border, and the OPO-Leader that connects the text area's border with the Label's port.
+We use a depth-first-search algorithm with the Graph $G$ and the set of sites $V_{ann} \subset V$ as input.
+The Algorithm \change{prioritizes routing to not yet visited vertices located above the current vertex} and terminates either when reaching the right text border or if all possible paths failed to reach the text border, and the resulting SP-path will be monotonically increasing.
+\change{If the routing for any site $v_{ann} \in V_{ann}$ failed, the algorithm returns $\bot$ for that site.
+To ensure that routings remain crossing-free, leaders are forbidden to incorporate any vertices that are either located directly above any vertices of the last successfully routed leader, or not between that leader's site and the border vertices.} This path is split into two parts: The path within the graph $(P \subset V)$, leading from the source to the text area's border, and the OPO-Leader that connects the text area's border with the Label's port.
 The former is implemented as a list of vertices $V_{Path} \subset V$ the leader travels through, whereas the latter is determined by the rightmost vertex of $V_{Path}$, the Port, and the location of the P-Segment.
-\change{If the routing for any given site failed, the algorithm returns $\bot$.} (For an illustration in Pseudocode see Alg.~\ref{alg:DFS}.)
+(For an illustration in Pseudocode see Alg.~\ref{alg:DFS}.)
 
@@
 
- \uIf{$($\G.\Top{\Curr}$\neq null)\land \neg$\Back}{
+ \uIf{$($\G.\Top{\Curr}$\neq null)\land \neg$\Back and previous Vertex not part of other leaders}{
   \Path.\AddVert{\Curr}\;
@@
 To reach this time, we'd have to try and use every single edge in the graph, reaching each of the graph's vertices in the process.
-However, as our algorithm can't use any edges leading away from its target, our search space is reduced to only the vertices either in lower-numbered $s \in S$ (``above'') or those in the same $s$, but located between the source and the border vertices.
+However, as our algorithm can't use any edges leading away from its target, our search space is reduced to only the vertices either in lower-numbered $s \in S$ (those ``above'' the current position) or those in the same $s$, but located between the source and the border vertices.
 If we also take into account that leaders aren't allowed to cross one another, this further limits our search space, removing all vertices above the previous leader's vertices.
@@
 
+To prove that the algorithm only creates crossing-free routings, we will examine the rules given in section~\ref{sec:AlgDesc}: All nodes that are either located directly above any nodes of the last successfully routed leader (whose path is represented by $P_{old}$) or not between the site $s_{old} \in P_{old}$ and the border vertices.
+For a crossing to occur, both the currently routed path $(P_{new})$ and $P_{old}$ must have one or several consecutive vertices in common, before which the nodes of one were below the other, and afterwardsthe opposite is true.
+This leaves us with two options:
+\begin{itemize}
+ \item $P_{new}$ crosses from below to above $P_{old}$: For this to happen, $P_{new}$ must at some point start to include vertices that were also used in $P_{old}$.
+While this is allowed, the routing algorithm was explicitely forbidden to use any vertices located directly above any vertices of $P_{old}$, which makes crossings impossible.
+ \item $P_{new}$ crosses from above to below $P_{old}$: While this type of crossing would be possible if $P_{new}$ managed to incorporate vertices that are above and not between $P_{old}$ and the border vertices, this requires the inclusion of nodes that are not between $s_{old}$ and the border vertices, as the currently routed leader's site $s_{new}$ is either located between $s_{old}$ and the border vertices, which wouldn't make this routing possible without violating the monotonicity constraint, or below $s_{old}$.
+Since the inclusion of these nodes was also explicitely forbidden, this type of crossing is impossible as well.
+\end{itemize}
+Therefore, the algorithm only produces crossing-free routings.
+\\~\\
+Finally, we want to prove that the algorithm only returns a highest legal path - a path that is both monotonically increasing and crossing-free, ends at the highest possible border vertex $b_{high} \in B$ reachable from a given source $s$, and contains no other border vertices than $b_{high}$.
+The path must also contain exactly one border vertex.
+The first two criteria are given, as the algorithm only returns monotonically increasing paths, and the existence of crossing was disproven above.
+To show that there are no reachable border vertices above $b_{high}$, we can assume there exists a vertex $b_{top} \in B$ that is located above $b_{high}$ and legally reachable via the path $P_{top}$.
+As the algorithm prioritizes routing to vertices directly above, the only way for the path to $b_{high}$ to reach higher vertices than $P_{top}$ would be that either $P_{top}$ would generate crossings with previously routed leaders or violate the the monotonicity constraint, which both render $P_{top}$ illegal.
+Since we also cannot include more than one border vertex in our path, we must conclude that $b_{top}$ cannot exist.
+
 %TODO: Re-Write!

### 2018-03-09 LaTeX/Paper.tex
Paper: Re-worded definition of SP-Paths and clarified how the sub-components of the text are related. Program: no changes.

@@
 \usepackage{subcaption}
+\usepackage{amsthm}
 
@@
 
+\newtheorem{lemma}{Lemma}
 \newcommand{\change}[1]{\textcolor{red}{#1}}
@@
 \chapter{Introduction}
-Whenever additional information neets to be inserted into an existing document without altering the original text, we can make use of annotations.
+Whenever additional information needs to be inserted into an existing document without altering the original text, we can make use of annotations.
 They usually take the form of footnotes, which require only a minimal reference in the main text, and are used for a variety of reasons - for example, to provide additional information that would hinder the text's flow if inserted directly, or as a result of a commenting tool that is used for communication between an author and their editor.
@@
 \section{Terminology and Fundamentals} 
-Boundary Labeling (or equivalent concepts) can be applied to a space with different geometry or more dimensions, but this thesis will only concern itself with two-dimensonal, Euclidean space.
+Boundary Labeling (or equivalent concepts) can be applied to a space with different geometry or more dimensions, but this thesis will only concern itself with two-dimensional, Euclidean space.
 To easily reference important concepts, some additional terminology will be introduced as well. (See Fig.~\ref{fig:term} for a visual explanation)
@@
 
-\change{A \emph{Polyline} is a curve defined by a sequence of points $O=p_1, p_2, ...,p_n$, connecting each point to its successor with straight lines.
-Each point $p \in O$ cannot be equal to any other point in $O$.
-We will call a polyline \emph{monotone}, if it satisfies certain conditions. For \emph{monotonically increasing} polylines, each two vertices must satisfy the following: \\
-$\forall p_i, p_j \in O: i<j \iff (X(p_i)\leq X(p_j) \land Y(p_i)\leq Y(p_j)\land p_i \neq p_j)$, where $X(.)$ returns a point's X-coordinate, and $Y(.)$ returns its Y-coordinate.
-For \emph{monotonically decreasing} polylines, the inverse must be true - each point must have at least one coordinate with a value smaller than its predecessor.}
+\change{A \emph{polyline} is a sequence of points $O=p_1, p_2, ...,p_n$, connecting each point to its successor with straight lines. Each point $p \in O$ cannot be equal to any other point in $O$.
+We will call a polyline $O$ \emph{monotonically increasing}, if any two vertices $p_i, p_j \in O$ satisfy the following: $\forall p_i, p_j \in O: i<j \iff (X(p_i)\leq X(p_j) \land Y(p_i)\leq Y(p_j)\land p_i \neq p_j)$, where $X(.)$ returns a point's X-coordinate, and $Y(.)$ returns its Y-coordinate.}
+
 \emph{Labels} hold additional information and are represented as boxes containing this information.
@@
 
-These restrictions are implemented as follows: We divided the text $T$ into separate lines $L=\{W,H\}$ which are in turn made up of words $W$ and whitespace $H$.
-The remaining space between lines is called $S=\{s_1, ...
+\change{These restrictions are implemented as follows: We divided the text $T=\{W,H\}$ into separate lines $L=\{l_1 ,l_2 , ...
+,l_n\}$ which are in turn made up of words $w \in W$ and whitespace $h \in H$, which alternate, starting and ending with a word, which creates a sequence $l=(w_1,h_1,w_2, ...
+, h_{m-1},w_m)$, where $m$ is the number of words in $l$.} The remaining space between lines is called $S=\{s_1, ...
 s_{n+1}\}$, with $n$ being the number of lines in $L$.
@@
 All $s \in S$ are the same height and width, even $s_1$ and $s_{n+1}$, which only have lines on one of their sides.
-For each word $w \in W$, we define $R(w)$  as its bounding rectangle, which marks the space leaders aren't allowed to cross.
+\change{The $s \in S$ and $l \in L$ are arranged in alternating order in the text $T=\{s_1, l_1, s_2, l_2, ..., s_n, l_n, s_{n+1}\}$.}  For each word $w \in W$, we define $R(w)$  as its bounding rectangle, which marks the space leaders aren't allowed to cross.
 For a graphical representation see Fig.~\ref{fig:wbound}.
@@
 The resulting graph looks similar to Fig.~\ref{fig:rgraph}.
-\change{In the following, we define vertex $v_a \in s_i$ to be \emph{above} another vertex $v_b \in s_j$, if $i<j$.
+In the following, we define vertex $v_a \in s_i$ to be \emph{above} another vertex $v_b \in s_j$, if $i<j$.
 If $v_a$ and $v_b$ are adjacent to each other, $v_a$ is \emph{directly above} $v_b$.
@@
 Furthermore,we will call the vertices closest to the text area's border to the routing area \emph{Border Vertices} $B \subset V$.
-They serve as a goal for the depth-first search, and are located at the end of each line in our implementation.}
+They serve as a goal for the depth-first search, and are located at the end of each line in our implementation.
 
@@
 
-\change{The Routing Area will be used to connect each site's path with its source, using OPO-Leaders.
-Combined with the path $P=\{s, v_1, v_2, ..., v_b\}; P \subset V$ leading from the source $s \in S$ to the graph's border $(v_b \in B)$ it creates an unbroken connection between a source and its port, which shall be called a \emph{Source-Port-Path}, or \emph{SP-Path}.}
-%For ease of visualisation, we limited the number of leaders per edge to one, and forbid the leaders to cross one another, as this could be mistaken for two leaders touching each other at the corners.
-We also decided to place the annotations in the order in which they appear in the text, and always place their port above the leader's source, which means that crossings are inevitable if a leader is located above and to the left of an earlier leader.
-To ensure this never happens, leaders aren't allowed to enter the top left quadrant of a source vertex, unless no routing could be found for that vertex, which means there won't be any leaders to collide with.
-Additionally, to keep the leader's length minimal, each leader is only allowed to move towards the source, which places a hard limit on how far up a leader is allowed to go, and restricts the possible choices at each node to edges that allow the leader to move in that direction.
-%To maximize the space available to subsequent annotations, each label is placed as far up as possible.
-This also means that if there's no placement available that is above its leader's source, the annotation cannot be placed, as we previously restricted the leader's movement.
+The Routing Area will be used to connect each site's path with its source, using OPO-Leaders.
+Combined with the path \change{$P=\{s, v_1, v_2, ..., v_h, v_b\}; P \subset V; P \cap B = v_b$} leading from the source $s \in S$ to the graph's border $(v_b \in B)$ it creates an unbroken connection between a source and its port, which shall be called a \emph{Source-Port-Path}, or \emph{SP-Path}.
 
@@
 We use a depth-first-search algorithm with the Graph $G$ and the set of sites $V_{ann} \subset V$ as input.
-The Algorithm \change{prioritizes routing to not yet visited vertices located above the current vertex} and terminates either when reaching the right text border or if all possible paths failed to reach the text border, and the resulting SP-path will be monotonically increasing.
-\change{If the routing for any site $v_{ann} \in V_{ann}$ failed, the algorithm returns $\bot$ for that site.
-To ensure that routings remain crossing-free, leaders are forbidden to incorporate any vertices that are either located directly above any vertices of the last successfully routed leader, or not between that leader's site and the border vertices.} This path is split into two parts: The path within the graph $(P \subset V)$, leading from the source to the text area's border, and the OPO-Leader that connects the text area's border with the Label's port.
+The Algorithm prioritizes routing to not yet visited vertices located above the current vertex and terminates either when reaching the right text border or if all possible paths failed to reach the text border, and the resulting SP-path will be monotonically increasing.
+If the routing for any site $v_{ann} \in V_{ann}$ failed, the algorithm returns $\bot$ for that site.
+To ensure that routings remain crossing-free, leaders are forbidden to incorporate any vertices that are either located directly above any vertices of the last successfully routed leader, or not between that leader's site and the border vertices.
+This path is split into two parts: The path within the graph $(P \subset V)$, leading from the source to the text area's border, and the OPO-Leader that connects the text area's border with the Label's port.
 The former is implemented as a list of vertices $V_{Path} \subset V$ the leader travels through, whereas the latter is determined by the rightmost vertex of $V_{Path}$, the Port, and the location of the P-Segment.

### 2018-03-11 LaTeX/Paper.tex
Paper: Re-wrote Section 2.3, adjusted definition of 'forbidden' vertices from 2.2 to fit the re-write. Program: No changes.

@@
 \newtheorem{lemma}{Lemma}
+\newtheorem{theorem}{Theorem}
 \newcommand{\change}[1]{\textcolor{red}{#1}}
@@
 ,l_n\}$ which are in turn made up of words $w \in W$ and whitespace $h \in H$, which alternate, starting and ending with a word, which creates a sequence $l=(w_1,h_1,w_2, ...
-, h_{m-1},w_m)$, where $m$ is the number of words in $l$.} The remaining space between lines is called $S=\{s_1, ...
+, h_{m-1},w_m)$, where $m$ is the number of words in $l$.
+The width of each line is equal to the text area's total width in characters $(width)$, which means that $m \leq \frac{width}{2}$.} The remaining space between lines is called $S=\{s_1, ...
 s_{n+1}\}$, with $n$ being the number of lines in $L$.
@@
 If the routing for any site $v_{ann} \in V_{ann}$ failed, the algorithm returns $\bot$ for that site.
-To ensure that routings remain crossing-free, leaders are forbidden to incorporate any vertices that are either located directly above any vertices of the last successfully routed leader, or not between that leader's site and the border vertices.
-This path is split into two parts: The path within the graph $(P \subset V)$, leading from the source to the text area's border, and the OPO-Leader that connects the text area's border with the Label's port.
+\change{To ensure that routings remain crossing-free, the most recently successfully routed leader is used as a ceiling to the available search space.
+This means that all vertices that are both above any of that leader's vertices and not between one of that leader's vertices and the border vertices are forbidden to use.} This path is split into two parts: The path within the graph $(P \subset V)$, leading from the source to the text area's border, and the OPO-Leader that connects the text area's border with the Label's port.
 The former is implemented as a list of vertices $V_{Path} \subset V$ the leader travels through, whereas the latter is determined by the rightmost vertex of $V_{Path}$, the Port, and the location of the P-Segment.
@@
  \item Computation time of graph-creation
- \item Computation time of the routing algorithm for single leaders
- \item Overall computation time of the routing algorithm
+ \item Computation time of the routing algorithm
 \end{itemize}
@@
 \end{itemize}
+~\\
+\begin{lemma}
+	The creation of the routing graph is in O(n).
+	\label{lem:GraphCreat}
+\end{lemma}
 
-To determine the program's computation time, we will first take a look at the creation of the graph.
-As discussed in section~\ref{sec:ProbSpec}, we will be placing a pair of vertices per whitespace character, as well as one additional pair at both the start and end of each line, which amounts to $2\times|H|+4\times|L|$ vertices total.
+\begin{proof}
+	As discussed in section~\ref{sec:ProbSpec}, each line holds at most $\frac{width}{2}$ words, with an appropriate amount of whitespace in between.
+Since $width$ is a constant value, the amount of vertices placed is only dependent on the number of lines in the text, therefore $|V|=O(n)$.
+	Similarly, each vertex has at most 4 edges, representing the two possible placements of P- or O-Segments starting from that vertex, respectively.
+Therefore, $|E|=O(n)$.
+As both the creation of new vertices and edges can be done in a constant amount of time, the overall computation time is in $O(n)$.
+\end{proof}
+
+%As discussed in section~\ref{sec:ProbSpec}, we will be placing a pair of vertices per whitespace character, as well as one additional pair at both the start and end of each line, which amounts to $2\times|H|+4\times|L|$ vertices total.
 Together with the sites $I$, which are also represented as vertices, this results in the total number of vertices $|V|=|I|+2\times|H|+4\times|L|$.
@@
 Finally, each edge placed in a space $s \in S$ above or below a line will be connected to each other, which is solved by connecting each vertex to the next unconnected vertex remaining in $s$.
-This leads to the creation of $|s|-1$ edges representing possible P-segments, which leads to a total of $|V|-|S|$ additional edges, making the total number of edges $|E|=(3\times|V|-2\times|S|-|I|)/2$.%Todo: look up fractions.
-If we assume that the creation of edges and vertices both take a similar amount of time, we are looking at roughly $5/2 \times |V|$ operations, which means that the algorithm scales linearly with the amount of vertices placed, which in turn is directly proportional to the number of words $|W|$ in the text.
+This leads to the creation of $|s|-1$ edges representing possible P-segments, which leads to a total of $|V|-|S|$ additional edges, making the total number of edges $|E|=(3\times|V|-2\times|S|-|I|)/2$.
+%If we assume that the creation of edges and vertices both take a similar amount of time, we are looking at roughly $5/2 \times |V|$ operations, which means that the algorithm scales linearly with the amount of vertices placed, which in turn is directly proportional to the number of words $|W|$ in the text.
 For a representation of the graph-generation algorithm in pseudocode, see Alg.~\ref{alg:GraphCreation}.
-
-Next, we shall look at the routing algorithm - since we use depth-first search, which is a well-known algorithm, the worst-case computation time is known to be $|V|+|E|$.
-To reach this time, we'd have to try and use every single edge in the graph, reaching each of the graph's vertices in the process.
-However, as our algorithm can't use any edges leading away from its target, our search space is reduced to only the vertices either in lower-numbered $s \in S$ (those ``above'' the current position) or those in the same $s$, but located between the source and the border vertices.
-If we also take into account that leaders aren't allowed to cross one another, this further limits our search space, removing all vertices above the previous leader's vertices.
 
@@
 
-To prove that the algorithm only creates crossing-free routings, we will examine the rules given in section~\ref{sec:AlgDesc}: All nodes that are either located directly above any nodes of the last successfully routed leader (whose path is represented by $P_{old}$) or not between the site $s_{old} \in P_{old}$ and the border vertices.
-For a crossing to occur, both the currently routed path $(P_{new})$ and $P_{old}$ must have one or several consecutive vertices in common, before which the nodes of one were below the other, and afterwardsthe opposite is true.
+%Next, we shall inspect the routing algorithm - since we use depth-first search, which is a well-known algorithm, the worst-case computation time is known to be $|V|+|E|$.
+As both $|V|$ and $|E|$ are in $O(n)$, so is our routing algorithm.
+
+%To reach this time, we'd have to try and use every single edge in the graph, reaching each of the graph's vertices in the process.
+However, as our algorithm can't use any edges leading away from its target, our search space is reduced to only the vertices either in lower-numbered $s \in S$ (those ``above'' the current position) or those in the same $s$, but located between the source and the border vertices.
+If we also take into account that leaders aren't allowed to cross one another, this further limits our search space, removing all vertices above the previous leader's vertices.
+
+\begin{lemma}
+	The algorithm only creates crossing-free routings.
+	\label{lem:CrosFree}
+\end{lemma}
+\begin{proof}
+	As stated in section~\ref{sec:AlgDesc}, we are forbidden to use any vertices above any the vertices ($V_{last}$) of the last successfully routed leader $(P_{last})$, unless they are between any $v \in V_{last}$ and the border vertices.
+As this makes it impossible to construct a route that incorporates any vertex directly above any $v \in V_{last}$, those two leaders cannot possibly cross each other.
+However, as this is true for any two consecutive leaders, each leader also cannot cross any of the leaders routed before $P_{last}$, as those are located on the other side of $P_{last}$, which contains only forbidden vertices.
+	Therefore, our algorithm is crossing-free.
+\end{proof}
+%To prove that the algorithm only creates crossing-free routings, we will examine the rules given in section~\ref{sec:AlgDesc}: All nodes that are either located directly above any nodes of the last successfully routed leader (whose path is represented by $P_{old}$) or not between the site $s_{old} \in P_{old}$ and the border vertices.
+%For a crossing to occur, both the currently routed path $(P_{new})$ and $P_{old}$ must have one or several consecutive vertices in common, before which the nodes of one were below the other, and afterwardsthe opposite is true.
 This leaves us with two options:
-\begin{itemize}
- \item $P_{new}$ crosses from below to above $P_{old}$: For this to happen, $P_{new}$ must at some point start to include vertices that were also used in $P_{old}$.
+%\begin{itemize}
+% \item $P_{new}$ crosses from below to above $P_{old}$: For this to happen, $P_{new}$ must at some point start to include vertices that were also used in $P_{old}$.
 While this is allowed, the routing algorithm was explicitely forbidden to use any vertices located directly above any vertices of $P_{old}$, which makes crossings impossible.
- \item $P_{new}$ crosses from above to below $P_{old}$: While this type of crossing would be possible if $P_{new}$ managed to incorporate vertices that are above and not between $P_{old}$ and the border vertices, this requires the inclusion of nodes that are not between $s_{old}$ and the border vertices, as the currently routed leader's site $s_{new}$ is either located between $s_{old}$ and the border vertices, which wouldn't make this routing possible without violating the monotonicity constraint, or below $s_{old}$.
+% \item $P_{new}$ crosses from above to below $P_{old}$: While this type of crossing would be possible if $P_{new}$ managed to incorporate vertices that are above and not between $P_{old}$ and the border vertices, this requires the inclusion of nodes that are not between $s_{old}$ and the border vertices, as the currently routed leader's site $s_{new}$ is either located between $s_{old}$ and the border vertices, which wouldn't make this routing possible without violating the monotonicity constraint, or below $s_{old}$.
 Since the inclusion of these nodes was also explicitely forbidden, this type of crossing is impossible as well.
-\end{itemize}
-Therefore, the algorithm only produces crossing-free routings.
-\\~\\
-Finally, we want to prove that the algorithm only returns a highest legal path - a path that is both monotonically increasing and crossing-free, ends at the highest possible border vertex $b_{high} \in B$ reachable from a given source $s$, and contains no other border vertices than $b_{high}$.
-The path must also contain exactly one border vertex.
-The first two criteria are given, as the algorithm only returns monotonically increasing paths, and the existence of crossing was disproven above.
-To show that there are no reachable border vertices above $b_{high}$, we can assume there exists a vertex $b_{top} \in B$ that is located above $b_{high}$ and legally reachable via the path $P_{top}$.
+%\end{itemize}
+%Therefore, the algorithm only produces crossing-free routings.
+\begin{lemma}
+	Our algorithm only returns a highest legal path - that is a path that's both monotonically increasing and crossing-free, ends at the highest possible border vertex $b_{high} \in B$ reachable from a given source $s$.
+	\label{lem:High}
+\end{lemma}
+\begin{proof}
+	The first two criteria are given, due to the algorithm only returns monotonically increasing paths, and Lemma~\ref{lem:CrosFree}.
+To show that there are no reachable border vertices above $b_{high}$, we will assume there exists a vertex $b_{top} \in B$ that is located above $b_{high}$ and legally reachable via the path $P_{top}$.
 As the algorithm prioritizes routing to vertices directly above, the only way for the path to $b_{high}$ to reach higher vertices than $P_{top}$ would be that either $P_{top}$ would generate crossings with previously routed leaders or violate the the monotonicity constraint, which both render $P_{top}$ illegal.
-Since we also cannot include more than one border vertex in our path, we must conclude that $b_{top}$ cannot exist.
+Since we also cannot include more than one border vertex in our path, we conclude that $b_{top}$ cannot exist.
+\end{proof}
 
-%TODO: Re-Write!
+\begin{theorem}
+	The routing problem  can be solved in $O(n^2)$ for all sources.
+\end{theorem}
+\begin{proof}
+	As part of the proof for Lemma~\ref{lem:GraphCreat}, we have shown that the total number of vertices $|V|$ is in $O(n)$. Since $S \subset V$, $|S|$ must too be in $O(n)$. 
+	We have proven our algorithm's solutions to be correct in the Lemmas \ref{lem:CrosFree} and \ref{lem:High}.
+	The performance of our routing algorithm is easily calculated, as depth-first-search has a worst-case-performance of $|V|+|E|$, which is in $O(n)$ for our problem, as $|E|$ is in $O(n)$ as well.
+(See Lemma~\ref{lem:GraphCreat})
+	Therefore, our overall computation time comes out to $|S| \times O(n)$, or $O(n^2)$.
+\end{proof}
 

### 2018-03-16 LaTeX/Paper.tex
Paper: Re-wrote explanation of search-space limitation by previous routings, added graphic. Program: No changes.

@@
 
-\change{A \emph{polyline} is a sequence of points $O=p_1, p_2, ...,p_n$, connecting each point to its successor with straight lines. Each point $p \in O$ cannot be equal to any other point in $O$.
-We will call a polyline $O$ \emph{monotonically increasing}, if any two vertices $p_i, p_j \in O$ satisfy the following: $\forall p_i, p_j \in O: i<j \iff (X(p_i)\leq X(p_j) \land Y(p_i)\leq Y(p_j)\land p_i \neq p_j)$, where $X(.)$ returns a point's X-coordinate, and $Y(.)$ returns its Y-coordinate.}
+A \emph{polyline} is a sequence of points $O=p_1, p_2, ...,p_n$, connecting each point to its successor with straight lines. Each point $p \in O$ cannot be equal to any other point in $O$.
+We will call a polyline $O$ \emph{monotonically increasing}, if any two vertices $p_i, p_j \in O$ satisfy the following: $\forall p_i, p_j \in O: i<j \iff (X(p_i)\leq X(p_j) \land Y(p_i)\leq Y(p_j)\land p_i \neq p_j)$, where $X(.)$ returns a point's X-coordinate, and $Y(.)$ returns its Y-coordinate.
 
@@
 
-\change{These restrictions are implemented as follows: We divided the text $T=\{W,H\}$ into separate lines $L=\{l_1 ,l_2 , ...
+These restrictions are implemented as follows: We divided the text $T=\{W,H\}$ into separate lines $L=\{l_1 ,l_2 , ...
 ,l_n\}$ which are in turn made up of words $w \in W$ and whitespace $h \in H$, which alternate, starting and ending with a word, which creates a sequence $l=(w_1,h_1,w_2, ...
 , h_{m-1},w_m)$, where $m$ is the number of words in $l$.
-The width of each line is equal to the text area's total width in characters $(width)$, which means that $m \leq \frac{width}{2}$.} The remaining space between lines is called $S=\{s_1, ...
+The width of each line is equal to the text area's total width in characters $(width)$, which means that $m \leq \frac{width}{2}$.
+The remaining space between lines is called $S=\{s_1, ...
 s_{n+1}\}$, with $n$ being the number of lines in $L$.
@@
 All $s \in S$ are the same height and width, even $s_1$ and $s_{n+1}$, which only have lines on one of their sides.
-\change{The $s \in S$ and $l \in L$ are arranged in alternating order in the text $T=\{s_1, l_1, s_2, l_2, ..., s_n, l_n, s_{n+1}\}$.}  For each word $w \in W$, we define $R(w)$  as its bounding rectangle, which marks the space leaders aren't allowed to cross.
+The $s \in S$ and $l \in L$ are arranged in alternating order in the text $T=\{s_1, l_1, s_2, l_2, ..., s_n, l_n, s_{n+1}\}$.
+For each word $w \in W$, we define $R(w)$  as its bounding rectangle, which marks the space leaders aren't allowed to cross.
 For a graphical representation see Fig.~\ref{fig:wbound}.
@@
 The Routing Area will be used to connect each site's path with its source, using OPO-Leaders.
-Combined with the path \change{$P=\{s, v_1, v_2, ..., v_h, v_b\}; P \subset V; P \cap B = v_b$} leading from the source $s \in S$ to the graph's border $(v_b \in B)$ it creates an unbroken connection between a source and its port, which shall be called a \emph{Source-Port-Path}, or \emph{SP-Path}.
+Combined with the path $P=\{s, v_1, v_2, ..., v_h, v_b\}; P \subset V; P \cap B = v_b$ leading from the source $s \in S$ to the graph's border $(v_b \in B)$ it creates an unbroken connection between a source and its port, which shall be called a \emph{Source-Port-Path}, or \emph{SP-Path}.
 
@@
 
+\begin{figure}
+	\captionsetup{justification=centering, margin=0.75cm}
+	\centering
+	\includegraphics[scale=1]{Ceiling.png}
+	\caption{Illustration of the search-space limitation by previous leaders}
+	\label{fig:ceil}
+\end{figure}
+
 We use a depth-first-search algorithm with the Graph $G$ and the set of sites $V_{ann} \subset V$ as input.
@@
 If the routing for any site $v_{ann} \in V_{ann}$ failed, the algorithm returns $\bot$ for that site.
-\change{To ensure that routings remain crossing-free, the most recently successfully routed leader is used as a ceiling to the available search space.
-This means that all vertices that are both above any of that leader's vertices and not between one of that leader's vertices and the border vertices are forbidden to use.} This path is split into two parts: The path within the graph $(P \subset V)$, leading from the source to the text area's border, and the OPO-Leader that connects the text area's border with the Label's port.
+\change{To ensure that routings remain crossing-free, the most recently successfully routed leader $P_last$ is used as a ceiling to the available search space.
+This means that any vertex above the polyline created by the vertices in $P_last$ and the connection of $P_last$'s source with the text area's border opposite of the label area is forbidden to use.
+(See als Fig.~\ref{fig:ceil})} This path is split into two parts: The path within the graph $(P \subset V)$, leading from the source to the text area's border, and the OPO-Leader that connects the text area's border with the Label's port.
 The former is implemented as a list of vertices $V_{Path} \subset V$ the leader travels through, whereas the latter is determined by the rightmost vertex of $V_{Path}$, the Port, and the location of the P-Segment.
@@
 \begin{proof}
-	As discussed in section~\ref{sec:ProbSpec}, each line holds at most $\frac{width}{2}$ words, with an appropriate amount of whitespace in between.
+	As discussed in Section~\ref{sec:ProbSpec}, each line holds at most $\frac{width}{2}$ words, with an appropriate amount of whitespace in between.
 Since $width$ is a constant value, the amount of vertices placed is only dependent on the number of lines in the text, therefore $|V|=O(n)$.
-	Similarly, each vertex has at most 4 edges, representing the two possible placements of P- or O-Segments starting from that vertex, respectively.
+	Similarly, each vertex has at most 4 incident edges, representing the two possible placements of P- or O-Segments starting from that vertex, respectively.
 Therefore, $|E|=O(n)$.

### 2018-03-19 LaTeX/Paper.tex
Paper:Re-worked parts of section 2.3, to fit the new definition of forbidden search space. Program: No changes.

@@
 
-In this section we will analyze the following:
+In this section we will first analyze the computation time required to create the graph.
+Afterwards, we will prove the correctness of our routing algorithm, and finally we will determine the algorithm's runtime complexity.
 
-\begin{itemize}
- \item Computation time of graph-creation
- \item Computation time of the routing algorithm
-\end{itemize}
-~\\
-Furthermore, we will prove that:
-\begin{itemize}
-\item The algorithm generates only crossing-free routings
-\item The algorithm always returns a highest (define!) legal path
-
-NOTE: The algorithm keeps the annnotations in reading order
-\end{itemize}
-~\\
+%\begin{itemize}
+% \item Computation time of graph-creation
+% \item Computation time of the routing algorithm
+%\end{itemize}
+%~\\
+%Furthermore, we will prove that:
+%\begin{itemize}
+%\item The algorithm generates only crossing-free routings
+%\item The algorithm always returns a highest (define!) legal path
+%
+%NOTE: The algorithm keeps the annnotations in reading order
+%\end{itemize}
+%~\\
 \begin{lemma}
@@
 \begin{proof}
-	As stated in section~\ref{sec:AlgDesc}, we are forbidden to use any vertices above any the vertices ($V_{last}$) of the last successfully routed leader $(P_{last})$, unless they are between any $v \in V_{last}$ and the border vertices.
-As this makes it impossible to construct a route that incorporates any vertex directly above any $v \in V_{last}$, those two leaders cannot possibly cross each other.
-However, as this is true for any two consecutive leaders, each leader also cannot cross any of the leaders routed before $P_{last}$, as those are located on the other side of $P_{last}$, which contains only forbidden vertices.
-	Therefore, our algorithm is crossing-free.
+	As stated in Section~\ref{sec:AlgDesc}, we are forbidden to use any vertices above \change{the polyline created by the last successfully routed leader $(P_{last})$ and the direct connection of $P_{last}$'s source and the left text area's border.
+As this makes it impossible to construct a route that incorporates any vertex above or to the left of $P_{last}$, those two leaders cannot possibly cross each other.
+This is true for any two consecutive leaders, therefore each leader also cannot cross any of the leaders routed before $P_{last}$, as those are located on the other side of $P_{last}$, which contains only forbidden vertices.
+	Therefore, our algorithm produces crossing-free routings.}
 \end{proof}
-%To prove that the algorithm only creates crossing-free routings, we will examine the rules given in section~\ref{sec:AlgDesc}: All nodes that are either located directly above any nodes of the last successfully routed leader (whose path is represented by $P_{old}$) or not between the site $s_{old} \in P_{old}$ and the border vertices.
-%For a crossing to occur, both the currently routed path $(P_{new})$ and $P_{old}$ must have one or several consecutive vertices in common, before which the nodes of one were below the other, and afterwardsthe opposite is true.
-This leaves us with two options:
-%\begin{itemize}
-% \item $P_{new}$ crosses from below to above $P_{old}$: For this to happen, $P_{new}$ must at some point start to include vertices that were also used in $P_{old}$.
-While this is allowed, the routing algorithm was explicitely forbidden to use any vertices located directly above any vertices of $P_{old}$, which makes crossings impossible.
-% \item $P_{new}$ crosses from above to below $P_{old}$: While this type of crossing would be possible if $P_{new}$ managed to incorporate vertices that are above and not between $P_{old}$ and the border vertices, this requires the inclusion of nodes that are not between $s_{old}$ and the border vertices, as the currently routed leader's site $s_{new}$ is either located between $s_{old}$ and the border vertices, which wouldn't make this routing possible without violating the monotonicity constraint, or below $s_{old}$.
-Since the inclusion of these nodes was also explicitely forbidden, this type of crossing is impossible as well.
-%\end{itemize}
-%Therefore, the algorithm only produces crossing-free routings.
+
+\change{For the next part, we first need to define what we consider a \emph{highest legal path}: A \emph{legal path} is one that satisfies both the constraints of monotonicity and being crossing-free, whereas a \emph{highest path} reaches the highest possible border vertex reachable from a given source.}
+
 \begin{lemma}
-	Our algorithm only returns a highest legal path - that is a path that's both monotonically increasing and crossing-free, ends at the highest possible border vertex $b_{high} \in B$ reachable from a given source $s$.
+	Our algorithm only returns a highest legal path.
 	\label{lem:High}
@@
 \begin{proof}
-	The first two criteria are given, due to the algorithm only returns monotonically increasing paths, and Lemma~\ref{lem:CrosFree}.
-To show that there are no reachable border vertices above $b_{high}$, we will assume there exists a vertex $b_{top} \in B$ that is located above $b_{high}$ and legally reachable via the path $P_{top}$.
-As the algorithm prioritizes routing to vertices directly above, the only way for the path to $b_{high}$ to reach higher vertices than $P_{top}$ would be that either $P_{top}$ would generate crossings with previously routed leaders or violate the the monotonicity constraint, which both render $P_{top}$ illegal.
-Since we also cannot include more than one border vertex in our path, we conclude that $b_{top}$ cannot exist.
+	The legality of any generated path is given, due to the algorithm only returning monotonically increasing paths, and Lemma~\ref{lem:CrosFree}.
+To show that there are no reachable border vertices above the path's highest reachable border vertex $b_{high}$, we will assume there exists a vertex $b_{top} \in B$ that is located above $b_{high}$ and legally reachable via the path $P_{top}$.
+As the algorithm prioritizes routing to vertices directly above, the only way for the path to $b_{top}$ to reach higher vertices than $b_{high}$ would be that either $P_{top}$ would generate crossings with previously routed leaders or violate the the monotonicity constraint, which both render $P_{top}$ illegal.
+Since we also cannot include more than one border vertex in our path, we conclude that such a path cannot exist.
 \end{proof}
@@
 \begin{theorem}
-	The routing problem  can be solved in $O(n^2)$ for all sources.
+	The routing problem can be solved with a worst-case time complexity of $O(n^2)$ for all sources.
 \end{theorem}

### 2018-03-22 LaTeX/Paper.tex
Paper: Started writing Chapter 3 - Implementation. Program: No changes.

@@
 
-%First, we create the graph that will be used for the routing part of the algorithm.
-To achieve this, we go through the annotated text word by word in reading order, and measure the length $le(w_i)$ of each word $w_i$ that is not part of an annotation to determine its placement in the text.
-At this point, we can also assign a line $l_j$ to it.
-Afterwards, we create new vertices above and below the whitespace characters $h_k, h_{k+1}$ preceding and succeeding $w_i$.
-These vertices are located in the space $s_j, s_{j+1}$ between the word's line and its surrounding lines $l_{j-1}$ and $l_{j+1}$, as described in the previous section.
-Whenever an annotation is encountered, a single vertex is added above the last non-annotation word in $s_j$ at $pos(h_k)+le(w_i)/2$.
-%For a representation of this process in pseudocode, see Alg.~\ref{alg:GraphCreation}.
-%To create leaders that exclusively use the space not taken up by a word's bounding rectangle, we decided to use a graph similar to the one Loose \cite{Loose2015} used.
-%As each vertex represents a physical location, they will have co-ordinates associated with them, and the vertices representing the sites will hold additional information regarding its leader and label.
-%The graph is constructed by placing vertices between the lines, located next to each corner of a word's bounding rectangle, with two consecutive words in a line sharing the two vertices associated with their adjacent corners.
-%For the sites, we inserted an additional vertex above the center of the word, which will serve as the leader's starting point.
-After placing all vertices, we created edges between each vertex and the closest horizontal neighbour to both sides, and between nodes that are located exactly above or below each other, and exactly one line apart.
-(For a representation in Pseudocode, see Alg.~\ref{alg:GraphCreation})
+\chapter{Implementation}
+The program was written in Java 1.8.0u40, using JGraphT 1.0.1\cite{JGraphT} as graph library. No other external libraries were used.
+The program itself can be found at~\cite{Program}.
 
-\chapter{Implementation}
-The program was written in Java 1.8.0u40, using JGraphT1.0.1\cite{JGraphT} as graph library.
-Since we only want to create leaders that don't intersect with the text, the graph was created alongside the placement of the words on the canvas.%, as it was easy to extract measurements at this point.
+\section{Architecture}
+ Most of the calculations are done side-by-side with the visualization, as the position of each word is both used in the creation of the routing graph and when displaying the final result.
+The routing itself is done separately from this process in its own class, as the finished routing graph suffices here.
+This allows for different  routing algorithms to be implemented separately.
+%Since we only want to create leaders that don't intersect with the text, the graph was created alongside the placement of the words on the canvas.%, as it was easy to extract measurements at this point.
 Due to some problems with Java's various methods of calculating linebreaks, this process was done completely by hand.
+\section{GUI}
+
+%TODO: Screenshot nehmen und beschriften!
+
 \section{Challenges}
+
+Displaying the text on-screen without losing the position of the whitespace characters turned out to be harder than expected.
+While Java offers a class that divides any string of characters in separate lines, the resulting object doesn't allow for the line's text to be extracted again, which makes it unusable for our purposes.
+This problem was solved by placing each word manually on the screen.
+For simplicity's sake, we only allowed line breaks between words.
+
+Another issue was the visualization of the routing graph and the leaders.
+While JGraphT offers ways to display graphs, it wasn't possible to influence the placement of individual nodes.
+Our solution was forgoing these options and drawing the graph manually, assigning coordinates to each vertex when created.
+Vertices are represented by small rectangles, whereas edges are represented with straight lines.
+To avoid confusion between crossing leaders and leaders that incorporate the same vertex in their path, but curve away from each other, all leaders are drawn with rounded edges.
+
+\section{Workflow}%<--What do I put here?
 

### 2018-03-23 LaTeX/Paper.tex
Paper: Added section 3.2 - GUI. Program: Source vertices are now always drawn. Updated incorrect comments.

@@
 
-%TODO: Screenshot nehmen und beschriften!
+The GUI of the program is very simplistic, as the main goal was to demonstrate the routing algorithm.
+As such, all customization options are located on the control panel above the drawing area, which contains the text, leaders and annotations.
+(See also Fig.~\ref{fig:ProgWindow})
+
+The control panel's first three options determine the font the program uses, whether it is bold or italicized and its size.
+The fourth and last option allows the selection of different routing algorithms.
+After changing any value, the program will automatically redraw the drawing area's contents, and re-do the routing.
+
+The drawing area itself can't be interacted with, it only serves to display the algorithm's results.
+\begin{figure}
+	\captionsetup{justification=centering, margin=0.75cm}
+	\centering
+	\includegraphics[scale=0.75]{ProgramWindowAnnotated.png}
+	\caption{<test>}
+	\label{fig:ProgWindow}
+\end{figure}
 

### 2018-03-23 Eclipse/Labeling/src/routingapp/GraphCreatorPanel.java

@@ -404,8 +404,9 @@ public class GraphCreatorPanel extends JComponent {
 	/**
 	 * Creates an instance of the requested routing type, if supported.
 	 * Currently recognized routing types are:
-	 * 		Greedy/Topmost			-		A greedy routing with backtracking, aiming to place the label as high up as possible. Can only place labels separately.
-	 * 		Greedy/Topmost (2-Pass)	-		Variation on the above, it will adjust all annotation's positions after finding a  place to move them closer to where the routing ends its path through the text. 
+	 * 		Greedy/Topmost				-		A greedy routing with backtracking, aiming to place the label as high up as possible. Can only place labels separately.
+	 * 		Greedy/Topmost (OPO-Leader)	-		Same as above, except it  will use OPO-Leaders to connect to the label instead of S-Leaders.  
+	 * 		Greedy/Topmost (2-Pass)		-		Variation on the above, it will adjust all annotation's positions after finding a  place to move them closer to where the routing ends its path through the text. 
 	 * @param routingType The parameter determining the type of routing that will be used.
 	 * @return
 	 * @throws
@@ -505,6 +506,9 @@ public class GraphCreatorPanel extends JComponent {
 			GraphTuple nextTuple=it.next();
 			GraphTuple currentTuple=nextTuple;
 			GraphTuple oldTuple=currentTuple;
+			
+			GraphTuple temp=info.getSource();
+			g.fillRect(temp.getX()-2,temp.getY()-2,4,4);
 			
 			GeneralPath route=new GeneralPath();
 			route.moveTo(nextTuple.getX(), nextTuple.getY());

### 2018-03-24 LaTeX/Paper.tex
Paper: Wrote Section 3.4, edited chapter 3. Program: No changes.

@@
 \section{Architecture}
- Most of the calculations are done side-by-side with the visualization, as the position of each word is both used in the creation of the routing graph and when displaying the final result.
-The routing itself is done separately from this process in its own class, as the finished routing graph suffices here.
-This allows for different  routing algorithms to be implemented separately.
-%Since we only want to create leaders that don't intersect with the text, the graph was created alongside the placement of the words on the canvas.%, as it was easy to extract measurements at this point.
+The program is divided into three parts. The GUI and program window only propagate user interaction to other parts of the program. It will be discussed further in the next section.
+The main drawing area handles the visualization of the results, calculates where to place the text on-screen, and also generates the routing graph, as this can be done alongside the text placement.
+The routing is done in a separate class, and can easily be switched out for another algorithm because of this.
+%The routing itself is handled separately from this process in its own class, as the finished routing graph suffices here.
+This facilitates the implementation of additional routing algorithms and allows for easy switching.
+The only other notable part of the program is the GUI, which will be discussed in the next section.%Since we only want to create leaders that don't intersect with the text, the graph was created alongside the placement of the words on the canvas.%, as it was easy to extract measurements at this point.
 Due to some problems with Java's various methods of calculating linebreaks, this process was done completely by hand.
 \section{GUI}
+\label{sec:GUI}
 
-The GUI of the program is very simplistic, as the main goal was to demonstrate the routing algorithm.
-As such, all customization options are located on the control panel above the drawing area, which contains the text, leaders and annotations.
+The GUI of the program is kept simple, as the main goal was to demonstrate the routing algorithm.
+All customization options are located on the control panel above the drawing area, which contains the text, leaders and annotations.
 (See also Fig.~\ref{fig:ProgWindow})
@@
 The fourth and last option allows the selection of different routing algorithms.
-After changing any value, the program will automatically redraw the drawing area's contents, and re-do the routing.
+After changing any value, the program will automatically be notified of the changes and react accordingly.
 
@@
 	\includegraphics[scale=0.75]{ProgramWindowAnnotated.png}
-	\caption{<test>}
+	\caption{Picture of the routing program}
 	\label{fig:ProgWindow}
@@
 Displaying the text on-screen without losing the position of the whitespace characters turned out to be harder than expected.
-While Java offers a class that divides any string of characters in separate lines, the resulting object doesn't allow for the line's text to be extracted again, which makes it unusable for our purposes.
+While Java offers a class that divides any string of characters in separate lines, the resulting line object doesn't allow for it's text to be extracted, which makes it unusable for our purposes.
 This problem was solved by placing each word manually on the screen.
-For simplicity's sake, we only allowed line breaks between words.
+For simplicity's sake, we only allowed line breaks between words, whenever a whitespace character was encountered.
 
 Another issue was the visualization of the routing graph and the leaders.
-While JGraphT offers ways to display graphs, it wasn't possible to influence the placement of individual nodes.
-Our solution was forgoing these options and drawing the graph manually, assigning coordinates to each vertex when created.
+While JGraphT offers ways to display graphs, it wasn't possible to influence the placement of individual vertices.
+Our solution was forgoing these options and drawing the graph manually, assigning coordinates to each vertex on creation.
 Vertices are represented by small rectangles, whereas edges are represented with straight lines.
-To avoid confusion between crossing leaders and leaders that incorporate the same vertex in their path, but curve away from each other, all leaders are drawn with rounded edges.
+To avoid confusion between crossing leaders and leaders that incorporate the same vertex in their path, but curve away from each other, all leaders are drawn with rounded corners.
+
+%Add Test data generation/retrieval here?
 
@@
 
+When the program starts, it first goes through the text word by word in reading order, placing each word $w \in W$  on-screen, and saving the content of each encountered annotation for later.
+As the words are placed, vertices are added to the graph after each word, and at the start and end of each line.
+Source vertices are inserted whenever an annotation is encountered, and contain a reference to their annotation in addition to the information contained in regular vertices.
+
+After the text has been processed, additional vertices are inserted into each line between the last word of that line and the right text border, located directly above or below vertices in the neighboring lines.
+Whenever a vertex is inserted exactly above or below another vertex in a neighboring line, they are immediately connected via an edge, whereas all vertices in the same space between lines $s \in S$ are connected in order of their x-coordinates after all vertices have been placed.
+
+Once the graph has been generated, a list of all source vertices in reading order is passed to the selected routing algorithm, which then determines leader routing and annotation placement for each of them and returns the results to the main program, which then draws the results.
+(For details on the routing algorithm, see section~\ref{sec:AlgDesc}.)
+
+If any of the parameters on the control panel (see section~\ref{sec:GUI}) are changed, the process re-starts from the beginning.
 %As the Graph's vertices represent fixed locations on the canvas, they each have coordinates associated with them, with the vertices representing the sites containing extra information, such as references to the corresponding annotation and to the leader connecting the two (if existent).

### 2018-04-01 Eclipse/Labeling/src/routingapp/TextGenerator.java
Paper: No changes. Program: Added TextGenerator.java(unfinished).

@@ -1,0 +1,12 @@
+package routingapp;
+/**
+ * String-generating class. Generates correctly formatted texts for use in the routing application.
+ * 
+ * @author Jakob Klinger
+ *
+ */
+public class TextGenerator {
+
+	int wordProb[]={0,};//TODO: Finish this!
+	
+}

### 2018-04-02 Eclipse/Labeling/src/routingapp/GraphCreatorPanel.java
Paper:No changes. Program: Finished TextGenerator.java and made it the source for the program's text.

@@ -24,8 +24,11 @@ import org.jgrapht.graph.ListenableUndirectedWeightedGraph;
 
 public class GraphCreatorPanel extends JComponent {
 
-	private String text=
-			"Dies ist ein Typoblindtext.\\note{This is an annotation} An ihm kann man sehen, ob alle Buchstaben da sind und wie sie aussehen. " +
+	private TextGenerator gen=new TextGenerator(0);//<--Change value to change the seed
+	private String text=gen.generateUniformText((short)10, /*(short)2,*/ 200);
+			
+			
+			/*"Dies ist ein Typoblindtext.\\note{This is an annotation} An ihm kann man sehen, ob alle Buchstaben da sind und wie sie aussehen. " +
 					"Manchmal benutzt man Worte wie Hamburgefonts, Rafgenduks\\note{This too} oder Handgloves, um\\note{This is an annotation with long words, like supercalifragilisticexpiralidocious} "+
 					"Schriften\\note{Source: http://www.blindtextgenerator.de/} zu testen. " +
 					"Manchmal Sätze, die alle Buchstaben des Alphabets enthalten - man nennt diese Sätze »Pangrams«. " +

### 2018-04-02 Eclipse/Labeling/src/routingapp/TextGenerator.java

@@ -1,4 +1,7 @@
 package routingapp;
+
+import java.util.Random;
+
 /**
  * String-generating class. Generates correctly formatted texts for use in the routing application.
  * 
@@ -7,6 +10,211 @@ package routingapp;
  */
 public class TextGenerator {
 
-	int wordProb[]={0,};//TODO: Finish this!
+	/*Probability that a word has at least this many characters:
+	 * 
+	 * 
+	 * For all analyzed words:
+	 * 1	1
+	 * 2	0.659639992
+	 * 3	0.20638369
+	 * 4	0.095874714
+	 * 5	0.06834893
+	 * 6	0.055361917
+	 * 7	0.047579259
+	 * 8	0.040930577
+	 * 9	0.035344251 
+	 * 10	0.030187643
+	 * 11	0.024911669
+	 * 12	0.020996467
+	 * 13	0.017296123
+	 * 14	0.013034759
+	 * 15	0.009859626
+	 * 16	0.006684492
+	 * >16	0.003342246
+	 * 
+	 * For Content words:
+	 * 1	1
+	 * 2	0.907959418
+	 * 3	0.803367849
+	 * 4	0.652755988
+	 * 5	0.537914444
+	 * 6	0.465327895
+	 * 7	0.404873967
+	 * 8	0.353205732
+	 * 9	0.305093609
+	 * 10	0.263047798
+	 * 11	0.218073423
+	 * 12	0.183662797
+	 * 13	0.15186696
+	 * 14	0.114213994
+	 * 15	0.086392637
+	 * 16	0.058571279
+	 * >16	0.02928564
+	 * 
+	 * Data taken from https://www.sciencedirect.com/science/article/pii/S0019995858902298 (Table III) and converted to percentages (See Tab3Calc.xslx for calculations.)
+	 */
 	
+	double wordProb[]={	1,
+						0.907959418,
+						0.803367849,
+						0.652755988,
+						0.537914444,
+						0.465327895,
+						0.404873967,
+						0.353205732,
+						0.305093609,
+						0.263047798,
+						0.218073423,
+						0.183662797,
+						0.15186696,
+						0.114213994,
+						0.086392637,
+						0.058571279,
+						0.02928564,
+						-1};//<--Probability is 0, but the RNG might still generate 0.0f 
+	
+	Random rng;
+	
+	/**
+	 * Creates a new instance of the text generator.
+	 * @param seed The seed used for the random number generator.
+	 */
[…]

### 2018-04-03 LaTeX/Paper.tex
Paper: Re-Wrote Chapter 3 to add separate sections for GUI, Drawing Area and the Routing Algorithm. Program: No changes.

@@
 \usepackage{amsthm}
+\usepackage[hyperfootnotes=false]{hyperref}
 
-%TODO: CHANGE!
 \setauthor{}{Jakob Klinger}{}{male}
@@
 \setthesis{bachelor}
-\setcurriculum{Bachelor's programme Software \& Information Engineering}{Bachelorstudium Software \& Information Engineering} %TODO!
+\setcurriculum{Bachelor's programme Software \& Information Engineering}{Bachelorstudium Software \& Information Engineering}
 
@@
 Boundary Labeling (or equivalent concepts) can be applied to a space with different geometry or more dimensions, but this thesis will only concern itself with two-dimensional, Euclidean space.
-To easily reference important concepts, some additional terminology will be introduced as well. (See Fig.~\ref{fig:term} for a visual explanation)
+To easily reference important concepts, some additional terminology will be introduced as well. (See Fig.~\ref*{fig:term} for a visual explanation)
 
@@
 \emph{S-Segments} are not required to have any particular orientation, and simply connect their start and ending points in a straight line.
-The leader's name is created by combining the name of the segments - for example, the blue leader from Fig~\ref{fig:term} would be classified as an OPOPO-Leader.
-The location where a leader connects to the label is called the \emph{port}. It can be restricted to pre-defined positions. %TODO: Re-insert monotonicity
+The leader's name is created by combining the name of the segments - for example, the blue leader from Fig~\ref*{fig:term} would be classified as an OPOPO-Leader.
+The location where a leader connects to the label is called the \emph{port}. It can be restricted to pre-defined positions.
 
@@
 \chapter{The Algorithm}
-%1 Zeile Einleitung hier einfügen - zb: Kapitelübersicht
+%TODO:1 Zeile Einleitung hier einfügen - zb: Kapitelübersicht
 \section{Problem Specification}
@@
 For each word $w \in W$, we define $R(w)$  as its bounding rectangle, which marks the space leaders aren't allowed to cross.
-For a graphical representation see Fig.~\ref{fig:wbound}.
+For a graphical representation see Fig.~\ref*{fig:wbound}.
 
@@
 The edges between the vertices are all perfectly horizontal or vertical, and do neither intersect with any bounding rectangle, nor any vertices other than their starting and ending vertex.
-The resulting graph looks similar to Fig.~\ref{fig:rgraph}.
+The resulting graph looks similar to Fig.~\ref*{fig:rgraph}.
 In the following, we define vertex $v_a \in s_i$ to be \emph{above} another vertex $v_b \in s_j$, if $i<j$.
@@
 If the routing for any site $v_{ann} \in V_{ann}$ failed, the algorithm returns $\bot$ for that site.
-\change{To ensure that routings remain crossing-free, the most recently successfully routed leader $P_last$ is used as a ceiling to the available search space.
-This means that any vertex above the polyline created by the vertices in $P_last$ and the connection of $P_last$'s source with the text area's border opposite of the label area is forbidden to use.
-(See als Fig.~\ref{fig:ceil})} This path is split into two parts: The path within the graph $(P \subset V)$, leading from the source to the text area's border, and the OPO-Leader that connects the text area's border with the Label's port.
+To ensure that routings remain crossing-free, the most recently successfully routed leader $P_{last}$ is used as a ceiling to the available search space.
+This means that any vertex above the polyline created by the vertices in $P_{last}$ and the horizontal connection of $P_{last}$'s source with the text area's border opposite of the label area is forbidden to use.
+(See also Fig.~\ref*{fig:ceil}) This path is split into two parts: The path within the graph $(P \subset V)$, leading from the source to the text area's border, and the OPO-Leader that connects the text area's border with the Label's port.
 The former is implemented as a list of vertices $V_{Path} \subset V$ the leader travels through, whereas the latter is determined by the rightmost vertex of $V_{Path}$, the Port, and the location of the P-Segment.
-(For an illustration in Pseudocode see Alg.~\ref{alg:DFS}.)
+(For an illustration in Pseudocode see Alg.~\ref*{alg:DFS}.)
 
@@
 
-%\begin{itemize}
-% \item Computation time of graph-creation
-% \item Computation time of the routing algorithm
-%\end{itemize}
-%~\\
-%Furthermore, we will prove that:
-%\begin{itemize}
-%\item The algorithm generates only crossing-free routings
-%\item The algorithm always returns a highest (define!) legal path
-%
-%NOTE: The algorithm keeps the annnotations in reading order
-%\end{itemize}
-%~\\
 \begin{lemma}
@@
 \begin{proof}
-	As discussed in Section~\ref{sec:ProbSpec}, each line holds at most $\frac{width}{2}$ words, with an appropriate amount of whitespace in between.
+	As discussed in Section~\ref*{sec:ProbSpec}, each line holds at most $\frac{width}{2}$ words, with an appropriate amount of whitespace in between.
 Since $width$ is a constant value, the amount of vertices placed is only dependent on the number of lines in the text, therefore $|V|=O(n)$.
@@
 
-%As discussed in section~\ref{sec:ProbSpec}, we will be placing a pair of vertices per whitespace character, as well as one additional pair at both the start and end of each line, which amounts to $2\times|H|+4\times|L|$ vertices total.
-Together with the sites $I$, which are also represented as vertices, this results in the total number of vertices $|V|=|I|+2\times|H|+4\times|L|$.
-Additionally, each vertex in a pair is connected to its counterpart, which creates $(|V|-|I|)/2$ edges, representing all possible P-Segments.
-Finally, each edge placed in a space $s \in S$ above or below a line will be connected to each other, which is solved by connecting each vertex to the next unconnected vertex remaining in $s$.
-This leads to the creation of $|s|-1$ edges representing possible P-segments, which leads to a total of $|V|-|S|$ additional edges, making the total number of edges $|E|=(3\times|V|-2\times|S|-|I|)/2$.
-%If we assume that the creation of edges and vertices both take a similar amount of time, we are looking at roughly $5/2 \times |V|$ operations, which means that the algorithm scales linearly with the amount of vertices placed, which in turn is directly proportional to the number of words $|W|$ in the text.
-For a representation of the graph-generation algorithm in pseudocode, see Alg.~\ref{alg:GraphCreation}.
+For a representation of the graph-generation algorithm in pseudocode, see Alg.~\ref*{alg:GraphCreation}.
 
@@
 
-%Next, we shall inspect the routing algorithm - since we use depth-first search, which is a well-known algorithm, the worst-case computation time is known to be $|V|+|E|$.
-As both $|V|$ and $|E|$ are in $O(n)$, so is our routing algorithm.
-
-%To reach this time, we'd have to try and use every single edge in the graph, reaching each of the graph's vertices in the process.
-However, as our algorithm can't use any edges leading away from its target, our search space is reduced to only the vertices either in lower-numbered $s \in S$ (those ``above'' the current position) or those in the same $s$, but located between the source and the border vertices.
-If we also take into account that leaders aren't allowed to cross one another, this further limits our search space, removing all vertices above the previous leader's vertices.
-
 \begin{lemma}
@@
 \begin{proof}
-	As stated in Section~\ref{sec:AlgDesc}, we are forbidden to use any vertices above \change{the polyline created by the last successfully routed leader $(P_{last})$ and the direct connection of $P_{last}$'s source and the left text area's border.
+	As stated in Section~\ref*{sec:AlgDesc}, we are forbidden to use any vertices above the polyline created by the last successfully routed leader $(P_{last})$ and the direct connection of $P_{last}$'s source and the left text area's border.
 As this makes it impossible to construct a route that incorporates any vertex above or to the left of $P_{last}$, those two leaders cannot possibly cross each other.
 This is true for any two consecutive leaders, therefore each leader also cannot cross any of the leaders routed before $P_{last}$, as those are located on the other side of $P_{last}$, which contains only forbidden vertices.
-	Therefore, our algorithm produces crossing-free routings.}
+	Therefore, our algorithm produces crossing-free routings.
 \end{proof}
 
-\change{For the next part, we first need to define what we consider a \emph{highest legal path}: A \emph{legal path} is one that satisfies both the constraints of monotonicity and being crossing-free, whereas a \emph{highest path} reaches the highest possible border vertex reachable from a given source.}
+For the next part, we first need to define what we consider a \emph{highest legal path}: A \emph{legal path} is one that satisfies both the constraints of monotonicity and being crossing-free, whereas a \emph{highest path} reaches the highest possible border vertex reachable from a given source.
 
@@
 \begin{proof}
-	The legality of any generated path is given, due to the algorithm only returning monotonically increasing paths, and Lemma~\ref{lem:CrosFree}.
+	The legality of any generated path is given, due to the algorithm only returning monotonically increasing paths, and Lemma~\ref*{lem:CrosFree}.
 To show that there are no reachable border vertices above the path's highest reachable border vertex $b_{high}$, we will assume there exists a vertex $b_{top} \in B$ that is located above $b_{high}$ and legally reachable via the path $P_{top}$.
@@
 \begin{proof}
-	As part of the proof for Lemma~\ref{lem:GraphCreat}, we have shown that the total number of vertices $|V|$ is in $O(n)$. Since $S \subset V$, $|S|$ must too be in $O(n)$. 
-	We have proven our algorithm's solutions to be correct in the Lemmas \ref{lem:CrosFree} and \ref{lem:High}.
+	As part of the proof for Lemma~\ref*{lem:GraphCreat}, we have shown that the total number of vertices $|V|$ is in $O(n)$. Since $S \subset V$, $|S|$ must too be in $O(n)$. 
+	We have proven our algorithm's solutions to be correct in the Lemmas \ref*{lem:CrosFree} and \ref*{lem:High}.
 	The performance of our routing algorithm is easily calculated, as depth-first-search has a worst-case-performance of $|V|+|E|$, which is in $O(n)$ for our problem, as $|E|$ is in $O(n)$ as well.
-(See Lemma~\ref{lem:GraphCreat})
+(See Lemma~\ref*{lem:GraphCreat})
 	Therefore, our overall computation time comes out to $|S| \times O(n)$, or $O(n^2)$.
@@
 \chapter{Implementation}
-The program was written in Java 1.8.0u40, using JGraphT 1.0.1\cite{JGraphT} as graph library. No other external libraries were used.
-The program itself can be found at~\cite{Program}.
+The program was written in Java 1.8.0u40, using JGraphT 1.0.1\footnote{\url{http://jgrapht.org/}} as graph library. No other external libraries were used.
+The program itself can be found at GitHub \footnote{\url{https://github.com/e1125755/AnnotationRouting}}.
 
@@
 The main drawing area handles the visualization of the results, calculates where to place the text on-screen, and also generates the routing graph, as this can be done alongside the text placement.
-The routing is done in a separate class, and can easily be switched out for another algorithm because of this.
-%The routing itself is handled separately from this process in its own class, as the finished routing graph suffices here.
-This facilitates the implementation of additional routing algorithms and allows for easy switching.
-The only other notable part of the program is the GUI, which will be discussed in the next section.%Since we only want to create leaders that don't intersect with the text, the graph was created alongside the placement of the words on the canvas.%, as it was easy to extract measurements at this point.
-Due to some problems with Java's various methods of calculating linebreaks, this process was done completely by hand.
+\change{The routing is done in a separate class, and can easily be exchanged with another algorithm because of this.}
 \section{GUI}
@@
 All customization options are located on the control panel above the drawing area, which contains the text, leaders and annotations.
-(See also Fig.~\ref{fig:ProgWindow})
+(See also Fig.~\ref*{fig:ProgWindow})
 
-The control panel's first three options determine the font the program uses, whether it is bold or italicized and its size.
-The fourth and last option allows the selection of different routing algorithms.
-After changing any value, the program will automatically be notified of the changes and react accordingly.
+The control panel's first three options determine the font the program uses for the main text, whether it is bold or italicized and its size.
+\change{The annotations' text currently uses a smaller version of this font.
+The fourth and last option allows the selection of different routing algorithms, although the current options are mostly variations on the routing algorithm described in Section~\ref*{sec:AlgDesc}.} After changing any value, the program will automatically be notified of the changes and react accordingly.
 
-The drawing area itself can't be interacted with, it only serves to display the algorithm's results.
-\begin{figure}
+\begin{figure}%TODO: Replace with better screenshot!
 	\captionsetup{justification=centering, margin=0.75cm}
@@
 
+\section{Drawing Area}
+\change{As this program is a proof of concept, the drawing area itself can't be interacted with, it only serves to display the algorithm's results.
+If we were to adapt this program for regular use, the user should be able to manually change routings or move labels around here.
+Due to reasons described below in Section~\ref*{sec:Challenge}, each word has to be placed individually, which allows us to create the routing graph alongside the darwing process. 
+After the text has been drawn completely, the routing algorithm is invoked from here, and its results are visualized.}
+
+\section{Routing Algorithm}
+
+\change{The Routing algorithm has already been described thoroughly in Section~\ref{sec:AlgDesc}, that process is implemented by creating ordered lists of the vertices traversed by the SP-Path, along with some additional information on where to draw the P-segment located in the Routing Area.
+While the algorithm offers methods to route each annotation individually, we currently route all annotations at the same time, which vastly decreases the required communication between the algorithm and the rest of the program.}
+
 \section{Challenges}
-
+\label{sec:Challenge}
 Displaying the text on-screen without losing the position of the whitespace characters turned out to be harder than expected.
-While Java offers a class that divides any string of characters in separate lines, the resulting line object doesn't allow for it's text to be extracted, which makes it unusable for our purposes.
-This problem was solved by placing each word manually on the screen.
+While Java offers a class that divides any string of characters into separate lines, the resulting line object does not allow for it's text to be extracted, which makes it unusable for our purposes.
+This problem was solved by placing each word on the screen individually.
 For simplicity's sake, we only allowed line breaks between words, whenever a whitespace character was encountered.
@@
 While JGraphT offers ways to display graphs, it wasn't possible to influence the placement of individual vertices.
-Our solution was forgoing these options and drawing the graph manually, assigning coordinates to each vertex on creation.
+Our solution was forgoing these options and drawing the graph \change{unassisted}, assigning coordinates to each vertex on creation.
 Vertices are represented by small rectangles, whereas edges are represented with straight lines.
-To avoid confusion between crossing leaders and leaders that incorporate the same vertex in their path, but curve away from each other, all leaders are drawn with rounded corners.
+To avoid confusion between crossing leaders and leaders that incorporate the same vertex in their path, but curve away from each other, all leaders are drawn with rounded corners, which can be seen in Fig.~\ref{fig:ProgWindow}.
 
@@
 
-When the program starts, it first goes through the text word by word in reading order, placing each word $w \in W$  on-screen, and saving the content of each encountered annotation for later.
+When the program starts, it first scans the text word by word in reading order, placing each word $w \in W$  on-screen, and saving the content of each encountered annotation for later.
 As the words are placed, vertices are added to the graph after each word, and at the start and end of each line.
@@
 Once the graph has been generated, a list of all source vertices in reading order is passed to the selected routing algorithm, which then determines leader routing and annotation placement for each of them and returns the results to the main program, which then draws the results.
-(For details on the routing algorithm, see section~\ref{sec:AlgDesc}.)
+(For details on the routing algorithm, see section~\ref*{sec:AlgDesc}.)
 
-If any of the parameters on the control panel (see section~\ref{sec:GUI}) are changed, the process re-starts from the beginning.
+If any of the parameters on the control panel (see section~\ref*{sec:GUI}) are changed, the process re-starts from the beginning.
+
+%TODO: Schematischen Graph (Flowchart?) von oben beschriebenem Ablauf erstellen!
 %As the Graph's vertices represent fixed locations on the canvas, they each have coordinates associated with them, with the vertices representing the sites containing extra information, such as references to the corresponding annotation and to the leader connecting the two (if existent).
@@
 %TODO: Schönheitsfeatures, etc.
-%TODO: Code genauer durchlesen und erweitern!
-
-% For demonstration purposes, we also included another algorithm that uses S-Leaders for this section.
 

### 2018-04-04 Eclipse/Labeling/src/routingapp/GraphCreatorPanel.java
Paper: No changes. Program:Added 'Testing mode' - currently unfinished.

@@ -23,12 +23,24 @@ import org.jgrapht.graph.GraphWalk;
 import org.jgrapht.graph.ListenableUndirectedWeightedGraph;
 
 public class GraphCreatorPanel extends JComponent {
+	
+	//DEBUG VALUES
+	private boolean testMode=true;//Toggles whether the program is in testing mode. If true, visualization is turned off, and multiple texts will be generated and routed. Overrides all other debug values.
+	private boolean showWordBoundaries=false;//Draws rectangles around detected word boundaries in main text, if set to true
+	private boolean showGraphGrid=false;//Draws the whole routing Graph 
+	private boolean hideLeaders=false; //Hides the leaders and unsuccessfully routed nodes, if some other feature needs to be inspected visually
+	private boolean hideText=false;//Disables drawing the text on-screen.
+	//DEBUG VALUES END
 
-	private TextGenerator gen=new TextGenerator(0);//<--Change value to change the seed
-	private String text=gen.generateUniformText((short)10, /*(short)2,*/ 200);
-			
-			
-			/*"Dies ist ein Typoblindtext.\\note{This is an annotation} An ihm kann man sehen, ob alle Buchstaben da sind und wie sie aussehen. " +
+	private TextGenerator gen=new TextGenerator(0);//<--Temporary value, will be changed before usage.
+	private int numberOfTests=100, textLength=200;
+	private int annMean=7, annSTDDevi=10; //Values for mean and standard deviation for normally distributed annotations. annMean is also used for uniformly distributed annotations.
+	private String textType="noramal"/**//*uniform"/**/;//Changes which type of text is used in the tests. Currently recognized values are "normal" and "uniform".
+	private String testResults;
+	
+	
+	private String text=
+					"Dies ist ein Typoblindtext.\\note{This is an annotation} An ihm kann man sehen, ob alle Buchstaben da sind und wie sie aussehen. " +
 					"Manchmal benutzt man Worte wie Hamburgefonts, Rafgenduks\\note{This too} oder Handgloves, um\\note{This is an annotation with long words, like supercalifragilisticexpiralidocious} "+
 					"Schriften\\note{Source: http://www.blindtextgenerator.de/} zu testen. " +
[…]

### 2018-04-04 Eclipse/Labeling/src/routingapp/GraphTuple.java

@@ -14,18 +14,18 @@ public class GraphTuple{
 	private Annotation annotation;
 	private RouteInfo routeinfo;
 	
-	public GraphTuple(int e1, int e2)
+	public GraphTuple(int xpos, int ypos)
 	{
-		x=e1;
-		y=e2;
+		x=xpos;
+		y=ypos;
 		name="Tuple";
 		annotation=null;
 		routeinfo=null;
 	}
-	public GraphTuple(String nm, int e1, int e2)
+	public GraphTuple(String name, int xpos, int ypos)
 	{
-		this(e1,e2);
-		name=nm;
+		this(xpos,ypos);
+		this.name=name;
 	}
 	
 	public void setAnnotation(Annotation ann)

### 2018-04-04 Eclipse/Labeling/src/routingapp/TextGenerator.java

@@ -85,15 +85,24 @@ public class TextGenerator {
 	}
 	
 	/**
+	 * Sets a new seed for the random number generator.
+	 * @param seed The seed used for the random number generator.
+	 */
+	public void setSeed(long seed)
+	{
+		rng.setSeed(seed);
+	}
+	
+	/**
 	 * Generates a text made up of random alphabetic characters with spaces and uniformly distributed annotations.
 	 * @param annFreq Determines the interval in which annotations appear in the text
 	 * @param textLength Length of the generated text in words
 	 * @return The text generated by this function as a string.
 	 */
-	String generateUniformText(short annFreq, int textLength)
+	String generateUniformText(int annFreq, int textLength)
 	{
 		String text="";
-		short annCounter=annFreq;
+		int annCounter=annFreq;
 		for(int i=0; i<textLength;i++)
 		{
 			String word="";
@@ -151,7 +160,7 @@ public class TextGenerator {
 	 * @param textLength The length of the text in words
 	 * @return The text generated by this function as a string.
 	 */
-	String generateNormalizedText(short annMean, short annDevi, int textLength)
+	String generateNormalizedText(int annMean, int annDevi, int textLength)
 	{
 		String text="";
 		short annCounter=(short)(rng.nextGaussian()*annDevi+annMean);

### 2018-04-05 Eclipse/Labeling/src/routingapp/GraphCreatorPanel.java
Paper: No changes. Program: Finished implementing Testing mode. Currently produces Human-readable output.

@@ -7,6 +7,14 @@ import java.awt.FontMetrics;
 import java.awt.Graphics;
 import java.awt.Graphics2D;
 import java.awt.geom.GeneralPath;
+import java.io.BufferedWriter;
+import java.io.FileOutputStream;
+import java.io.IOException;
+import java.io.OutputStreamWriter;
+import java.io.Writer;
+import java.time.Duration;
+import java.time.ZonedDateTime;
+import java.time.format.DateTimeFormatter;
 import java.util.ArrayList;
 import java.util.Iterator;
 import java.util.List;
@@ -34,8 +42,8 @@ public class GraphCreatorPanel extends JComponent {
 
 	private TextGenerator gen=new TextGenerator(0);//<--Temporary value, will be changed before usage.
 	private int numberOfTests=100, textLength=200;
-	private int annMean=7, annSTDDevi=10; //Values for mean and standard deviation for normally distributed annotations. annMean is also used for uniformly distributed annotations.
-	private String textType="noramal"/**//*uniform"/**/;//Changes which type of text is used in the tests. Currently recognized values are "normal" and "uniform".
+	private int annMean=7, annSTDDevi=30; //Values for mean and standard deviation for normally distributed annotations. annMean is also used for uniformly distributed annotations.
+	private String textType="normal"/**//*uniform"/**/;//Changes which type of text is used in the tests. Currently recognized values are "normal" and "uniform".
 	private String testResults;
 	
 	
@@ -130,6 +138,11 @@ public class GraphCreatorPanel extends JComponent {
 	 */
 	public void paintComponent(Graphics g) {
 		super.paintComponent(g);
+		ZonedDateTime startingTime=ZonedDateTime.now();
+		long routingStart = 0;
+		
+		g.setFont(this.getFont());
+		FontMetrics metrics = g.getFontMetrics();
 		
 		if(testMode)//Turn visualisation down as far as possible
 		{
@@ -137,28 +150,33 @@ public class GraphCreatorPanel extends JComponent {
 			hideText=true;
 			showGraphGrid=false;
 			showWordBoundaries=false;
-			testResults=	"Testing mode: "+textType+" distribution\n"+
[…]

### 2018-04-05 Eclipse/Labeling/src/routingapp/GreedyTopOpo.java

@@ -1,6 +1,5 @@
 package routingapp;
 
-import java.util.Iterator;
 import java.util.List;
 import java.util.TreeMap;
 

### 2018-04-05 Eclipse/Labeling/src/routingapp/OpoCalculator.java

@@ -53,7 +53,7 @@ public class OpoCalculator {
 					}
 					else
 					{
-						//TODO: implement cap? Might get silly with many parallel opo-leaders and little space to route through
+						//Maybe implement cap? Might get silly with many parallel opo-leaders and little space to route through
 						opoNumber++;
 						if(opoNumber>biggestOpoNumber) biggestOpoNumber=opoNumber;
 					}

### 2018-04-06 Eclipse/Labeling/src/routingapp/GraphCreatorPanel.java
Paper:No changes. Program: Fixed bug in test log naming, added CSV-Style output option. Misc: Added R script to evaluate CSV-Style testing data.

@@ -11,8 +11,6 @@ import java.io.BufferedWriter;
 import java.io.FileOutputStream;
 import java.io.IOException;
 import java.io.OutputStreamWriter;
-import java.io.Writer;
-import java.time.Duration;
 import java.time.ZonedDateTime;
 import java.time.format.DateTimeFormatter;
 import java.util.ArrayList;
@@ -34,6 +32,9 @@ public class GraphCreatorPanel extends JComponent {
 	
 	//DEBUG VALUES
 	private boolean testMode=true;//Toggles whether the program is in testing mode. If true, visualization is turned off, and multiple texts will be generated and routed. Overrides all other debug values.
+	private boolean humanReadableOutput=false;//Toggles whether the test results are raw data or with some additional descriptors. Does nothing if testMode==false 
+	
+	 //The following values are all overridden if testMode==true
 	private boolean showWordBoundaries=false;//Draws rectangles around detected word boundaries in main text, if set to true
 	private boolean showGraphGrid=false;//Draws the whole routing Graph 
 	private boolean hideLeaders=false; //Hides the leaders and unsuccessfully routed nodes, if some other feature needs to be inspected visually
@@ -143,6 +144,7 @@ public class GraphCreatorPanel extends JComponent {
 		
 		g.setFont(this.getFont());
 		FontMetrics metrics = g.getFontMetrics();
+		testResults="";
 		
 		if(testMode)//Turn visualisation down as far as possible
 		{
@@ -150,12 +152,20 @@ public class GraphCreatorPanel extends JComponent {
 			hideText=true;
 			showGraphGrid=false;
 			showWordBoundaries=false;
-			testResults=	"Test started at: "+startingTime.format(DateTimeFormatter.RFC_1123_DATE_TIME)+"\n"+
-							"Testing mode: "+textType+" distribution\n"+
-							"Font: "+this.getFont().getFontName()+", "+this.getFont().getSize()+" Pt\n"+
-							"Text length: "+textLength+" Words.\n"+
-							"Mean: "+annMean+"\n"+
-							"Standard Deviation: "+annSTDDevi+"\n";
+			
+			if(humanReadableOutput)
+			{
+				testResults+=	"Test started at: "+startingTime.format(DateTimeFormatter.RFC_1123_DATE_TIME)+"\n"+
[…]

### 2018-04-06 Eclipse/Labeling/test.r

@@ -1,0 +1,28 @@
+#establish error logging (Taken from: https://stackoverflow.com/questions/11666086/)
+zz <- file("Errorlog.Rout", open="wt")
+sink(zz, type="message")
+
+filename <- "Test_2018-04-06_17-37-47.log"#Replace with name of file to evaluate.
+
+results <- read.table(filename, sep=",", row.names=1, header=TRUE)
+results <- within(results, Sites.ratio <- Sites.successful/Sites.total)
+results <- within(results, Space.ratio <- Space.used/Space.total)
+results <- within(results, TimeToSites <- Time/Sites.total)
+
+boxplot(results$Time, main="Time elapsed (Outliers removed)", outline=FALSE)
+boxplot(results$Time, main="Time elapsed", outline=TRUE)
+boxplot(results$TimeToSites, main="Average time per site")
+boxplot(results$Sites.ratio, main="Percentage of successful routings")
+boxplot(results$Space.ratio, main="Percentage of Label Area usage")
+
+
+## reset message sink and close the file connection
+sink(type="message")
+close(zz)
+
+#User feedback
+colMax <- function(data) sapply(data, max, na.rm = TRUE)
+colMax(results)
+message("Calculations completed successfully.")
+message("Please do not close the window prematurely.")
+Sys.sleep(10)

### 2018-04-07 LaTeX/Paper.tex
Paper: Started Chapter 4 - Testing. Program: Added automatically generated R script for data visualisation. Misc: Updated .gitignore to disregard testing output.

@@
 
-\section{Data generation} %(Bin mit dem Titel unzufrieden, aber hier soll beschrieben werden, wie ich die einzelnen Testfälle erstellt habe.)
+\section{Data generation}
+
+To generate data, we created our own texts using word length statistics from Table III in "Length-frequency statistics for written English" by Miller et al.
+\cite{Miller1958}, which lists word frequencies by length.
+We opted to use the data from the "Content Words" column, as the combined statistics almost exclusively produced words with three or less letters in our tests.
+Using this data, we compiled our own fixed-length testing texts by creating words containing randomized alphabetic characters, taking care that the word length distribution matched our data.
+Annotation sites are inserted into the text on either a normal or uniform distribution, and their labels contain similarly generated texts of randomized length.
+
+To enable the inspection of individual texts, each text was generated from a randomized seed that is noted at time of creation.
 
@@
 
+To generate many datasets at once, each testing run consists of 100 randomly generated texts that are routed one after the other without drawing them on-screen.
+During this process, we measure the time taken to complete each text, as well as the amount of successful routings, the total space used by all labels and the amount of P-Segments within the text area.
+The results are saved to a CSV-style text file, and a R script is generated to visualize the data.
+
 \section{Results}
+
+%TODO:Run several tests with different parameters!
 

### 2018-04-07 Eclipse/Labeling/Eval_Template.r

@@ -1,17 +1,20 @@
+#THIS IS A TEMPLATE - DO NOT CHANGE ANYTHING!
+
 #establish error logging (Taken from: https://stackoverflow.com/questions/11666086/)
 zz <- file("Errorlog.Rout", open="wt")
 sink(zz, type="message")
 
-filename <- "Test_2018-04-06_17-37-47.log"#Replace with name of file to evaluate.
+filename <- REPLACEME
 
 results <- read.table(filename, sep=",", row.names=1, header=TRUE)
 results <- within(results, Sites.ratio <- Sites.successful/Sites.total)
 results <- within(results, Space.ratio <- Space.used/Space.total)
 results <- within(results, TimeToSites <- Time/Sites.total)
 
-boxplot(results$Time, main="Time elapsed (Outliers removed)", outline=FALSE)
-boxplot(results$Time, main="Time elapsed", outline=TRUE)
-boxplot(results$TimeToSites, main="Average time per site")
+with(results, plot(results[,'Time'], ylab="Time (ns)", main="Plot of time elapsed"))
+with(results, lines(results[,'Time']))
+boxplot(results$Time, ylab="Time (ns)", main="Time elapsed (Outliers removed)", outline=FALSE)
+boxplot(results$TimeToSites, main="Average time per site", outline=FALSE)
 boxplot(results$Sites.ratio, main="Percentage of successful routings")
 boxplot(results$Space.ratio, main="Percentage of Label Area usage")
 
@@ -25,4 +28,4 @@ colMax <- function(data) sapply(data, max, na.rm = TRUE)
 colMax(results)
 message("Calculations completed successfully.")
 message("Please do not close the window prematurely.")
-Sys.sleep(10)
+Sys.sleep(1)

### 2018-04-07 Eclipse/Labeling/src/routingapp/GraphCreatorPanel.java

@@ -7,8 +7,10 @@ import java.awt.FontMetrics;
 import java.awt.Graphics;
 import java.awt.Graphics2D;
 import java.awt.geom.GeneralPath;
+import java.io.BufferedReader;
 import java.io.BufferedWriter;
 import java.io.FileOutputStream;
+import java.io.FileReader;
 import java.io.IOException;
 import java.io.OutputStreamWriter;
 import java.time.ZonedDateTime;
@@ -44,7 +46,7 @@ public class GraphCreatorPanel extends JComponent {
 	private TextGenerator gen=new TextGenerator(0);//<--Temporary value, will be changed before usage.
 	private int numberOfTests=100, textLength=200;
 	private int annMean=7, annSTDDevi=30; //Values for mean and standard deviation for normally distributed annotations. annMean is also used for uniformly distributed annotations.
-	private String textType="normal"/**//*uniform"/**/;//Changes which type of text is used in the tests. Currently recognized values are "normal" and "uniform".
+	private String textType=/*"normal"/**/"uniform"/**/;//Changes which type of text is used in the tests. Currently recognized values are "normal" and "uniform".
 	private String testResults;
 	
 	
@@ -176,7 +178,7 @@ public class GraphCreatorPanel extends JComponent {
 			g.fillRect(0, 0, width, height);
 			g.setColor(Color.black);
 			
-			//Uncomment the following to inspect specific seeds in non-testing mode
+			//Uncomment the following lines to inspect specific seeds in non-testing mode
 			/*gen.setSeed(173466998178271232L);
 			text=gen.generateNormalizedText(annMean, annSTDDevi, textLength);/**/
 			
@@ -466,7 +468,7 @@ public class GraphCreatorPanel extends JComponent {
 				testResults+=evaluateResults(results);
 			}
 		}
-		if(testMode)//Save results to log file.
+		if(testMode)//Save results to log file and generate R file to graph results
 		{
 			try
 			{
@@ -483,6 +485,40 @@ public class GraphCreatorPanel extends JComponent {
 				writer.close();
 				g.setColor(new Color(0x00A000));
 				g.drawString("Results successfully logged under: "+filename, 30, 30);
+				
+				
+				
+				String rfile;//Create R file to plot results
+				BufferedReader br = new BufferedReader(new FileReader("Eval_Template.r"));
+				
+				try {
[…]

### 2018-04-09 LaTeX/Paper.tex
Paper: Started work on Chapter 4 - Testing. Program: Removed Human-readable output-mode, added extra information from that mode into a separate output file.

@@
 
-To enable the inspection of individual texts, each text was generated from a randomized seed that is noted at time of creation.
+To enable the inspection of individual texts, each text was generated from a randomized seed that is saved at time of creation.
 
@@
 
+All tests were run under 64-bit Windows 7 on a dual-core laptop\footnote{Dell Latitude E6400, Intel Core2 Duo, 4047 MB Virtual Memory on a SSD}.
+
 To generate many datasets at once, each testing run consists of 100 randomly generated texts that are routed one after the other without drawing them on-screen.
 During this process, we measure the time taken to complete each text, as well as the amount of successful routings, the total space used by all labels and the amount of P-Segments within the text area.
-The results are saved to a CSV-style text file, and a R script is generated to visualize the data.
+The results are saved to a CSV-style text file, and a R script is generated to visualize the data. Additional information regarding test parameters are logged to a separate file.
+Each test run can be repeated, its seed is part of the testing parameters.
 
@@
 
-%TODO:Run several tests with different parameters!
+\begin{figure}
+	\centering
+	\begin{subfigure}[b]{\textwidth}
+		\centering
+		\includegraphics[page=1, scale=0.50]{Rplots_n(5-0).pdf}
+		\caption{\label{fig:plot1_n5,0}}
+	\end{subfigure}
+	\\
+	~\\%There's probably a better way to force multiple newlines...
+	\begin{subfigure}[b]{\textwidth}
+		\centering
+		\includegraphics[page=1, scale=0.50]{Rplots_n(5-0)_repeat.pdf}
+		\caption{\label{fig:plot2_n5,0}}
+	\end{subfigure}
+	\caption{Two plots of the same test seed's time taken per text. Note the massive differences caused by the computer's scheduling.}
+	\label{fig:timeplot}
+\end{figure}
+
+Across every test run, the first two instances always took significantly longer to complete than subsequent ones.
+As this was replicated even when using the same text for all 100 passes, we believe this to be an issue caused by the implementation rather than the algorithm itself.
+(This effect can be seen in Fig.~\ref*{fig:timeplot})
+Similarly, the biggest fluctuations in computation time for the remaining instances were caused by the computer's scheduling, as Fig.~\ref*{fig:plot1_n5,0} and Fig.~\ref*{fig:plot2_n5,0} both depict time statistics for the same run, executed twice in a row.
+
+Unsurprisingly, the amount of labeling area space used by annotations increases with the total amount of annotations in the text, but the ratio of successful routings to total sites goes down. 
+%While increasing the standard deviation for the annotation placement process also allowed some texts to reach a better label area usage, the average results became worse for that category, whereas the percentage of successful routings moves closer to the 50\% mark.
 

### 2018-04-09 Eclipse/Labeling/Eval_Template.r

@@ -1,22 +1,26 @@
-#THIS IS A TEMPLATE - DO NOT CHANGE ANYTHING!
+#THIS IS A TEMPLATE - CHANGE AT YOUR OWN RISK!
 
 #establish error logging (Taken from: https://stackoverflow.com/questions/11666086/)
 zz <- file("Errorlog.Rout", open="wt")
 sink(zz, type="message")
 
+#Don't manipulate this line - it is automatically replaced by the program!
 filename <- REPLACEME
 
 results <- read.table(filename, sep=",", row.names=1, header=TRUE)
 results <- within(results, Sites.ratio <- Sites.successful/Sites.total)
 results <- within(results, Space.ratio <- Space.used/Space.total)
 results <- within(results, TimeToSites <- Time/Sites.total)
+results <- within(results, SpaceToSites <- Space.used/Sites.total)
 
-with(results, plot(results[,'Time'], ylab="Time (ns)", main="Plot of time elapsed"))
-with(results, lines(results[,'Time']))
+plot(results[,'Time'], ylab="Time (ns)", type="b", main="Plot of time elapsed")
 boxplot(results$Time, ylab="Time (ns)", main="Time elapsed (Outliers removed)", outline=FALSE)
-boxplot(results$TimeToSites, main="Average time per site", outline=FALSE)
+boxplot(results$TimeToSites, main="Average time per site (Outliers removed)", outline=FALSE)
 boxplot(results$Sites.ratio, main="Percentage of successful routings")
 boxplot(results$Space.ratio, main="Percentage of Label Area usage")
+plot(results$Sites.total,results$Space.used, xlab="Sites in text", ylab="Label area space used", main="Label area usage by number of sites")
+#with(results, abline(lm(Sites.total ~ Space.used)))
+
 
 
 ## reset message sink and close the file connection
@@ -24,8 +28,9 @@ sink(type="message")
 close(zz)
 
 #User feedback
+message("Maximum Values for:")
 colMax <- function(data) sapply(data, max, na.rm = TRUE)
 colMax(results)
 message("Calculations completed successfully.")
-message("Please do not close the window prematurely.")
+message("Please do not close the window prematurely.")#Closing the window early seems to corrupt the PDF with the plots
 Sys.sleep(1)

### 2018-04-09 Eclipse/Labeling/src/demos/GraphCreatorApplet.java

@@ -37,9 +37,9 @@ implements ChangeListener, ItemListener {
 	int sizeChoice = 12;
 	
 	//Names for the algorithms - they MUST be the same as in GraphCreatorPanel.getRouter()!
-	String algNames[]={	"Greedy/Topmost",
+	String algNames[]={	"Greedy/Topmost (2-Pass)",
 						"Greedy/Topmost (OPO-Leader)",
-						"Greedy/Topmost (2-Pass)"
+						"Greedy/Topmost"
 	};
 	String algChoice=algNames[0];
 

### 2018-04-09 Eclipse/Labeling/src/routingapp/GraphCreatorPanel.java

@@ -19,6 +19,7 @@ import java.util.ArrayList;
 import java.util.Iterator;
 import java.util.List;
 import java.util.Map.Entry;
+import java.util.Random;
 import java.util.Set;
 import java.util.TreeMap;
 
@@ -33,10 +34,9 @@ import org.jgrapht.graph.ListenableUndirectedWeightedGraph;
 public class GraphCreatorPanel extends JComponent {
 	
 	//DEBUG VALUES
-	private boolean testMode=true;//Toggles whether the program is in testing mode. If true, visualization is turned off, and multiple texts will be generated and routed. Overrides all other debug values.
-	private boolean humanReadableOutput=false;//Toggles whether the test results are raw data or with some additional descriptors. Does nothing if testMode==false 
+	private boolean testMode=true;//Toggles whether the program is in testing mode. If true, visualization is turned off, and multiple texts will be generated and routed. Overrides all other debug values. 
 	
-	 //The following values are all overridden if testMode==true
+	//The following values are all overridden if testMode==true
 	private boolean showWordBoundaries=false;//Draws rectangles around detected word boundaries in main text, if set to true
 	private boolean showGraphGrid=false;//Draws the whole routing Graph 
 	private boolean hideLeaders=false; //Hides the leaders and unsuccessfully routed nodes, if some other feature needs to be inspected visually
@@ -44,10 +44,8 @@ public class GraphCreatorPanel extends JComponent {
 	//DEBUG VALUES END
 
 	private TextGenerator gen=new TextGenerator(0);//<--Temporary value, will be changed before usage.
-	private int numberOfTests=100, textLength=200;
-	private int annMean=7, annSTDDevi=30; //Values for mean and standard deviation for normally distributed annotations. annMean is also used for uniformly distributed annotations.
-	private String textType=/*"normal"/**/"uniform"/**/;//Changes which type of text is used in the tests. Currently recognized values are "normal" and "uniform".
-	private String testResults;
+	private int numberOfTests=100, textLength=300;
[…]

### 2018-04-09 Eclipse/Labeling/src/routingapp/TextGenerator.java

@@ -95,6 +95,7 @@ public class TextGenerator {
 	
 	/**
 	 * Generates a text made up of random alphabetic characters with spaces and uniformly distributed annotations.
+	 * Currently unused, since generateNormalizedText produces the same results with a Standard Deviation of 0.
 	 * @param annFreq Determines the interval in which annotations appear in the text
 	 * @param textLength Length of the generated text in words
 	 * @return The text generated by this function as a string.

### 2018-04-11 LaTeX/Paper.tex
Paper: Updated system specs. Program: No changes.

@@
 
-All tests were run under 64-bit Windows 7 on a dual-core laptop\footnote{Dell Latitude E6400, Intel Core2 Duo, 4047 MB Virtual Memory on a SSD}.
+All tests were run under 64-bit Windows 7 Professional Version 6.1.7601 Service Pack 1 Build 7601 on a consumer-grade laptop using Java 1.8.0u40.
+The laptop is a Dell Latitude E6400 with an Intel Core2 Duo dual-core processor running at 2.53 GHz and 4 GB DDR2 RAM.
+There's 4047 MB virtual memory available on the SSD as well.
 
@@
 	\end{subfigure}
-	\\
-	~\\%There's probably a better way to force multiple newlines...
+
 	\begin{subfigure}[b]{\textwidth}

### 2018-04-12 Eclipse/Labeling/src/routingapp/GraphCreatorPanel.java
Paper: No changes. Program: Re-worked Text generation to have annotations distributed randomly instead of in random intervals. Normalized text generation currently unfinished.

@@ -34,8 +34,17 @@ import org.jgrapht.graph.ListenableUndirectedWeightedGraph;
 public class GraphCreatorPanel extends JComponent {
 	
 	//DEBUG VALUES
-	private boolean testMode=true;//Toggles whether the program is in testing mode. If true, visualization is turned off, and multiple texts will be generated and routed. Overrides all other debug values. 
-	
+	private boolean testMode=false;//Toggles whether the program is in testing mode. If true, visualization is turned off, and multiple texts will be generated and routed. Overrides all other debug values.
+	private String annDistribution=	"uniform";		//Only used if testMode==true - determines which type of annotation distribution is used for the randomized texts. 
+									//"top-left";	//Uncomment whatever value you want to use - "uniform" creates truly random annotations, whereas the others are normally distributed around a region.
+									//"top";
+									//"top-right";
+									//"left";
+									//"center";
+									//"right";
+									//"bottom-left";
+									//"bottom";
+									//"bottom-right";
 	//The following values are all overridden if testMode==true
 	private boolean showWordBoundaries=false;//Draws rectangles around detected word boundaries in main text, if set to true
 	private boolean showGraphGrid=false;//Draws the whole routing Graph 
@@ -43,9 +52,9 @@ public class GraphCreatorPanel extends JComponent {
 	private boolean hideText=false;//Disables drawing the text on-screen.
 	//DEBUG VALUES END
 
-	private TextGenerator gen=new TextGenerator(0);//<--Temporary value, will be changed before usage.
+	private TextGenerator gen=new TextGenerator(0);//<--Temporary seed, will be changed before use.
 	private int numberOfTests=100, textLength=300;
-	private int annMean=5, annSTDDevi=40; //Values for mean and standard deviation for normally distributed annotations. annMean is also used for uniformly distributed annotations.
+	private int annCount=15; //Values for mean and standard deviation for normally distributed annotations. annMean is also used for uniformly distributed annotations.
 	
 	
 	private String text=
@@ -163,9 +172,9 @@ public class GraphCreatorPanel extends JComponent {
 						"Testing batch seed:"+generatorseed+"\n"+
 						"Font: "+this.getFont().getFontName()+", "+this.getFont().getSize()+" Pt\n"+
[…]

### 2018-04-12 Eclipse/Labeling/src/routingapp/TextGenerator.java

@@ -1,6 +1,13 @@
 package routingapp;
 
+import java.awt.Canvas;
+import java.awt.Font;
+import java.awt.FontMetrics;
+import java.text.AttributedCharacterIterator;
+import java.util.ArrayList;
+import java.util.Map;
 import java.util.Random;
+import java.util.TreeMap;
 
 /**
  * String-generating class. Generates correctly formatted texts for use in the routing application.
@@ -71,7 +78,7 @@ public class TextGenerator {
 						0.086392637,
 						0.058571279,
 						0.02928564,
-						-1};//<--Probability is 0, but the RNG might still generate 0.0f 
+						-1};//<--Actual Probability is 0, but the RNG might still generate 0.0f 
 	
 	Random rng;
 	
@@ -95,15 +102,24 @@ public class TextGenerator {
 	
 	/**
 	 * Generates a text made up of random alphabetic characters with spaces and uniformly distributed annotations.
-	 * Currently unused, since generateNormalizedText produces the same results with a Standard Deviation of 0.
-	 * @param annFreq Determines the interval in which annotations appear in the text
+	 * @param annNum Determines the number of annotations that appear in the text - at most one per word. 
 	 * @param textLength Length of the generated text in words
 	 * @return The text generated by this function as a string.
 	 */
-	String generateUniformText(int annFreq, int textLength)
+	String generateUniformText(int annNum, int textLength)
 	{
 		String text="";
-		int annCounter=annFreq;
+		if(annNum>textLength)annNum=textLength; //There can't be more than one annotation per word
+		ArrayList<Integer> annLocations=new ArrayList<Integer>(annNum);
+		for(int i=0;i<annNum;i++)
+		{
+			int rand=rng.nextInt(textLength);
+			while(annLocations.contains(rand))
+			{
+				rand=rng.nextInt(textLength);
+			}
+			annLocations.add(rand);
+		}
 		for(int i=0; i<textLength;i++)
 		{
 			String word="";
@@ -120,14 +136,10 @@ public class TextGenerator {
 					word+=(char)('a'+rng.nextInt(25));
 				}
 				result=rng.nextFloat();
-				//System.out.println("Rolled "+result);
 			}
-			//System.out.println(word);
 			text+=word+" ";
-			annCounter--;
-			//System.out.println("AnnCounter="+annCounter);
 			
[…]

### 2018-04-13 Eclipse/Labeling/src/routingapp/GraphCreatorPanel.java
Paper: No changes. Program: Fixed bug that caused words with annotations to be measured differently than those without. Normalized text generation is now usable, but results are not yet as intended.

@@ -35,12 +35,12 @@ public class GraphCreatorPanel extends JComponent {
 	
 	//DEBUG VALUES
 	private boolean testMode=false;//Toggles whether the program is in testing mode. If true, visualization is turned off, and multiple texts will be generated and routed. Overrides all other debug values.
-	private String annDistribution=	"uniform";		//Only used if testMode==true - determines which type of annotation distribution is used for the randomized texts. 
+	private String annDistribution=	//"uniform";		//Only used if testMode==true - determines which type of annotation distribution is used for the randomized texts. 
 									//"top-left";	//Uncomment whatever value you want to use - "uniform" creates truly random annotations, whereas the others are normally distributed around a region.
 									//"top";
 									//"top-right";
 									//"left";
-									//"center";
+									"center";
 									//"right";
 									//"bottom-left";
 									//"bottom";
@@ -190,7 +190,7 @@ public class GraphCreatorPanel extends JComponent {
 			
 			//Uncomment the following lines to inspect specific seeds in non-testing mode
 			gen.setSeed(2457556750776815616L);
-			text=gen.generateNormalizedText(annCount, textLength, rightTextBorder-leftTextBorder, metrics, annDistribution);/**/
+			text=gen.generateNormalizedText(annCount, textLength, rightTextBorder-leftTextBorder-1, metrics, annDistribution);/**/
 			
 			if(testMode)
 			{
@@ -227,7 +227,7 @@ public class GraphCreatorPanel extends JComponent {
 			//While it is possible to have line breaks in more places than after each word, this will be ignored for ease of implementation.
 			for(int i=0;i<words.length; i++)
 			{
-				if((metrics.stringWidth(words[i].split("\\\\")[0])+x)>rightTextBorder)//Four backslashes, since both the String and the Regex use '\' as escape character
+				if((metrics.stringWidth(words[i].split("\\\\")[0]+" ")+x)>rightTextBorder)//Four backslashes, since both the String and the Regex use '\' as escape character
 				{
 					//Add one more node-pair after the last word
 					x-=metrics.stringWidth(" ");
@@ -243,7 +243,7 @@ public class GraphCreatorPanel extends JComponent {
 					
 					lineEnds.add(x);
 					
-					//Create single tuple at text border, it will be propagated downwards
[…]

### 2018-04-13 Eclipse/Labeling/src/routingapp/GraphTuple.java

@@ -1,8 +1,9 @@
 package routingapp;
 /**
  * Custom Graph node for the annotation routing program.
- * It currently contains its location as coordinates, and can be named.
- * Optionally, an Annotation can be set, which should in turn have this node listed as its associated node. 
+ * It contains its location as coordinates, and can be named.
+ * Optionally, an Annotation can be set, which should in turn have this node listed as its associated node.
+ * The routeinfo attribute contains all routing information from this site to its label. This implies that an annotation is set was well, although it is not enforced.
  * @author Jakob Klinger
  *
  */

### 2018-04-13 Eclipse/Labeling/src/routingapp/TextGenerator.java

@@ -1,16 +1,15 @@
 package routingapp;
 
-import java.awt.Canvas;
-import java.awt.Font;
 import java.awt.FontMetrics;
-import java.text.AttributedCharacterIterator;
 import java.util.ArrayList;
-import java.util.Map;
+import java.util.Comparator;
+import java.util.Iterator;
 import java.util.Random;
-import java.util.TreeMap;
+import java.util.TreeSet;
 
 /**
  * String-generating class. Generates correctly formatted texts for use in the routing application.
+ * NOTE: Generating words is done the same way in all methods - it should be put into a separate method.
  * 
  * @author Jakob Klinger
  *
@@ -168,7 +167,9 @@ public class TextGenerator {
 
 	
 	/**
-	 * Generates a text with words made of random alphabetic characters and normally distributed annotations, focused on a region of the text. 
+	 * Generates a text with words made of random alphabetic characters and normally distributed annotations, focused on a region of the text.
+	 * To do so, this method must use the same process of dividing the text into separate lines as GraphCreatorPanel.java.
+	 * Any changes to that process must be applied here as well, or unintended results will happen!  
 	 * @param textLength The length of the text in words
 	 * @param annNum How many annotations the text will contain. 
 	 * @param textWidth Space available to the text, measured in pixels - needed to split the text into lines to define text regions.
@@ -183,7 +184,7 @@ public class TextGenerator {
 	{
 		String text="";
 		double vertMean=0.5, horizMean=0.5;//Default values - remain unchanged if annMode==center
-		double vertStDev=0.33, horizStDev=0.33;
+		double vertStDev=0.33/2, horizStDev=0.33/2;
 		
 		if(annMode.contains("top"))
 		{
@@ -232,27 +233,67 @@ public class TextGenerator {
 		
 		String textByLine[]=text.split("\n");
 		
-	
-		//generate Annotations
-		/*text+="\\note{";
+		//Generate annotations
+		//NOTE: Using GraphTuples for positions mostly because Java has no in-built Tuple class. Coordinates used here are based on lines and character count, whereas 
+		TreeSet<GraphTuple> locations=new TreeSet<GraphTuple>(new Comparator<GraphTuple>(){
+
+			@Override
[…]

### 2018-04-22 LaTeX/Paper.tex
Paper: Re-wrote 4.1 to reflect new method of data generation. Program: Changed R script to include a combined histogram/boxplot of a test batch, grouped by number of successful routings.

@@
 Using this data, we compiled our own fixed-length testing texts by creating words containing randomized alphabetic characters, taking care that the word length distribution matched our data.
-Annotation sites are inserted into the text on either a normal or uniform distribution, and their labels contain similarly generated texts of randomized length.
+\change{A fixed number of annotation sites is inserted into the text either by distributing their locations uniformly across the text, or by clustering them regionally.}
+
+\change{For the clustering approach, we divided the text vertically and horizontally into three subsections each - top/center/bottom and left/center/right respectively.
+To determine the site's location, we then determined the line the site will be in and its exact word by generating numbers from two normal distributions, each with a standard deviation of $1/6$th of the text's width or height, and a mean located at 25\%, 50\% or 75\% of that parameter, depending on the chosen region.}
+
+\change{The uniform distribution is achieved by choosing each annotated word in advance, regardless of its relative location within the text.}
 
 To enable the inspection of individual texts, each text was generated from a randomized seed that is saved at time of creation.
+\change{To enable the repetition of full test batches, each of them was also generated using a seed, which is recorded as well.}
 
-\section{Testing methods}
+\section{Experimental Setup}
 

### 2018-04-22 Eclipse/Labeling/Eval_Template.r

@@ -4,12 +4,14 @@
 zz <- file("Errorlog.Rout", open="wt")
 sink(zz, type="message")
 
+library(ggplot2)
+
 #Don't manipulate this line - it is automatically changed by the program to point at the current output file!
 filename <- REPLACEME
 
 results <- read.table(filename, sep=",", row.names=1, header=TRUE)
-results <- within(results, Sites.ratio <- Sites.successful/Sites.total)
-results <- within(results, Space.ratio <- Space.used/Space.total)
+results <- within(results, Sites.ratio <- Sites.successful/Sites.total*100)
+results <- within(results, Space.ratio <- Space.used/Space.total*100)
 results <- within(results, TimeToSites <- Time/Sites.total)
 results <- within(results, SpaceToSites <- Space.used/Sites.total)
 
@@ -18,8 +20,16 @@ boxplot(results$Time, ylab="Time (ns)", main="Time elapsed (Outliers removed)", 
 boxplot(results$TimeToSites, main="Average time per site (Outliers removed)", outline=FALSE)
 boxplot(results$Sites.ratio, main="Percentage of successful routings")
 boxplot(results$Space.ratio, main="Percentage of Label Area usage")
-plot(results$Sites.successful,results$Space.ratio, xlab="Successfully routed sites", ylab="Label area space used (%)", main="Label area usage by number of successful routings")
-#with(results, abline(lm(Sites.successful ~ Space.ratio)))
+a <- ggplot(results)
+a <- a+geom_bar(aes(factor(results$Sites.successful)),fill="lightgoldenrod2")
+a <- a+stat_boxplot(aes(factor(results$Sites.successful),results$Space.ratio), geom="errorbar")#Workaround to get T-shaped whiskers
+a <- a+geom_boxplot(aes(factor(results$Sites.successful),results$Space.ratio))
+
+a
+
+
+#plot(results$Sites.successful,results$Space.ratio, xlab="Successfully routed sites", ylab="Label area space used (%)", main="Label area usage by number of successful routings")
+
 
 
 
@@ -32,5 +42,5 @@ message("Maximum Values for:")
 colMax <- function(data) sapply(data, max, na.rm = TRUE)
 colMax(results)
 message("Calculations completed successfully.")
-message("Please do not close the window prematurely.")#Closing the window early seems to corrupt the PDF with the plots
+message("Please do not close the window prematurely.")#Closing the shell window early seems to corrupt the PDF with the plots
 Sys.sleep(1)

### 2018-04-22 Eclipse/Labeling/src/routingapp/GraphCreatorPanel.java

@@ -34,7 +34,7 @@ import org.jgrapht.graph.ListenableUndirectedWeightedGraph;
 public class GraphCreatorPanel extends JComponent {
 	
 	//DEBUG VALUES
-	private boolean testMode=true;//Toggles whether the program is in testing mode. If true, visualization is turned off, and multiple texts will be generated and routed. Overrides all other debug values.
+	private boolean testMode=false;//Toggles whether the program is in testing mode. If true, visualization is turned off, and multiple texts will be generated and routed. Overrides all other debug values.
 	private String annDistribution=	//"uniform";		//Only used if testMode==true - determines which type of annotation distribution is used for the randomized texts. 
 									//"top-left";	//Uncomment whatever value you want to use - "uniform" creates truly random annotations, whereas the others are normally distributed around a region.
 									//"top";
@@ -47,7 +47,7 @@ public class GraphCreatorPanel extends JComponent {
 									"bottom-right";
 	//The following values are all overridden if testMode==true
 	private boolean showWordBoundaries=false;//Draws rectangles around detected word boundaries in main text, if set to true
-	private boolean showGraphGrid=false;//Draws the whole routing Graph 
+	private boolean showGraphGrid=true;//Draws the whole routing Graph 
 	private boolean hideLeaders=false; //Hides the leaders and unsuccessfully routed nodes, if some other feature needs to be inspected visually
 	private boolean hideText=false;//Disables drawing the text on-screen.
 	//DEBUG VALUES END

### 2018-04-22 LaTeX/Paper.tex
Paper:Re-wrote Section 4.3. Program: Minor changes/fixes.

@@
 
-Unsurprisingly, the amount of labeling area space used by annotations increases with the total amount of annotations in the text, but the ratio of successful routings to total sites goes down. 
+\change{Adding more sources to the text increases the both the space usage of the labeling area and the number of successfully routes, but there seems to be a limit, as increasing the amount of sources only brings marginal improvements after a certain point.
+For our experiments, this was reached at 15 sources in a 300-word text.
+Doubling their amount improved both the space usage and the amount of routed leaders for several test instances, but the overall highest amount of successful routings increased only by 1, whereas the increase from 10 to 15 sources impacted the results much stronger.
+(See Fig.~\ref*{fig:uniformstats} for details.)}
+
+ \begin{figure}
+ 	\centering
+ 	\begin{subfigure}[b]{\textwidth}
+ 		\centering
+ 		\includegraphics[scale=0.8]{Rplot_10u.png}
+ 		\caption{\label{fig:10u}}
+ 	\end{subfigure}
+
+ 	\begin{subfigure}[b]{\textwidth}
+ 		\centering
+ 		\includegraphics[scale=0.8]{Rplot_15u.png}
+ 		\caption{\label{fig:15u}}
+ 	\end{subfigure}
+    \begin{subfigure}[b]{\textwidth}
+		 \centering
+		 \includegraphics[scale=0.8]{Rplot_30u.png}
+		 \caption{\label{fig:30u}}
+    \end{subfigure}
+
+ 	\caption{Three plots of test batches with identical seeds. The only difference is the number of sources in the generated texts: 10, 15 and 30 respectively.}
+ 	\label{fig:uniformstats}
+ \end{figure}
+
+\change{While the placement of sources definitely had an impact on the amount of successful routings, the influence of horizontal placement is far smaller than that of the vertical placement.
+This is likely due to the annotations being placed vertically, which means that the more annotations are placed, the less likely it becomes for sources close to the beginning of the text to be routed.
+Together with the Routing Area, which allows for labels to be placed further up, this means that placements in the first few lines are disadvantaged compared to those further down.}
+
+\change{The effects of clustering are currently minor, and mostly outweighed by the influence of vertical placement, as the average word length is short enough that there is enough whitespace in each line to avoid leaders from blocking each other.
+Increasing the word length would make this effect more severe, and it would be interesting to see the influence on the results.}
 %While increasing the standard deviation for the annotation placement process also allowed some texts to reach a better label area usage, the average results became worse for that category, whereas the percentage of successful routings moves closer to the 50\% mark.

### 2018-04-22 Eclipse/Labeling/Eval_Template.r

@@ -24,8 +24,10 @@ a <- ggplot(results)
 a <- a+geom_bar(aes(factor(results$Sites.successful)),fill="lightgoldenrod2")
 a <- a+stat_boxplot(aes(factor(results$Sites.successful),results$Space.ratio), geom="errorbar")#Workaround to get T-shaped whiskers
 a <- a+geom_boxplot(aes(factor(results$Sites.successful),results$Space.ratio))
+a <- a+xlab("Number of successful routings")+ylab("Successful routings/Used space")+ggtitle("Label space usage by routings")
 
 a
+
 
 
 #plot(results$Sites.successful,results$Space.ratio, xlab="Successfully routed sites", ylab="Label area space used (%)", main="Label area usage by number of successful routings")

### 2018-04-22 Eclipse/Labeling/src/routingapp/GraphCreatorPanel.java

@@ -34,7 +34,7 @@ import org.jgrapht.graph.ListenableUndirectedWeightedGraph;
 public class GraphCreatorPanel extends JComponent {
 	
 	//DEBUG VALUES
-	private boolean testMode=false;//Toggles whether the program is in testing mode. If true, visualization is turned off, and multiple texts will be generated and routed. Overrides all other debug values.
+	private boolean testMode=true;//Toggles whether the program is in testing mode. If true, visualization is turned off, and multiple texts will be generated and routed. Overrides all other debug values.
 	private String annDistribution=	//"uniform";		//Only used if testMode==true - determines which type of annotation distribution is used for the randomized texts. 
 									//"top-left";	//Uncomment whatever value you want to use - "uniform" creates truly random annotations, whereas the others are normally distributed around a region.
 									//"top";
@@ -47,16 +47,16 @@ public class GraphCreatorPanel extends JComponent {
 									"bottom-right";
 	//The following values are all overridden if testMode==true
 	private boolean showWordBoundaries=false;//Draws rectangles around detected word boundaries in main text, if set to true
-	private boolean showGraphGrid=true;//Draws the whole routing Graph 
+	private boolean showGraphGrid=false;//Draws the whole routing Graph 
 	private boolean hideLeaders=false; //Hides the leaders and unsuccessfully routed nodes, if some other feature needs to be inspected visually
 	private boolean hideText=false;//Disables drawing the text on-screen.
 	//DEBUG VALUES END
 
 	private TextGenerator gen=new TextGenerator(0);//<--Temporary seed, will be changed before use.
 	private int numberOfTests=100, textLength=300;
-	private int annCount=10; //Values for mean and standard deviation for normally distributed annotations. annMean is also used for uniformly distributed annotations.
+	private int annCount=30; //Number of annotations placed in the generated text.
 	
-	
+	//Default text, overwritten if a generated text is used.
 	private String text=
 					"Dies ist ein Typoblindtext.\\note{This is an annotation} An ihm kann man sehen, ob alle Buchstaben da sind und wie sie aussehen. " +
 					"Manchmal benutzt man Worte wie Hamburgefonts, Rafgenduks\\note{This too} oder Handgloves, um\\note{This is an annotation with long words, like supercalifragilisticexpiralidocious} "+

### 2018-04-29 LaTeX/Paper.tex
Paper: Removed old graphics displaying time fluctuations, replaced them with a verbal description instead. Program: Added way to generate longer words in TextGenerator.java.

--- 2018-04-11 LaTeX/Paper.tex
@@
 The main drawing area handles the visualization of the results, calculates where to place the text on-screen, and also generates the routing graph, as this can be done alongside the text placement.
-\change{The routing is done in a separate class, and can easily be exchanged with another algorithm because of this.}
+The routing is done in a separate class, and can easily be exchanged with another algorithm because of this.
+
 \section{GUI}
 \label{sec:GUI}
-
 The GUI of the program is kept simple, as the main goal was to demonstrate the routing algorithm.
@@
 The control panel's first three options determine the font the program uses for the main text, whether it is bold or italicized and its size.
-\change{The annotations' text currently uses a smaller version of this font.
-The fourth and last option allows the selection of different routing algorithms, although the current options are mostly variations on the routing algorithm described in Section~\ref*{sec:AlgDesc}.} After changing any value, the program will automatically be notified of the changes and react accordingly.
+The annotations' text currently uses a smaller version of this font.
+The fourth and last option allows the selection of different routing algorithms, although the current options are mostly variations on the routing algorithm described in Section~\ref*{sec:AlgDesc}.
+After changing any value, the program will automatically be notified of the changes and react accordingly.
 
@@
 \section{Drawing Area}
-\change{As this program is a proof of concept, the drawing area itself can't be interacted with, it only serves to display the algorithm's results.
+As this program is a proof of concept, the drawing area itself can't be interacted with, it only serves to display the algorithm's results.
 If we were to adapt this program for regular use, the user should be able to manually change routings or move labels around here.
 Due to reasons described below in Section~\ref*{sec:Challenge}, each word has to be placed individually, which allows us to create the routing graph alongside the darwing process. 
-After the text has been drawn completely, the routing algorithm is invoked from here, and its results are visualized.}
+After the text has been drawn completely, the routing algorithm is invoked from here, and its results are visualized.
 
@@
 
-\change{The Routing algorithm has already been described thoroughly in Section~\ref{sec:AlgDesc}, that process is implemented by creating ordered lists of the vertices traversed by the SP-Path, along with some additional information on where to draw the P-segment located in the Routing Area.
-While the algorithm offers methods to route each annotation individually, we currently route all annotations at the same time, which vastly decreases the required communication between the algorithm and the rest of the program.}
+The Routing algorithm has already been described thoroughly in Section~\ref*{sec:AlgDesc}, that process is implemented by creating ordered lists of the vertices traversed by the SP-Path, along with some additional information on where to draw the P-segment located in the Routing Area.
+While the algorithm offers methods to route each annotation individually, we currently route all annotations at the same time, which vastly decreases the required communication between the algorithm and the rest of the program.
 
@@
 While JGraphT offers ways to display graphs, it wasn't possible to influence the placement of individual vertices.
-Our solution was forgoing these options and drawing the graph \change{unassisted}, assigning coordinates to each vertex on creation.
+Our solution was forgoing these options and drawing the graph unassisted, assigning coordinates to each vertex on creation.
 Vertices are represented by small rectangles, whereas edges are represented with straight lines.
@@
 
-%Add Test data generation/retrieval here?
-
-\section{Workflow}%<--What do I put here?
+\section{Workflow}
 
@@
 %TODO: Schematischen Graph (Flowchart?) von oben beschriebenem Ablauf erstellen!
-%As the Graph's vertices represent fixed locations on the canvas, they each have coordinates associated with them, with the vertices representing the sites containing extra information, such as references to the corresponding annotation and to the leader connecting the two (if existent).
-%The additional information stored in the vertices is used for both routing and drawing the leaders: The routing was done via a depth-first search algorithm that prioritized vertices located further up if possible, and vertices to the right otherwise.
-After the algorithm terminates, it returns  a vertex-based graph walk describing the route from the site to the border of the text area alongside information about where to draw the OPO-segment in the buffer zone if successful, or a walk containing only the starting vertex, if no valid path was found.
-Using this information alongside the positional data stored in the vertices, the polyline representing the leader can be drawn.
-
-%TODO: Schönheitsfeatures, etc.
 
@@
 Using this data, we compiled our own fixed-length testing texts by creating words containing randomized alphabetic characters, taking care that the word length distribution matched our data.
-Annotation sites are inserted into the text on either a normal or uniform distribution, and their labels contain similarly generated texts of randomized length.
+A fixed number of annotation sites is inserted into the text either by distributing their locations uniformly across the text, or by clustering them regionally.
+
+For the clustering approach, we divided the text vertically and horizontally into three subsections each - top/center/bottom and left/center/right respectively.
+To determine the site's location, we then determined the line the site will be in and its exact word by generating numbers from two normal distributions, each with a standard deviation of $1/6$th of the text's width or height, and a mean located at 25\%, 50\% or 75\% of that parameter, depending on the chosen region.
+
+The uniform distribution is achieved by choosing each annotated word in advance, regardless of its relative location within the text.
 
 To enable the inspection of individual texts, each text was generated from a randomized seed that is saved at time of creation.
+To enable the repetition of full test batches, each of them was also generated using a seed, which is recorded as well.
 
-\section{Testing methods}
+\section{Experimental Setup}
 
@@
 
+Across every test run, the first two instances always took significantly longer to complete than subsequent ones.
+As this was replicated even when using the same text for all 100 passes, we believe this to be an issue caused by the implementation rather than the algorithm itself.
+\change{Another issue was a significantly increased execution time for some instances, which could not be reproduced in subsequent tests under the same conditions and are assumed to be caused by differences in scheduling or Java's garbage collection.
+Disregarding these outliers, the worst time measured across all tests was about two milliseconds.}%Wie kann ich einbauen, dass der Computer heiß läuft?
+
+Adding more sources to the text increases the both the space usage of the labeling area and the number of successful routes, but there seems to be a limit, as increasing the amount of sources only brings marginal improvements after a certain point.
+For our experiments, this was reached at 15 sources in a 300-word text.
+Doubling their amount improved both the space usage and the amount of routed leaders for several test instances, but the overall highest amount of successful routings increased only by 1, whereas the increase from 10 to 15 sources impacted the results much stronger.
+(See Fig.~\ref*{fig:uniformstats} for details.)
+
  \begin{figure}
@@
  		\centering
-		\includegraphics[page=1, scale=0.50]{Rplots_n(5-0).pdf}
-		\caption{\label{fig:plot1_n5,0}}
+ 		\includegraphics[scale=0.8]{Rplot_10u.png}
+ 		\caption{\label{fig:10u}Results for 10 sources}
  	\end{subfigure}
@@
  		\centering
-		\includegraphics[page=1, scale=0.50]{Rplots_n(5-0)_repeat.pdf}
-		\caption{\label{fig:plot2_n5,0}}
+ 		\includegraphics[scale=0.8]{Rplot_15u.png}
+ 		\caption{\label{fig:15u}Results for 15 sources}
  	\end{subfigure}
-	\caption{Two plots of the same test seed's time taken per text. Note the massive differences caused by the computer's scheduling.}
-	\label{fig:timeplot}
+    \begin{subfigure}[b]{\textwidth}
+		 \centering
+		 \includegraphics[scale=0.8]{Rplot_30u.png}
+		 \caption{\label{fig:30u}Results for 30 sources}
+    \end{subfigure}
+
+ 	\caption{Three plots of test batches with identical seeds. \change{Each instance used a uniform distribution of sources and a 300-word text.}}
+ 	\label{fig:uniformstats}
  \end{figure}
 
-Across every test run, the first two instances always took significantly longer to complete than subsequent ones.
-As this was replicated even when using the same text for all 100 passes, we believe this to be an issue caused by the implementation rather than the algorithm itself.
-(This effect can be seen in Fig.~\ref*{fig:timeplot})
-Similarly, the biggest fluctuations in computation time for the remaining instances were caused by the computer's scheduling, as Fig.~\ref*{fig:plot1_n5,0} and Fig.~\ref*{fig:plot2_n5,0} both depict time statistics for the same run, executed twice in a row.
+While the placement of sources definitely had an impact on the amount of successful routings, the influence of horizontal placement is far smaller than that of the vertical placement.
+This is likely due to the annotations being placed vertically, which means that the more annotations are placed, the less likely it becomes for sources close to the beginning of the text to be routed.
+Together with the Routing Area, which allows for labels to be placed further up, this means that placements in the first few lines are disadvantaged compared to those further down.
 
-Unsurprisingly, the amount of labeling area space used by annotations increases with the total amount of annotations in the text, but the ratio of successful routings to total sites goes down. 
+The effects of clustering are currently minor, and mostly outweighed by the influence of vertical placement, as the average word length is short enough that there is enough whitespace in each line to avoid leaders from blocking each other.
+Increasing the word length would make this effect more severe, and it would be interesting to see the influence on the results.
 %While increasing the standard deviation for the annotation placement process also allowed some texts to reach a better label area usage, the average results became worse for that category, whereas the percentage of successful routings moves closer to the 50\% mark.

### 2018-04-29 Eclipse/Labeling/Eval_Template.r

@@ -12,11 +12,13 @@ filename <- REPLACEME
 results <- read.table(filename, sep=",", row.names=1, header=TRUE)
 results <- within(results, Sites.ratio <- Sites.successful/Sites.total*100)
 results <- within(results, Space.ratio <- Space.used/Space.total*100)
-results <- within(results, TimeToSites <- Time/Sites.total)
+results <- within(results, TimeToSites <- Routing.Time/Sites.total)
 results <- within(results, SpaceToSites <- Space.used/Sites.total)
+results <- within(results, TimeTotal <- Graph.Time+Routing.Time)
 
-plot(results[,'Time'], ylab="Time (ns)", type="b", main="Plot of time elapsed")
-boxplot(results$Time, ylab="Time (ns)", main="Time elapsed (Outliers removed)", outline=FALSE)
+plot(results[,'Graph.Time'], ylab="Time (ns)", type="b", main="Plot of time elapsed for Graph construction")
+plot(results[,'Routing.Time'], ylab="Time (ns)", type="b", main="Plot of time elapsed for Routing")
+boxplot(results$Routing.Time, ylab="Time (ns)", main="Routing: Time elapsed (Outliers removed)", outline=FALSE)
 boxplot(results$TimeToSites, main="Average time per site (Outliers removed)", outline=FALSE)
 boxplot(results$Sites.ratio, main="Percentage of successful routings")
 boxplot(results$Space.ratio, main="Percentage of Label Area usage")

### 2018-04-29 Eclipse/Labeling/src/routingapp/GraphCreatorPanel.java

@@ -34,9 +34,9 @@ import org.jgrapht.graph.ListenableUndirectedWeightedGraph;
 public class GraphCreatorPanel extends JComponent {
 	
 	//DEBUG VALUES
-	private boolean testMode=true;//Toggles whether the program is in testing mode. If true, visualization is turned off, and multiple texts will be generated and routed. Overrides all other debug values.
-	private String annDistribution=	//"uniform";		//Only used if testMode==true - determines which type of annotation distribution is used for the randomized texts. 
-									//"top-left";	//Uncomment whatever value you want to use - "uniform" creates truly random annotations, whereas the others are normally distributed around a region.
+	private boolean testMode=false;//Toggles whether the program is in testing mode. If true, visualization is turned off, and multiple texts will be generated and routed. Overrides all other debug values.
+	private String annDistribution=	"uniform";		//Only used if testMode==true - determines which type of annotation distribution is used for the randomized texts. 
+									//"top-left";		//Uncomment whatever value you want to use - "uniform" creates truly random annotations, whereas the others are normally distributed around a region.
 									//"top";
 									//"top-right";
 									//"left";
@@ -44,7 +44,7 @@ public class GraphCreatorPanel extends JComponent {
 									//"right";
 									//"bottom-left";
 									//"bottom";
-									"bottom-right";
+									//"bottom-right";
 	//The following values are all overridden if testMode==true
 	private boolean showWordBoundaries=false;//Draws rectangles around detected word boundaries in main text, if set to true
 	private boolean showGraphGrid=false;//Draws the whole routing Graph 
@@ -54,7 +54,7 @@ public class GraphCreatorPanel extends JComponent {
 
 	private TextGenerator gen=new TextGenerator(0);//<--Temporary seed, will be changed before use.
 	private int numberOfTests=100, textLength=300;
-	private int annCount=30; //Number of annotations placed in the generated text.
+	private int annCount=10; //Number of annotations placed in the generated text.
 	
 	//Default text, overwritten if a generated text is used.
 	private String text=
[…]

### 2018-04-29 Eclipse/Labeling/src/routingapp/TextGenerator.java

@@ -58,9 +58,46 @@ public class TextGenerator {
 	 * >16	0.02928564
 	 * 
 	 * Data taken from https://www.sciencedirect.com/science/article/pii/S0019995858902298 (Table III) and converted to percentages (See Tab3Calc.xslx for calculations.)
+	 * 
+	 * Custom "longer words" probability:
+	 * 1	1
+	 * 2	0.957959418
+	 * 3	0.903367849
+	 * 4	0.852755988
+	 * 5	0.737914444
+	 * 6	0.665327895
+	 * 7	0.604873967
+	 * 8	0.553205732
+	 * 9	0.405093609
+	 * 10	0.363047798
+	 * 11	0.318073423
+	 * 12	0.283662797
+	 * 13	0.21186696
+	 * 14	0.154213994
+	 * 15	0.126392637
+	 * 16	0.088571279
+	 * >16	0.02928564
 	 */
 	
-	double wordProb[]={	1,
+	double wordProb[]=/*{	1,
+	 					0.957959418,
+	 					0.903367849,
+	 					0.852755988,
+	 					0.737914444,
+	 					0.665327895,
+	 					0.604873967,
+	 					0.553205732,
+	 					0.405093609,
+	 					0.363047798,
+	 					0.318073423,
+	 					0.283662797,
+	 					0.21186696,
+	 					0.154213994,
+	 					0.126392637,
+	 					0.088571279,
+	 					0.02928564,
+						-1};
+					/*/{	1,
 						0.907959418,
 						0.803367849,
 						0.652755988,
@@ -77,7 +114,7 @@ public class TextGenerator {
 						0.086392637,
 						0.058571279,
 						0.02928564,
-						-1};//<--Actual Probability is 0, but the RNG might still generate 0.0f 
+						-1};/**///<--Actual Probability is 0, but the RNG might still generate 0.0f 
 	
 	Random rng;
 	

### 2018-04-29 LaTeX/Paper.tex
Paper: No changes. Program: Fixed bug in uniform text generation. Added 'longer words' mode. Modified R script to better label the combined histogram/boxplot.

@@
 \change{Another issue was a significantly increased execution time for some instances, which could not be reproduced in subsequent tests under the same conditions and are assumed to be caused by differences in scheduling or Java's garbage collection.
-Disregarding these outliers, the worst time measured across all tests was about two milliseconds.}%Wie kann ich einbauen, dass der Computer heiß läuft?
+Disregarding these outliers, the worst time measured across all tests was about two milliseconds.}%Wie kann ich sagen, dass der Computer heiß läuft?
 
@@
 
-The effects of clustering are currently minor, and mostly outweighed by the influence of vertical placement, as the average word length is short enough that there is enough whitespace in each line to avoid leaders from blocking each other.
-Increasing the word length would make this effect more severe, and it would be interesting to see the influence on the results.
-%While increasing the standard deviation for the annotation placement process also allowed some texts to reach a better label area usage, the average results became worse for that category, whereas the percentage of successful routings moves closer to the 50\% mark.
+The effects of clustering are minor, and mostly outweighed by the influence of vertical placement, as the average word length is short enough that there is enough whitespace in each line to avoid leaders from blocking each other.
+%Increasing the word length would make this effect more severe, and it would be interesting to see the influence on the results.
 

### 2018-04-29 Eclipse/Labeling/Eval_Template.r

@@ -26,7 +26,7 @@ a <- ggplot(results)
 a <- a+geom_bar(aes(factor(results$Sites.successful)),fill="lightgoldenrod2")
 a <- a+stat_boxplot(aes(factor(results$Sites.successful),results$Space.ratio), geom="errorbar")#Workaround to get T-shaped whiskers
 a <- a+geom_boxplot(aes(factor(results$Sites.successful),results$Space.ratio))
-a <- a+xlab("Number of successful routings")+ylab("Successful routings/Used space")+ggtitle("Label space usage by routings")
+a <- a+xlab("Number of successful routings")+ylab("Results per category              Space usage (%)       ")+ggtitle("Labeling area space usage by routings")
 
 a
 

### 2018-04-29 Eclipse/Labeling/src/routingapp/GraphCreatorPanel.java

@@ -34,8 +34,8 @@ import org.jgrapht.graph.ListenableUndirectedWeightedGraph;
 public class GraphCreatorPanel extends JComponent {
 	
 	//DEBUG VALUES
-	private boolean testMode=false;//Toggles whether the program is in testing mode. If true, visualization is turned off, and multiple texts will be generated and routed. Overrides all other debug values.
-	private String annDistribution=	"uniform";		//Only used if testMode==true - determines which type of annotation distribution is used for the randomized texts. 
+	private boolean testMode=true;//Toggles whether the program is in testing mode. If true, visualization is turned off, and multiple texts will be generated and routed. Overrides all other debug values.
+	private String annDistribution=	//"uniform";		//Only used if testMode==true - determines which type of annotation distribution is used for the randomized texts. 
 									//"top-left";		//Uncomment whatever value you want to use - "uniform" creates truly random annotations, whereas the others are normally distributed around a region.
 									//"top";
 									//"top-right";
@@ -44,7 +44,7 @@ public class GraphCreatorPanel extends JComponent {
 									//"right";
 									//"bottom-left";
 									//"bottom";
-									//"bottom-right";
+									"bottom-right";
 	//The following values are all overridden if testMode==true
 	private boolean showWordBoundaries=false;//Draws rectangles around detected word boundaries in main text, if set to true
 	private boolean showGraphGrid=false;//Draws the whole routing Graph 
@@ -189,7 +189,7 @@ public class GraphCreatorPanel extends JComponent {
 			g.setColor(Color.black);
 			
 			//Uncomment the following lines to inspect specific seeds in non-testing mode
-			gen.setSeed(-8765336103901329443L);
+			gen.setSeed(-6742867441916902193L);
 			if(!annDistribution.equals("uniform"))text=gen.generateNormalizedText(annCount, textLength, rightTextBorder-leftTextBorder, metrics, annDistribution);
 			else text=gen.generateUniformText(annCount, textLength);/**/
 			

### 2018-04-29 Eclipse/Labeling/src/routingapp/TextGenerator.java

@@ -79,7 +79,7 @@ public class TextGenerator {
 	 * >16	0.02928564
 	 */
 	
-	double wordProb[]=/*{	1,
+	double wordProb[]= /*{1,//Longer words
 	 					0.957959418,
 	 					0.903367849,
 	 					0.852755988,
@@ -97,7 +97,7 @@ public class TextGenerator {
 	 					0.088571279,
 	 					0.02928564,
 						-1};
-					/*/{	1,
+				/*/	  {	1,//Content words
 						0.907959418,
 						0.803367849,
 						0.652755988,
@@ -173,7 +173,7 @@ public class TextGenerator {
 				}
 				result=rng.nextFloat();
 			}
-			text+=word+" ";
+			text+=word;
 			
 			if(annLocations.contains(i))//generate Annotation
 			{
@@ -194,9 +194,10 @@ public class TextGenerator {
 					}
 					text+=word+" ";
 				}while(rng.nextFloat()>0.2);
-				text=text.substring(0,text.lastIndexOf(' '))+"}";
+				text=text.substring(0,text.lastIndexOf(' ')-1)+"}";
 				annLocations.remove((Integer)i);//Might not be necessary
 			}
+			text+=" ";
 			
 		}
 		return text;

### 2018-04-29 LaTeX/Paper.tex
Paper:Completed Section4, wrote Section 5 - Conclusion, added abstract. Program: No changes.

@@
 \setregnumber{1125755}
-\setdate{25}{09}{2017}%TODO: Update
+\setdate{29}{04}{2018}%TODO: Update
 
@@
 
-%\begin{abstract}
-%\todo{Enter your text here.}
-%\end{abstract}
+\begin{abstract}
+	In this paper, we present an algorithm capable of solving the one-sided boundary-labeling problem for crossing-free, monotonous OPO-Leaders in $O(n^2)$ using only the space between words and lines.
+We prove its correctness and evaluate its performance under different conditions.
+\end{abstract}
 
@@
 
-%Beispiele von suboptimalen Lösungen einfügen? (Bilder)
 Boundary labeling in text documents however, is rarely discussed, and only few papers exist about this topic.
@@
 If we were to adapt this program for regular use, the user should be able to manually change routings or move labels around here.
-Due to reasons described below in Section~\ref*{sec:Challenge}, each word has to be placed individually, which allows us to create the routing graph alongside the darwing process. 
+Due to reasons described below in Section~\ref*{sec:Challenge}, each word has to be placed individually, which allows us to create the routing graph alongside the drawing process. 
 After the text has been drawn completely, the routing algorithm is invoked from here, and its results are visualized.
@@
 We opted to use the data from the "Content Words" column, as the combined statistics almost exclusively produced words with three or less letters in our tests.
+\change{Furthermore, we created a custom distribution based on these values to test the algorithm's behavior in texts with longer words.}
 Using this data, we compiled our own fixed-length testing texts by creating words containing randomized alphabetic characters, taking care that the word length distribution matched our data.
+\change{The results of both distributions can be seen in Fig.~\ref{fig:wordlength}}
 A fixed number of annotation sites is inserted into the text either by distributing their locations uniformly across the text, or by clustering them regionally.
 
+\begin{figure}
+	\centering
+	\begin{subfigure}[b]{\textwidth}
+		\centering
+		\includegraphics[scale=0.65]{ContentWords.png}
+		\caption{A generated text using the "Content Words" distribution from Miller et al.\cite{Miller1958}.}
+	\end{subfigure}
+	~\\
+	\begin{subfigure}[b]{\textwidth}
+		\centering
+		\includegraphics[scale=0.65]{LongerWords.png}
+		\caption{A generated text using our custom distribution.}
+	\end{subfigure}
+	\label{fig:wordlength}
+	\caption{Results for both text generation distributions, with annotations.}
+\end{figure}	
 For the clustering approach, we divided the text vertically and horizontally into three subsections each - top/center/bottom and left/center/right respectively.
@@
 For our experiments, this was reached at 15 sources in a 300-word text.
-Doubling their amount improved both the space usage and the amount of routed leaders for several test instances, but the overall highest amount of successful routings increased only by 1, whereas the increase from 10 to 15 sources impacted the results much stronger.
+Doubling their amount improved both the space usage and the number of routed leaders for several test instances, but the overall highest amount of successful routings increased only by 1, whereas the increase from 10 to 15 sources impacted the results much stronger.
 (See Fig.~\ref*{fig:uniformstats} for details.)
@@
  		\centering
- 		\includegraphics[scale=0.8]{Rplot_10u.png}
+ 		\includegraphics[scale=0.6]{Rplot_10u.png}
  		\caption{\label{fig:10u}Results for 10 sources}
@@
  		\centering
- 		\includegraphics[scale=0.8]{Rplot_15u.png}
+ 		\includegraphics[scale=0.6]{Rplot_15u.png}
  		\caption{\label{fig:15u}Results for 15 sources}
@@
 		 \centering
-		 \includegraphics[scale=0.8]{Rplot_30u.png}
+		 \includegraphics[scale=0.6]{Rplot_30u.png}
 		 \caption{\label{fig:30u}Results for 30 sources}
@@
 
-The effects of clustering are minor, and mostly outweighed by the influence of vertical placement, as the average word length is short enough that there is enough whitespace in each line to avoid leaders from blocking each other.
+The effects of clustering are minor, and mostly outweighed by the influence of vertical placement, as the average word length is short enough that there is sufficient whitespace in each line to avoid leaders from blocking each other.
+
+\newpage
+\change{Re-running the tests with an increased word length showed several changes: Most noticable was a higher variance in results - whereas the histogram in Fig.~\ref*{fig:15u} has a noticable peak at 10 successful routings, the values in Fig.~\ref*{fig:15uLong} are more spread out, reaching both better and worse end results.
+The bottom left clustering also no longer achieves obviously better results than other vertex placements.
+While Fig.~\ref*{fig:15blLong} has less results with nine or less successful routings then Fig.~\ref*{fig:15uLong}, it also has less results in the 12+ range.
+Finally, the vertical placement of vertices has a noticable impact - Fig.~\ref*{fig:15blLong} contains less results in below 10 successful routings than Fig.~\ref*{fig:15brLong}, and even has some instances with 13 successes.}
+
+\begin{figure}
+	\centering
+	\begin{subfigure}[b]{\textwidth}
+		\centering
+		\includegraphics[scale=0.6]{Rplot_15u_longer.png}
+		\caption{\label{fig:15uLong}Results for uniform distribution.}
+	\end{subfigure}
+	\begin{subfigure}[b]{\textwidth}
+		\centering
+		\includegraphics[scale=0.6]{Rplot_15bl_longer.png}
+		\caption{\label{fig:15blLong}Results for clustered distribution in the bottom left.}
+	\end{subfigure}
+	\begin{subfigure}[b]{\textwidth}
+		\centering
+		\includegraphics[scale=0.6]{Rplot_15br_longer.png}
+		\caption{\label{fig:15brLong}Results for clustered distribution in the bottom right.}
+	\end{subfigure}
+	\caption{Test results for instances with 15 differently distributed sources over a 300-word text with increased word length.}
+	\label{fig:longTests}
+\end{figure}
 %Increasing the word length would make this effect more severe, and it would be interesting to see the influence on the results.
@@
 
+\change{In this paper we presented an algorithm capable of solving the one-sided boundary labeling problem for monotonous, crossing-free OPO-Leaders using only the whitespace between words and lines.
+We have shown it to run in $O(n^2)$ and proven its correctness.
+Finally, we have evaluated its performance under different conditions.}
+
 \section{Further notes}
 
-%Section Ideas: The Program/Framework (Modellerklärung), The Algorithm(s), Implementation, Evaluation, Conclusion
+\change{An obvious weakness of our algorithm is its inability to create routings for source nodes above the next free annotation positions.
+Adding leaders that are only allowed to go down instead of up would allow for these positions to be routed, but would require extra precautions to avoid intersecting leaders.
+Alternatively, leaders could be restricted by how many P-Segments can be included before reaching the labeling area to make them easier to follow.}
 

### 2018-04-29 Eclipse/Labeling/src/routingapp/GraphCreatorPanel.java

@@ -35,8 +35,8 @@ public class GraphCreatorPanel extends JComponent {
 	
 	//DEBUG VALUES
 	private boolean testMode=true;//Toggles whether the program is in testing mode. If true, visualization is turned off, and multiple texts will be generated and routed. Overrides all other debug values.
-	private String annDistribution=	//"uniform";		//Only used if testMode==true - determines which type of annotation distribution is used for the randomized texts. 
-									//"top-left";		//Uncomment whatever value you want to use - "uniform" creates truly random annotations, whereas the others are normally distributed around a region.
+	private String annDistribution=	"uniform";		//Only used if testMode==true - determines which type of annotation distribution is used for the randomized texts. 
+									//"top-left";	//Uncomment whatever value you want to use - "uniform" creates truly random annotations, whereas the others are normally distributed around a region.
 									//"top";
 									//"top-right";
 									//"left";
@@ -44,7 +44,7 @@ public class GraphCreatorPanel extends JComponent {
 									//"right";
 									//"bottom-left";
 									//"bottom";
-									"bottom-right";
+									//"bottom-right";
 	//The following values are all overridden if testMode==true
 	private boolean showWordBoundaries=false;//Draws rectangles around detected word boundaries in main text, if set to true
 	private boolean showGraphGrid=false;//Draws the whole routing Graph 
@@ -54,7 +54,7 @@ public class GraphCreatorPanel extends JComponent {
 
 	private TextGenerator gen=new TextGenerator(0);//<--Temporary seed, will be changed before use.
 	private int numberOfTests=100, textLength=300;
-	private int annCount=10; //Number of annotations placed in the generated text.
+	private int annCount=15; //Number of annotations placed in the generated text.
 	
 	//Default text, overwritten if a generated text is used.
 	private String text=

### 2018-06-10 LaTeX/Paper.tex
Paper: Style and presentation changes. Program: No changes.

@@
 \mainmatter
-
-%TODO: Abstract: \begin{abstract}...\end{abstract}
 
@@
 
-\section{GUI}
-\label{sec:GUI}
+\section{User Interface}
+\label{sec:UI}
 The GUI of the program is kept simple, as the main goal was to demonstrate the routing algorithm.
@@
 
-If any of the parameters on the control panel (see section~\ref*{sec:GUI}) are changed, the process re-starts from the beginning.
+If any of the parameters on the control panel (see section~\ref*{sec:UI}) are changed, the process re-starts from the beginning.
 
@@
 We opted to use the data from the "Content Words" column, as the combined statistics almost exclusively produced words with three or less letters in our tests.
-\change{Furthermore, we created a custom distribution based on these values to test the algorithm's behavior in texts with longer words.}
+Furthermore, we created a custom distribution based on these values to test the algorithm's behavior in texts with longer words.
 Using this data, we compiled our own fixed-length testing texts by creating words containing randomized alphabetic characters, taking care that the word length distribution matched our data.
-\change{The results of both distributions can be seen in Fig.~\ref{fig:wordlength}}
+The results of both distributions can be seen in Fig.~\ref{fig:wordlength}
 A fixed number of annotation sites is inserted into the text either by distributing their locations uniformly across the text, or by clustering them regionally.
@@
 As this was replicated even when using the same text for all 100 passes, we believe this to be an issue caused by the implementation rather than the algorithm itself.
-\change{Another issue was a significantly increased execution time for some instances, which could not be reproduced in subsequent tests under the same conditions and are assumed to be caused by differences in scheduling or Java's garbage collection.
-Disregarding these outliers, the worst time measured across all tests was about two milliseconds.}%Wie kann ich sagen, dass der Computer heiß läuft?
+Another issue was a significantly increased execution time for some instances, which could not be reproduced in subsequent tests under the same conditions and are assumed to be caused by differences in scheduling or Java's garbage collection.
+Disregarding these outliers, the worst time measured across all tests was about two milliseconds.%Wie kann ich sagen, dass der Computer heiß läuft?
 
@@
  	\centering
- 	\begin{subfigure}[b]{\textwidth}
+ 	\begin{subfigure}[b]{\textwidth*2/7}
  		\centering
- 		\includegraphics[scale=0.6]{Rplot_10u.png}
+ 		\includegraphics[scale=0.35]{Rplot_10u.png}
  		\caption{\label{fig:10u}Results for 10 sources}
  	\end{subfigure}
-
- 	\begin{subfigure}[b]{\textwidth}
+ 	\begin{subfigure}[b]{\textwidth*2/7}
  		\centering
- 		\includegraphics[scale=0.6]{Rplot_15u.png}
+ 		\includegraphics[scale=0.35]{Rplot_15u.png}
  		\caption{\label{fig:15u}Results for 15 sources}
  	\end{subfigure}
-    \begin{subfigure}[b]{\textwidth}
+    \begin{subfigure}[b]{\textwidth*2/7}
 		 \centering
-		 \includegraphics[scale=0.6]{Rplot_30u.png}
+		 \includegraphics[scale=0.35]{Rplot_30u.png}
 		 \caption{\label{fig:30u}Results for 30 sources}
@@
 
- 	\caption{Three plots of test batches with identical seeds. \change{Each instance used a uniform distribution of sources and a 300-word text.}}
+ 	\caption{Three plots of test batches with identical seeds. Each instance used a uniform distribution of sources and a 300-word text.}
  	\label{fig:uniformstats}
@@
 
-\newpage
-\change{Re-running the tests with an increased word length showed several changes: Most noticable was a higher variance in results - whereas the histogram in Fig.~\ref*{fig:15u} has a noticable peak at 10 successful routings, the values in Fig.~\ref*{fig:15uLong} are more spread out, reaching both better and worse end results.
+Re-running the tests with an increased word length showed several changes: Most noticable was a higher variance in results - whereas the histogram in Fig.~\ref*{fig:15u} has a noticable peak at 10 successful routings, the values in Fig.~\ref*{fig:15uLong} are more spread out, reaching both better and worse end results.
 The bottom left clustering also no longer achieves obviously better results than other vertex placements.
 While Fig.~\ref*{fig:15blLong} has less results with nine or less successful routings then Fig.~\ref*{fig:15uLong}, it also has less results in the 12+ range.
-Finally, the vertical placement of vertices has a noticable impact - Fig.~\ref*{fig:15blLong} contains less results in below 10 successful routings than Fig.~\ref*{fig:15brLong}, and even has some instances with 13 successes.}
+Finally, the vertical placement of vertices has a noticable impact - Fig.~\ref*{fig:15blLong} contains less results in below 10 successful routings than Fig.~\ref*{fig:15brLong}, and even has some instances with 13 successes.
 
@@
 	\centering
-	\begin{subfigure}[b]{\textwidth}
+	\begin{subfigure}[b]{\textwidth*5/11}
 		\centering
-		\includegraphics[scale=0.6]{Rplot_15u_longer.png}
+		\includegraphics[scale=0.55]{Rplot_15u_longer.png}
 		\caption{\label{fig:15uLong}Results for uniform distribution.}
 	\end{subfigure}
-	\begin{subfigure}[b]{\textwidth}
+
+	\begin{subfigure}[b]{\textwidth*5/11}
 		\centering
-		\includegraphics[scale=0.6]{Rplot_15bl_longer.png}
+		\includegraphics[scale=0.55]{Rplot_15bl_longer.png}
 		\caption{\label{fig:15blLong}Results for clustered distribution in the bottom left.}
-	\end{subfigure}
-	\begin{subfigure}[b]{\textwidth}
+	\end{subfigure}~~%This is here to increase the space between the figures
+	\begin{subfigure}[b]{\textwidth*5/11}
 		\centering
-		\includegraphics[scale=0.6]{Rplot_15br_longer.png}
+		\includegraphics[scale=0.55]{Rplot_15br_longer.png}
 		\caption{\label{fig:15brLong}Results for clustered distribution in the bottom right.}
@@
 
-\change{In this paper we presented an algorithm capable of solving the one-sided boundary labeling problem for monotonous, crossing-free OPO-Leaders using only the whitespace between words and lines.
+In this paper we presented an algorithm capable of solving the one-sided boundary labeling problem for monotonous, crossing-free OPO-Leaders using only the whitespace between words and lines.
 We have shown it to run in $O(n^2)$ and proven its correctness.
-Finally, we have evaluated its performance under different conditions.}
+Finally, we have evaluated its performance under different conditions.
 
-\section{Further notes}
-
-\change{An obvious weakness of our algorithm is its inability to create routings for source nodes above the next free annotation positions.
+An obvious weakness of our algorithm is its inability to create routings for source nodes above the next free annotation positions.
 Adding leaders that are only allowed to go down instead of up would allow for these positions to be routed, but would require extra precautions to avoid intersecting leaders.
-Alternatively, leaders could be restricted by how many P-Segments can be included before reaching the labeling area to make them easier to follow.}
+Alternatively, leaders could be restricted by how many P-Segments can be included before reaching the labeling area to make them easier to follow.
 

### 2018-06-17 LaTeX/Paper.tex
Paper: Edited Abstract. Program: No changes.

@@
 \begin{abstract}
-	In this paper, we present an algorithm capable of solving the one-sided boundary-labeling problem for crossing-free, monotonous OPO-Leaders in $O(n^2)$ using only the space between words and lines.
+	The applications of Boundary Labeling in tests are largely unexplored and existing solutions often impact readability or produce ambiguous labelings.
+
+	In this paper, we present an algorithm that produces crossing-free, monotonous OPO-Leaders and solves the one-sided boundary-labeling problem in $O(n^2)$.
+	We also prove its correctness and evaluate the performance under different circumstances.
+	%In this paper, we present an algorithm capable of solving the one-sided boundary-labeling problem for crossing-free, monotonous OPO-Leaders in $O(n^2)$ using only the space between words and lines.
 We prove its correctness and evaluate its performance under different conditions.

### 2018-06-21 LaTeX/Paper.tex
Paper: Expanded Sec. 4.3, added Lists of Figures and Algorithms. Program: No changes.

@@
 Vertices are represented by small rectangles, whereas edges are represented with straight lines.
-To avoid confusion between crossing leaders and leaders that incorporate the same vertex in their path, but curve away from each other, all leaders are drawn with rounded corners, which can be seen in Fig.~\ref{fig:ProgWindow}.
+To avoid confusion between crossing leaders and leaders that incorporate the same vertex in their path, but curve away from each other, all leaders are drawn with rounded corners, which can be seen in Fig.~\ref*{fig:ProgWindow}.
 
@@
 Using this data, we compiled our own fixed-length testing texts by creating words containing randomized alphabetic characters, taking care that the word length distribution matched our data.
-The results of both distributions can be seen in Fig.~\ref{fig:wordlength}
+The results of both distributions can be seen in Fig.~\ref*{fig:wordlength}
 A fixed number of annotation sites is inserted into the text either by distributing their locations uniformly across the text, or by clustering them regionally.
@@
 	\end{subfigure}
-	\label{fig:wordlength}
-	\caption{Results for both text generation distributions, with annotations.}
+	\caption{\label{fig:wordlength}Results for both text generation distributions, with annotations.}
 \end{figure}	
@@
 To enable the inspection of individual texts, each text was generated from a randomized seed that is saved at time of creation.
-To enable the repetition of full test batches, each of them was also generated using a seed, which is recorded as well.
+To facilitate the repetition of full test batches, each of them was also generated using a seed, which is recorded as well.
 
@@
 Another issue was a significantly increased execution time for some instances, which could not be reproduced in subsequent tests under the same conditions and are assumed to be caused by differences in scheduling or Java's garbage collection.
-Disregarding these outliers, the worst time measured across all tests was about two milliseconds.%Wie kann ich sagen, dass der Computer heiß läuft?
+Disregarding these outliers, the worst time measured across all tests was about two milliseconds, which allows us to display the routings almost instantly or run sequences of multiple tests very quickly.
 
@@
 For our experiments, this was reached at 15 sources in a 300-word text.
-Doubling their amount improved both the space usage and the number of routed leaders for several test instances, but the overall highest amount of successful routings increased only by 1, whereas the increase from 10 to 15 sources impacted the results much stronger.
+Doubling them improved both the space usage and the average number of routed leaders over all test instances, but the overall highest amount of successful routings increased only slightly, whereas the increase from 10 to 15 sources impacted the results much stronger.
+This shows that there is an upper bound to the amount of successful routings in our tests, and that the amount of routed leaders increase less than linearly with the number of available sources.
 (See Fig.~\ref*{fig:uniformstats} for details.)
@@
  		\centering
- 		\includegraphics[scale=0.35]{Rplot_10u.png}
+ 		\includegraphics[scale=0.36]{Rplot_10u.png}
  		\caption{\label{fig:10u}Results for 10 sources}
@@
  		\centering
- 		\includegraphics[scale=0.35]{Rplot_15u.png}
+ 		\includegraphics[scale=0.36]{Rplot_15u.png}
  		\caption{\label{fig:15u}Results for 15 sources}
@@
 		 \centering
-		 \includegraphics[scale=0.35]{Rplot_30u.png}
+		 \includegraphics[scale=0.36]{Rplot_30u.png}
 		 \caption{\label{fig:30u}Results for 30 sources}
@@
 This is likely due to the annotations being placed vertically, which means that the more annotations are placed, the less likely it becomes for sources close to the beginning of the text to be routed.
-Together with the Routing Area, which allows for labels to be placed further up, this means that placements in the first few lines are disadvantaged compared to those further down.
+Together with the Routing Area, which allows for labels to be placed further up, but not down, this means that placements in the first few lines are disadvantaged compared to those below them.
 
@@
 
-Re-running the tests with an increased word length showed several changes: Most noticable was a higher variance in results - whereas the histogram in Fig.~\ref*{fig:15u} has a noticable peak at 10 successful routings, the values in Fig.~\ref*{fig:15uLong} are more spread out, reaching both better and worse end results.
-The bottom left clustering also no longer achieves obviously better results than other vertex placements.
-While Fig.~\ref*{fig:15blLong} has less results with nine or less successful routings then Fig.~\ref*{fig:15uLong}, it also has less results in the 12+ range.
-Finally, the vertical placement of vertices has a noticable impact - Fig.~\ref*{fig:15blLong} contains less results in below 10 successful routings than Fig.~\ref*{fig:15brLong}, and even has some instances with 13 successes.
+Re-running the tests with an increased word length showed several changes: Most noticeable was a higher variance in results - whereas the histogram of uniform source placements in Fig.~\ref*{fig:15u} has a noticeable peak at 10 successful routings, the values in Fig.~\ref*{fig:15uLong}, which depicts the results of a bottom-left clustering, are more spread out, reaching both better and worse end results.
+This tells us that while there are more overall paths to route a source placed in the bottom left, the dense clustering of multiple sources quickly exhausts the possible routings away from the cluster, which lowers the maximum amount of successful routings.
+
+The bottom left clustering also no longer achieves obviously better results than other vertex placements - while Fig.~\ref*{fig:15blLong} has less results with nine or less successful routings then Fig.~\ref*{fig:15uLong}, it also has less results in the 12+ range.
+
+Finally, the vertical placement of vertices has a noticeable impact now - Fig.~\ref*{fig:15blLong} contains less results in below 10 successful routings than the bottom-right clusterings depicted in Fig.~\ref*{fig:15brLong}, and even has some instances with 13 successes.
+
+Therefore, text styles or languages that frequently use long words benefit from spread-out source placements, whereas those with shorter average word length allow for more sources to be placed where the algorithm has the most options for routing individual sources, disregarding any previous leaders.
 
@@
 An obvious weakness of our algorithm is its inability to create routings for source nodes above the next free annotation positions.
-Adding leaders that are only allowed to go down instead of up would allow for these positions to be routed, but would require extra precautions to avoid intersecting leaders.
+Adding leaders that are only allowed to go down instead of up would allow for these positions to be routed, but would require extra precautions to avoid intersections.
 Alternatively, leaders could be restricted by how many P-Segments can be included before reaching the labeling area to make them easier to follow.
 
-\backmatter %TODO: Es gibt mehr nützliche Kommandos in VUTINFTH für zb Listen von Bildern, etc.!
+\backmatter
+
+\listoffigures
+
+\listofalgorithms
 
